### 2016-05-12 Paper_v1.tex
Initial commit

%% BioMed_Central_Tex_Template_v1.06
%%                                      %
%  bmc_article.tex            ver: 1.06 %
%                                       %

%%IMPORTANT: do not delete the first line of this template
%%It must be present to enable the BMC Submission system to
%%recognise this template!!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                     %%
%%  LaTeX template for BioMed Central  %%
%%     journal article submissions     %%
%%                                     %%
%%          <8 June 2012>              %%
%%                                     %%
%%                                     %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% For instructions on how to fill out this Tex template           %%
%% document please refer to Readme.html and the instructions for   %%
%% authors page on the biomed central website                      %%
%% http://www.biomedcentral.com/info/authors/                      %%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%% BioMed Central currently use the MikTex distribution of         %%
%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
%% http://www.miktex.org                                           %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% additional documentclass options:
%  [doublespacing]
%  [linenumbers]   - put the line numbers on margins

%%% loading packages, author definitions

%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
\documentclass{bmcart}

%%% Load packages
\usepackage{amsthm,amsmath}
\usepackage{graphicx}
\usepackage[margin=0.1pt,font=footnotesize,labelfont=bf]{caption}

% Imports required for p-code of JuPOETs
\usepackage[linesnumbered]{algorithm2e}
%\usepackage{algorithm}
%\RequirePackage{natbib}
%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
%\RequirePackage{hyperref}
\usepackage[utf8]{inputenc} %unicode support
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                             %%
%%  If you wish to display your graphics for   %%
%%  your own use using includegraphic or       %%
%%  includegraphics, then comment out the      %%
%%  following two lines of code.               %%
%%  NB: These line *must* be included when     %%
%%  submitting to BMC.                         %%
%%  All figure files must be submitted as      %%
%%  separate graphics through the BMC          %%
%%  submission process, not included in the    %%
%%  submitted article.                         %%
%%                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\def\includegraphic{}
\def\includegraphics{}

%%% Put your definitions there:
\startlocaldefs
\endlocaldefs

%%% Begin ...
\begin{document}

%%% Start of article front matter
\begin{frontmatter}

\begin{fmbox}
\dochead{Software}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the title of your article here     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{JuPOETs: A Constrained Multiobjective Optimization Approach to Estimate Biochemical Model Ensembles in the Julia Programming Language}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors here                   %%
%%                                          %%
%% Specify information, if available,       %%
%% in the form:                             %%
%%   <key>={<id1>,<id2>}                    %%
%%   <key>=                                 %%
%% Comment or delete the keys which are     %%
%% not used. Repeat \author command as much %%
%% as required.                             %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author[
   addressref={aff2},
   email={dmb457@cornell.edu}
]{\inits{DB}\fnm{David} \snm{Bassen}}
\author[
   addressref={aff2},
   email={jtb47@cornell.edu}
]{\inits{JB}\fnm{Jonathan T} \snm{Butcher}}
\author[
   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
   corref={aff1},                       % id of corresponding address, if any
   %noteref={n1},                        % id's of article notes, if any
   email={jdv27@cornell.edu}   % email address
]{\inits{JV}\fnm{Jeffrey D} \snm{Varner}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors' addresses here        %%
%%                                          %%
%% Repeat \address commands as much as      %%
%% required.                                %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\address[id=aff1]{%                           % unique id
  \orgname{Department of Chemical and Biomolecular Engineering}, % university, etc
  \street{Cornell University},                     %
  \postcode{14806}                              % post or zip code
  \city{Ithaca NY},                             % city
  \cny{USA}                                    % country
}
\address[id=aff2]{%
  \orgname{Department of Biomedical Engineering},
  \street{Cornell University},                     %
  \postcode{14806}                              % post or zip code
  \city{Ithaca NY},                             % city
  \cny{USA}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter short notes here                   %%
%%                                          %%
%% Short notes will be after addresses      %%
%% on first page.                           %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{artnotes}
%\note{Sample of title note}     % note to the article
%\note[id=n1]{Equal contributor} % note, connected to author
\end{artnotes}

\end{fmbox}% comment this for two column layout

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Abstract begins here                 %%
%%                                          %%
%% Please refer to the Instructions for     %%
%% authors on http://www.biomedcentral.com  %%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstractbox}

\begin{abstract} % abstract
\parttitle{First part title} %if any
Text for this section.

\parttitle{Second part title} %if any
Text for this section.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The keywords begin here                  %%
%%                                          %%
%% Put each keyword in separate \kwd{}.     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{keyword}
\kwd{sample}
\kwd{article}
\kwd{author}
\end{keyword}

% MSC classifications codes, if any
%\begin{keyword}[class=AMS]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

\end{abstractbox}
%
%\end{fmbox}% uncomment this for twcolumn layout

\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Main Body begins here                %%
%%                                          %%
%% Please refer to the instructions for     %%
%% authors on:                              %%
%% http://www.biomedcentral.com/info/authors%%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%% See the Results and Discussion section   %%
%% for details on how to create sub-sections%%
%%                                          %%
%% use \cite{...} to cite references        %%
%%  \cite{koon} and                         %%
%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
% <put your article body there>

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section*{Background}
Models of signal transduction networks may exhibit complex relationships between model performance and parameter values \cite{Brown:2004}.
It is often not possible to uniquely identify parameters from noisy experimental measurements, even when given extensive training data \cite{Gadkar:2005ad}.
Thus, despite significant advances in standardizing model identification \cite{Gennemark:2009on},
the core problem of estimating model parameters from noisy data remains.
Ensemble approaches have been used to address parameter uncertainty in systems biology and other fields like weather prediction \cite{Battogtokh:2002,Kuepfer:2007, Brown:2003,Palmer:2005}.
In an ensemble approach, a \emph{family} of approximate parameter sets is identified instead of a single best-fit parameter set.
Parameter families can be selected based on simulation error, along with other criterion such as diversity.
Simulations using parameter ensembles can estimate confidence intervals on model variables, and robustly constrain model predictions,
despite having many poorly constrained parameters \cite{Gutenkunst2007,Song:2009b}.
There are several techniques to generate parameter ensembles.
Battogtokh et al., Brown et al., and later Tasseff et al. generated experimentally constrained parameter ensembles using a
Metropolis-type random walk through parameter space \cite{Battogtokh:2002,Brown:2003,Tasseff:2010aa,Tasseff:2011aa}.
Other strategies could also be adapted to generate parameter ensembles \cite{Moles:2003}.
However, the unifying component of all of these previous strategies was the minimization of a single objective function.
Identification of large-scale models with hundreds, thousands or tens of thousands states and parameters may not be tractable with single objective formulations.
Models at these scales require significant training data perhaps taken from diverse sources, for example different laboratories or cell-lines.
These diverse training data would likely be heterogenous and could contain intrinsic conflicts.
Another strategy to formulate large-scale models is to integrate many small pathway modules into a single large model.
For example, Diamond and coworkers developed a model of phosphoinositide and calcium signaling in human platelets composed of many
separate submodels identified using different types of training data \cite{Purvis:2008}.
Taken together, the problem of identifying parameter ensembles for large-scale models with many parameters posses a unique challenge.
Parameter ensemble estimation techniques which optimally balance tradeoffs between submodels or conflicts in training data
might lead to more robust model performance. One class of such techniques is multiobjective optimization.
Previously, we developed the Pareto Optimal Ensemble Technique (POETs) algorithm to address the challenge of competing or conflicting objectives.
POETs, which integrates simulated annealing and multiobjective optimization through the notion of Pareto rank, finds a family of parameter estimates which optimally trade-off between
competing (and potentially conflicting) experimental objectives \cite{Song:2010aa}.
POETs has been used in several biochemical and signal transduction studies to estimate parameter ensembles \cite{Song:2009aa,Lequieu:2011aa}.
However, the previous implementation of POETs, in the Octave programming language \cite{Octave}, suffered from poor performance, and did not offer user definable
functions such as custom cooling schedules, parameter constraints, or custom search logic. It was also not well integrated into a modern package or
source code management (SCM) system.

\section*{Implementation}

In this study, we present JuPOETs, an updated version of the Pareto Optimal Ensemble Technique (POETs) ported to Julia.
JuPOETs offers many advantages and improvements compared to the previous implementation of POETs.
First, JuPOETs takes advantage of the unique performance features of the Julia programming language.
Julia, which has performance comparable to C but with syntax similar to MATLAB/Octave and Python, is a cross-platform, high-performance programming language for technical computing \cite{Julia}. Julia offers a sophisticated compiler, distributed parallel execution, numerical accuracy, and an extensive mathematical function library.
Additionally, Julia offers a built-in package manager which is directly integrated with GitHub, a popular web-based Git repository hosting service which offers distributed revision control and source code management.
Next, because Julia can natively call other languages such as Python or C, JuPOETs can be used with models implemented in a variety of languages on many platforms.
Lastly, the architecture of JuPOETs takes advantage of the first-class function type in Julia allowing user definable behavior for all key aspects of the algorithm, including objective functions, custom search logic, linear/non-linear parameter constraints (and parameter bounds constraints) as well as custom cooling schedule functions. Thus, JuPOETs can easily be adapted to solve many problem types, including mixed binary and continuous variable types, without the need to change the base algorithm (which was not true of the previous POETs implementation).

\subsection*{Optimization problem formulation.}
JuPOETs solves the $\mathcal{K}-$dimensional constrained multiobjective optimization problem:
\begin{equation}
\min_{\mathbf{p}}
\begin{cases}
  O_{1}\left(\mathbf{x}(t,\mathbf{p}),\mathbf{p}\right) \\
  O_{2}\left(\mathbf{x}(t,\mathbf{p}),\mathbf{p}\right) \\
  \vdots & \\
  O_{\mathcal{K}}\left(\mathbf{x}(t,\mathbf{p}),\mathbf{p}\right) \\
\end{cases}
\end{equation}
subject to:
\begin{eqnarray}\nonumber
  \frac{d\mathbf{x}}{dt} & = & \mathbf{f}(t,\mathbf{x}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p}) \\\nonumber
  g_{1}\left(t,\mathbf{x}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p}\right) &\geq& 0 \\\nonumber
  g_{2}\left(t,\mathbf{x}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p}\right) &\geq& 0 \\\nonumber
  &\vdots& \\\nonumber
  g_{\mathcal{C}}\left(t,\mathbf{x}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p}\right) &\geq& 0
\end{eqnarray}and parameter bound constraints:
\begin{equation}\nonumber
  \mathcal{L} \leq \mathbf{p} \leq \mathcal{U}
\end{equation}
The term t denotes time, $g_i\left(t,\mathbf{x}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p}\right)$ denotes the ith constraint on the search, where
$\mathbf{x}\left(t,\mathbf{p}\right)$ denotes the model state variable vector with an initial state $\mathbf{x}_{0}$, and $\mathbf{u}(t)$ is an input vector.
The terms $\mathbf{f}(t,\mathbf{x}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p})$ denote the system of model equations (e.g., differential equations,
differential algebraic equations or algebraic constraints) and $\mathbf{p}$ denotes the unknown parameter vector ($\mathcal{P}\times~1$).
The parameter search can be subject to parameter bound constraints where $\mathcal{L}$ and $\mathcal{U}$ denote the lower and upper parameter bounds, respectively.

POETs integrate Simulated Annealing (SA) with Pareto optimality to estimate parameter sets on or near the optimal tradeoff surface between competing training objectives (Fig. \ref{fig-MOSAalgorithm}).
The central idea underlying POETs is a mapping between the value of the objective functions evaluated at $\mathbf{p}_{i+1}$ (parameter guess at iteration $i+1$) and the Pareto rank.
We computed the Pareto rank of $\mathbf{p}_{i+1}$ by comparing the simulation error at iteration $i+1$ against an archive of simulation errors compiled up to iteration $i$ $\mathbf{E}_{i}$.
We used the Fonseca and Fleming scheme to compute the Pareto rank, $r$ \cite{RANKING}:
\begin{equation}\label{eqn_rank}
rank\left(\mathbf{p}_{i+1}\mid \mathbf{E}_{i}\right)=r
\end{equation} where $r$ denotes the rank, or the number of parameter sets that dominate parameter set $\mathbf{p}_{i+1}$.
Parameter sets on the optimal trade-off surface between the objectives have a rank equal to $0$ (no other current parameter sets are better).
Sets with increasing non-zero rank are progressively further away from the optimal trade-off surface.
Parameter sets on the optimal trade-off surface have a rank equal to $0$ (no other current parameter sets are better).
Sets with increasing non-zero rank are progressively farther away from the optimal trade-off surface.
Thus, a parameter set with a $rank = 0$ is \emph{better} in a trade-off sense than $rank>0$.
We used the Pareto rank to inform the SA calculation.
The parameter set $\mathbf{k}_{i+1}$ was accepted or rejected by the SA using the acceptance probability $\mathcal{P}\left(\mathbf{k}_{i+1}\right)$:
\begin{equation}\label{eqn_costMOSA}
\mathcal{P}(\mathbf{k}_{i+1}) \equiv \exp{\{-rank\left(\mathbf{k}_{i+1} \mid \mathbf{K}_{i} \right)/T\}}
\end{equation}
where $T$ is the computational annealing temperature. As $rank\left(\mathbf{k}_{i+1}\mid \mathbf{K}_{i}\right)\rightarrow{0}$, the acceptance probability moved toward one,
ensuring that we explored parameter sets along the Pareto surface.
Occasionally (depending upon $T$) a parameter set with a high Pareto rank was accepted by the SA allowing a more diverse search of the parameter space.
However, as $T$ was reduced, the probability of this occurring decreased.
Parameter sets could be accepted by the SA and $\emph{not}$ archived in $\mathbf{K}_{i}$. Only
parameter sets with $rank\leq{2}$ were included in $\mathbf{K}_{i}$ to ensure that we characterized the neighborhood near the trade-off surface.
The general scheme of JuPOETs followed conventional simulated annealing except that the current archive of Pareto-optimal ensemble was updated whenever a new proposal solution is accepted JuPOETs expands upon the original POETs implementation by using the function type in Julia to allow user defined functions for all major components of the approach (Algorithm \ref{algo:JuPOETs-pcode}).

\begin{algorithm}

  \SetKwFunction{objective}{objective}
  \SetKwFunction{neighbor}{neighbor}
  \SetKwFunction{acceptance}{acceptance}
  \SetKwFunction{rank}{rank}

  \SetKwInOut{Input}{input}
  \SetKwInOut{Output}{Output}

  \Input{User specified neighbor, objective, acceptance and cooling functions. Initial parameter guess ($\mathcal{P}\times~1$)}
  \Output{Rank archive $\mathcal{R}$, solution archive $\mathcal{S}$ and objective archive $\mathcal{O}$}

  \BlankLine
  \BlankLine
  initialize: $\mathcal{R}$, $\mathcal{S}$ and $\mathcal{O}$ using initial guess\;
  initialize: T $\leftarrow$1.0\;
  initialize: $T_{min}\leftarrow$1/10000\;
  initialize: Maximum number of steps per temperature $\mathcal{I}$\;

  \BlankLine
  \BlankLine
 \While{$T>T_{min}$}{
    i $\leftarrow$  1\;
    \While{i$<\mathcal{I}$}{
      \BlankLine
      \BlankLine
      \tcp{Generate a new parameter solution using user neighbor function}
      $\mathbf{p}_{i+1}$ $\leftarrow$ user-function::\neighbor{$\mathbf{p}^{*}$}\;
      \BlankLine
      \BlankLine
      \tcp{Evaluate $\mathbf{p}_{i+1}$ using user objective function}
      $\mathbf{o}_{i+1}$ $\leftarrow$ user-function::\objective{$\mathbf{p}_{i+1}$}\;
      \BlankLine
      \BlankLine
      Add $\mathbf{p}_{i+1}$ to solution archive $\mathcal{S}$\;
      Add $\mathbf{o}_{i+1}$ to objective archive $\mathcal{O}$\;
      \BlankLine
      \BlankLine
      \tcp{Calculate Pareto rank of solutions in $\mathcal{O}$ using builtin rank function}
      $\mathcal{R}\leftarrow$ builtin-function::\rank{$\mathcal{O}$}\;

      \BlankLine
      \tcp{Accept $\mathbf{p}_{i+1}$ into the archive with user defined probability}
      $\mathcal{P}$ $\leftarrow$ user-function::\acceptance{$\mathcal{R}$,$T$}\;
      \eIf {$\mathcal{P}>$rand}{

          \BlankLine
          \tcp{Update the best solution with $\mathbf{p}_{i+1}$}
          $\mathbf{p}^{*}$ $\leftarrow$ $\mathbf{p}_{i+1}$\;
          prune $\mathcal{S}$, $\mathcal{R}$ and $\mathcal{O}$ of all solutions above a rank threshold\;
      }{
        Remove $\mathbf{p}_{i+1}$ from solution archive $\mathcal{S}$\;
        Remove $\mathbf{o}_{i+1}$ from error archive $\mathcal{O}$\;
      }

      \BlankLine
      i $\leftarrow$  i + 1\;
    }

    \BlankLine
    \tcp{Update $T$ using the user cooling function}
    $T$ $\leftarrow$ user-function::cooling($T$)\;
 }
\vspace{0.1in}
\caption{Pseudo-code for the run-loop of the JuPOETs algorithm. }\label{algo:JuPOETs-pcode}
\end{algorithm}

\section*{Results and Discussion}
We compared the performance of JuPOETs versus POETs implemented in Octave for a suite of multiobjective test functions (Table ZZ).
In all cases, the same cooling, neighbor, acceptance, search and objective functions were employed, all other parameters were the same between the runs.

\section*{Conclusions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Backmatter begins here                   %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{backmatter}

\section*{Competing interests}
  The authors declare that they have no competing interests.

\section*{Funding}
This study was supported by an award from the National Science Foundation (NSF CBET-0955172) and the National Institutes of Health (NIH HL110328) to J.B, and by
a National Science Foundation Graduate Research Fellowship (NSGRF) to D.B.

\section*{Author's contributions}
    J.V developed the software presented in this study. The manuscript was prepared and edited for publication by D.B, J.B. and J.V.

%\section*{Acknowledgements}
%  Text for this section \ldots
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%
%%  Bmc_mathpys.bst  will be used to                       %%
%%  create a .BBL file for submission.                     %%
%%  After submission of the .TEX file,                     %%
%%  you will be prompted to submit your .BBL file.         %%
%%                                                         %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %%
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %%
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% if your bibliography is in bibtex format, use those commands:
\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliography{References_v1}      % Bibliography file (usually '*.bib' )
% for author-year bibliography (bmc-mathphys or spbasic)
% a) write to bib file (bmc-mathphys only)
% @settings{label, options="nameyear"}
% b) uncomment next line
%\nocite{label}

% or include bibliography directly:
% \begin{thebibliography}
% \bibitem{b1}
% \end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Figures                       %%
%%                               %%
%% NB: this is for captions and  %%
%% Titles. All graphics must be  %%
%% submitted separately and NOT  %%
%% included in the Tex document  %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% Do not use \listoffigures as most will included as separate files

\clearpage
\section*{Figures}

\begin{figure}[h]
  \caption{Schematic of multiobjective parameter mapping. }
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Tables                        %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Use of \listoftables is discouraged.
%%
\clearpage
\section*{Tables}

% \begin{table}[h!]
% \caption{Sample table title. This is where the description of the table should go.}
%       \begin{tabular}{cccc}
%         \hline
%            & B1  &B2   & B3\\ \hline
%         A1 & 0.1 & 0.2 & 0.3\\
%         A2 & ... & ..  & .\\
%         A3 & ..  & .   & .\\ \hline
%       \end{tabular}
% \end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Additional Files              %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \section*{Additional Files}
%   \subsection*{Additional file 1 --- Sample additional file title}
%     Additional file descriptions text (including details of how to
%     view the file, if it is in a non-standard format or the file extension).  This might
%     refer to a multi-page table or a figure.
%
%   \subsection*{Additional file 2 --- Sample additional file title}
%     Additional file descriptions text.

\end{backmatter}
\end{document}

### 2016-05-12 data/binh_korn_lib.jl

@@ -1,0 +1,77 @@
+# some global parameters -
+BIG = 1e10
+
+# Evaluates the objective function values -
+function objective_function(parameter_array)
+
+  # Alias the species -
+  x = parameter_array[1]
+  y = parameter_array[2]
+
+  # Calculate the objective function array -
+  obj_array = BIG*ones(2,1)
+
+  # f1 and f2 -
+  obj_array[1] = 4.0*(x^2)+4.0*(y^2)
+  obj_array[2] = (x - 5)^2 + (y - 5)^2
+
+  # Constraints are implemented as a penaltly on obj value
+  lambda_value = 100.0
+
+  # How much do we violate the constraints?
+  violation_constraint_1 = 25 - (x-5.0)^2 - y^2
+  violation_constraint_2 = (x-8.0)^2 + (y-3.0)^2 - 7.7
+  penaltly_array = zeros(2)
+  penaltly_array[1] = lambda_value*(min(0,violation_constraint_1))^2
+  penaltly_array[2] = lambda_value*(min(0,violation_constraint_2))^2
+
+  # return the obj_array -
+  return obj_array+penaltly_array
+end
+
+# Generates new parameter array, given current array -
+function neighbor_function(parameter_array)
+
+  SIGMA = 0.05
+  number_of_parameters = length(parameter_array)
+
+  # calculate new parameters -
+  new_parameter_array = parameter_array.*(1+SIGMA*randn(number_of_parameters))
+
+  # Check the bound constraints -
+  LOWER_BOUND = [0,0]
+  UPPER_BOUND = [5,3]
+
+  # return the corrected parameter arrays -
+  return parameter_bounds_function(new_parameter_array,LOWER_BOUND,UPPER_BOUND)
+end
+
+function acceptance_probability_function(rank_array,temperature)
+
+  return (exp(-rank_array[end]/temperature))
+
+end
+
+# Helper functions -
+function parameter_bounds_function(parameter_array,lower_bound_array,upper_bound_array)
+
+  # reflection_factor -
+  epsilon = 0.01
+
+  # iterate through and fix the parameters -
+  new_parameter_array = copy(parameter_array)
+  for (index,value) in enumerate(parameter_array)
+
+    lower_bound = lower_bound_array[index]
+    upper_bound = upper_bound_array[index]
+
+    if (value<lower_bound)
+      new_parameter_array[index] = lower_bound + epsilon*upper_bound
+    elseif (value>upper_bound)
[…]

### 2016-05-12 data/fonseca_fleming_lib.jl

@@ -1,0 +1,79 @@
+using Debug
+
+# some global parameters -
+BIG = 1e10
+
+# Evaluates the objective function values -
+function objective_function(parameter_array)
+
+  # Alias the species -
+  number_of_parameters = length(parameter_array)
+
+  # Calculate the objective function array -
+  obj_array = BIG*ones(2,1)
+
+  # calculate the objectibe array -
+  # calculate the sums
+  sum_1 = 0.0
+  sum_2 = 0.0
+  for index in collect(1:number_of_parameters)
+
+    sum_1 = sum_1 + (parameter_array[index] - 1/sqrt(number_of_parameters))^2
+    sum_2 = sum_2 + (parameter_array[index] + 1/sqrt(number_of_parameters))^2
+  end
+
+  # objectives -
+  obj_array[1] = 1 - exp(-1*sum_1)
+  obj_array[2] = 1 - exp(-1*sum_2)
+
+  # return -
+  return obj_array
+end
+
+# Generates new parameter array, given current array -
+function neighbor_function(parameter_array)
+
+  SIGMA = 0.05
+  number_of_parameters = length(parameter_array)
+
+  # correct for zero -
+  idx_zero = find(parameter_array.==0.0)
+  parameter_array[idx_zero] = randn(length(idx_zero))
+  
+  # calculate new parameters -
+  new_parameter_array = parameter_array.*(1+SIGMA*randn(number_of_parameters))
+
+  # Check the bound constraints -
+  LOWER_BOUND = -4.0
+  UPPER_BOUND = 4.0
+
+  # return the corrected parameter arrays -
+  return parameter_bounds_function(new_parameter_array,LOWER_BOUND*ones(number_of_parameters),UPPER_BOUND*ones(number_of_parameters))
+end
+
+function acceptance_probability_function(rank_array,temperature)
+  return (exp(-rank_array[end]/temperature))
+end
+
+# Helper functions -
+function parameter_bounds_function(parameter_array,lower_bound_array,upper_bound_array)
+
+  # reflection_factor -
+  epsilon = 0.01
+
+  # iterate through and fix the parameters -
+  new_parameter_array = copy(parameter_array)
+  for (index,value) in enumerate(parameter_array)
+
+    lower_bound = lower_bound_array[index]
+    upper_bound = upper_bound_array[index]
+
+    if (value<lower_bound)
+      new_parameter_array[index] = lower_bound
+    elseif (value>upper_bound)
[…]

### 2016-05-12 data/test_binh_korn.jl

@@ -1,0 +1,19 @@
+# Script to generate test data from the bihn and korn function
+include("binh_korn_lib.jl")
+
+number_of_subdivisions = 10
+initial_state_array = zeros(2)
+x_initial_array = linspace(0.01,5,number_of_subdivisions)
+y_initial_array = linspace(0.01,3,number_of_subdivisions)
+ec_array = zeros(2)
+pc_array = zeros(2)
+for index in collect(1:number_of_subdivisions)
+
+  initial_state_array[1] = float(x_initial_array[index])
+  initial_state_array[2] = float(y_initial_array[index])
+  (EC,PC,RA) = estimate_ensemble(objective_function,neighbor_function,acceptance_probability_function,initial_state_array;rank_cutoff=4,maximum_number_of_iterations=10)
+
+  ec_array = [ec_array EC]
+  pc_array = [pc_array PC]
+
+end

### 2016-05-12 data/test_fonseca_fleming.jl

@@ -1,0 +1,26 @@
+# Script to generate test data from the bihn and korn function
+include("fonseca_fleming_lib.jl")
+
+number_of_subdivisions = 10
+number_of_states = 4
+initial_state_array = zeros(number_of_states)
+x_initial_array = linspace(-4,4,number_of_subdivisions)
+iz = find(x_initial_array==0.0)
+x_initial_array[iz] = randn(length(iz))
+
+
+ec_array = zeros(2)
+pc_array = zeros(number_of_states)
+for index in collect(1:number_of_subdivisions)
+
+  initial_state_array[1] = float(x_initial_array[index])
+  initial_state_array[2] = float(x_initial_array[index])
+  (EC,PC,RA) = estimate_ensemble(objective_function,neighbor_function,acceptance_probability_function,initial_state_array;rank_cutoff=4,maximum_number_of_iterations=12)
+
+  ec_array = [ec_array EC]
+  pc_array = [pc_array PC]
+
+end
+
+# rerank -
+final_rank_array = rank_function(ec_array)

### 2016-05-12 octave_poets/POETs.m

@@ -1,0 +1,104 @@
+function [pareto_rank_array,error_cache,parameter_cache] = POETs(objective_function,neighbor_function,acceptance_probability_function,initial_state,maximum_number_of_iterations,rank_cutoff,show_trace)
+
+  % internal parameters -
+  temperature = 1.0;
+  temperature_min = temperature/10000;
+  alpha = 0.90;
+
+  % Grab the initial parameters -
+  parameter_array_best = initial_state;
+
+  % initialize error cache -
+  error_cache = objective_function(parameter_array_best);
+
+  % initialize parameter_cache -
+  parameter_cache = parameter_array_best;
+
+  % initialize the Pareto rank array from the error_cache -
+  pareto_rank_array = rank_function(error_cache);
+
+  % how many objectives do we have?
+  number_of_objectives = length(error_cache);
+
+  %keyboard
+
+  % main loop -
+  while (temperature>temperature_min)
+
+    should_loop_continue = true;
+    iteration_index = 1;
+    while (should_loop_continue)
+
+      % generate a new solution -
+      test_parameter_array = neighbor_function(parameter_array_best);
+
+      %keyboard
+
+      % evaluate the new solution -
+      test_error = objective_function(test_parameter_array);
+
+      % Add the test error to the error cache -
+      error_cache = [error_cache test_error];
+
+      % Add parameters to parameter cache -
+      parameter_cache = [parameter_cache test_parameter_array];
+
+      % compute the Pareto rank for the error_cache -
+      pareto_rank_array = rank_function(error_cache);
+
+      % do we accept the new solution?
+      acceptance_probability = acceptance_probability_function(pareto_rank_array,temperature);
+      if (acceptance_probability>rand())
+
+        % Select the rank -
+        archive_select_index = find(pareto_rank_array<rank_cutoff);
+
+        % update the caches -
+        error_cache = error_cache(:,archive_select_index);
+        parameter_cache = parameter_cache(:,archive_select_index);
+
+        % update the parameters -
+        parameter_array_best = test_parameter_array;
+
+        if (show_trace == true)
+          msg = ['iteration_index = ',num2str(iteration_index),' temperature = ',num2str(temperature)];
+          disp(msg)
+        end
+      end
+
+      %msg = ['iteration_index = ',num2str(iteration_index),' temperature = ',num2str(temperature)]
+      %disp(msg)
+
[…]

### 2016-05-12 octave_poets/acceptance.m

@@ -1,0 +1,3 @@
+function [probability] = acceptance(rank_array,temperature)
+  probability = exp(-rank_array(end)/temperature);
+return;

### 2016-05-12 octave_poets/binh_korn_objective.m

@@ -1,0 +1,28 @@
+function [total_objective_array] = binh_korn_objective(parameter_array)
+
+  BIG = 1e10;
+
+  % Alias the species -
+  x = parameter_array(1,1);
+  y = parameter_array(2,1);
+
+  % Calculate the objective function array -
+  obj_array = BIG*ones(2,1);
+
+  % f1 and f2 -
+  obj_array(1,1) = 4.0*(x^2)+4.0*(y^2);
+  obj_array(2,1) = (x - 5)^2 + (y - 5)^2;
+
+  % Constraints are implemented as a penaltly on obj value
+  lambda_value = 100.0;
+
+  % How much do we violate the constraints?
+  violation_constraint_1 = 25 - (x-5.0)^2 - y^2;
+  violation_constraint_2 = (x-8.0)^2 + (y-3.0)^2 - 7.7;
+  penaltly_array = zeros(2,1);
+  penaltly_array(1,1) = lambda_value*(min(0,violation_constraint_1))^2;
+  penaltly_array(2,1) = lambda_value*(min(0,violation_constraint_2))^2;
+
+  % return the obj_array -
+  total_objective_array = obj_array+penaltly_array;
+return;

### 2016-05-12 octave_poets/binh_korn_test.m

@@ -1,0 +1,48 @@
+function [rank_archive,error_cache,parameter_cache,elapsed_time] = binh_korn_test()
+
+  # start -
+  id = tic();
+
+  % Ok, we need to split the domain into a number of subdivisions -
+  number_of_subdivisions = 10;
+  x_initial_array = linspace(0.01,5,number_of_subdivisions);
+  y_initial_array = linspace(0.01,3,number_of_subdivisions);
+
+  % Setup the function pointers -
+  pObjectiveFunction = @binh_korn_objective;
+  pNeighborFunction = @neighbor;
+  pCoolingFunction = @cooling;
+  pAcceptanceFunction = @acceptance;
+
+  % setup interations et al
+  maximum_number_of_iterations = 10;
+  rank_cutoff = 5;
+  show_trace = false;
+
+  rank_archive = [];
+  error_cache = [];
+  parameter_cache = [];
+
+  % main run loop -
+  for domain_index = 1:number_of_subdivisions
+
+    domain_index
+
+    % setup inital guess -
+    initial_state = zeros(2,1);
+    initial_state(1,1) = x_initial_array(domain_index);
+    initial_state(2,1) = y_initial_array(domain_index);
+
+    % call POETs
+    [RA,EC,PC] = POETs(pObjectiveFunction,pNeighborFunction,pAcceptanceFunction,initial_state,maximum_number_of_iterations,rank_cutoff,show_trace);
+
+    % cache -
+    rank_archive = [rank_archive RA];
+    error_cache = [error_cache EC];
+    parameter_cache = [parameter_cache PC];
+
+  end
+
+  elapsed_time = toc(id);
+
+return

### 2016-05-12 octave_poets/neighbor.m

@@ -1,0 +1,39 @@
+function [new_parameter_array] = neighbor(parameter_array)
+
+  SIGMA = 0.05;
+  number_of_parameters = length(parameter_array);
+
+  % calculate new parameters -
+  new_parameter_array = parameter_array.*(1+SIGMA*randn(number_of_parameters));
+
+  % Check the bound constraints -
+  LOWER_BOUND = [0,0];
+  UPPER_BOUND = [5,3];
+
+  % return the corrected parameter arrays -
+  new_parameter_array =  parameter_bounds_function(new_parameter_array,LOWER_BOUND,UPPER_BOUND);
+return
+
+function [new_parameter_array] = parameter_bounds_function(parameter_array,lower_bound_array,upper_bound_array)
+
+  % reflection_factor -
+  epsilon = 0.01;
+
+  % iterate through and fix the parameters -
+  new_parameter_array = parameter_array;
+  number_of_parameters = length(new_parameter_array);
+  for index = 1:number_of_parameters
+
+    value = parameter_array(index,1);
+
+    lower_bound = lower_bound_array(index);
+    upper_bound = upper_bound_array(index);
+
+    if (value<lower_bound)
+      new_parameter_array(index,1) = lower_bound + epsilon*upper_bound;
+    elseif (value>upper_bound)
+      new_parameter_array(index,1) = upper_bound - epsilon*lower_bound;
+    end
+  end
+
+return;

### 2016-05-13 Paper_v1.tex
Example timing functions

@@
 However, the unifying component of all of these previous strategies was the minimization of a single objective function.
-Identification of large-scale models with hundreds, thousands or tens of thousands states and parameters may not be tractable with single objective formulations.
+
+Identification of large models with hundreds or even thousands of states and parameters may not be tractable as a single objective formulation.
 Models at these scales require significant training data perhaps taken from diverse sources, for example different laboratories or cell-lines.
-These diverse training data would likely be heterogenous and could contain intrinsic conflicts.
-Another strategy to formulate large-scale models is to integrate many small pathway modules into a single large model.
+These diverse training data are likely heterogenous, and contain intrinsic conflicts.
+On the other hand, large models can also be constructed from many smaller modules.
 For example, Diamond and coworkers developed a model of phosphoinositide and calcium signaling in human platelets composed of many
@@
 
-In this study, we present JuPOETs, an updated version of the Pareto Optimal Ensemble Technique (POETs) ported to Julia.
+In this study, we present JuPOETs, an open-source implementation of the Pareto Optimal Ensemble Technique (POETs) in the Julia programming language.
 JuPOETs offers many advantages and improvements compared to the previous implementation of POETs.
@@
 However, as $T$ was reduced, the probability of this occurring decreased.
-Parameter sets could be accepted by the SA and $\emph{not}$ archived in $\mathbf{K}_{i}$. Only
-parameter sets with $rank\leq{2}$ were included in $\mathbf{K}_{i}$ to ensure that we characterized the neighborhood near the trade-off surface.
-The general scheme of JuPOETs followed conventional simulated annealing except that the current archive of Pareto-optimal ensemble was updated whenever a new proposal solution is accepted JuPOETs expands upon the original POETs implementation by using the function type in Julia to allow user defined functions for all major components of the approach (Algorithm \ref{algo:JuPOETs-pcode}).
+Parameter sets could be accepted by the SA and $\emph{not}$ archived in $\mathbf{K}_{i}$.
+Only parameter sets with rank less than or equal to a user defined threshold (rank$\leq$4 by default) were included in $\mathbf{K}_{i}$ to ensure that we characterized the neighborhood near the trade-off surface (Algorithm \ref{algo:JuPOETs-pcode}).
 
-\begin{algorithm}
+\begin{algorithm}[h!]
 
@@
  }
-\vspace{0.1in}
-\caption{Pseudo-code for the run-loop of the JuPOETs algorithm. }\label{algo:JuPOETs-pcode}
+\vspace{0.05in}
+\caption{Pseudo-code for the run-loop of the JuPOETs algorithm. The user specifies the neighbor, acceptance, cooling and objective function pointers along with a initial parameter guess.
+The rank archive $\mathcal{R}$, solution archive $\mathcal{S}$ and objective archive $\mathcal{O}$ are initialized from the initial guess.
+The initial guess is then perturbed in the neighbor function, which generates a new solution whose performance is evaluated using the user supplied objective function.
+The new solution and objective values are then added to the respective archives and ranked using the builtin rank function.
+If the new solution accepted (based upon a probability calculated with the user supplied acceptance function) it is added to the solution and objective archive.
+This solution is then perturbed during the next iteration of the algorithm.
+However, if the solution is not accepted, it is removed from the archive and discarded.
+The computational temperature is adjusted using the user supplied cooling function after each $\mathcal{I}$ iterations.
+}\label{algo:JuPOETs-pcode}
 \end{algorithm}
+
+\clearpage
 
@@
 We compared the performance of JuPOETs versus POETs implemented in Octave for a suite of multiobjective test functions (Table ZZ).
-In all cases, the same cooling, neighbor, acceptance, search and objective functions were employed, all other parameters were the same between the runs.
+In all cases, the same cooling, neighbor, acceptance, search and objective logic was employed, with all other parameters held constant.
+For each test function, the search domain was partitioned into a fixed number of segments, where an initial parameter guess was drawn from each the partitions.
 

### 2016-05-13 data/run_binh_korn_test.jl

@@ -1,0 +1,13 @@
+include("test_binh_korn.jl")
+
+
+number_of_trials = 10
+etime_array = zeros(number_of_trials)
+for index in collect(1:number_of_trials)
+
+  tic()
+  test_binh_korn()
+  elapsed_time = toc()
+  etime_array[index] = elapsed_time
+
+end

### 2016-05-13 data/schaffer_lib.jl

@@ -5,27 +5,23 @@ BIG = 1e10
 function objective_function(parameter_array)
 
   # Alias the species -
-  number_of_parameters = length(parameter_array)
+  x = parameter_array[1]
 
   # Calculate the objective function array -
   obj_array = BIG*ones(2,1)
 
-  # calculate the objectibe array -
-  # calculate the sums
-  sum_1 = 0.0
-  sum_2 = 0.0
-  for index in collect(1:number_of_parameters)
+  # f1 and f2 -
+  obj_array[1] = (x^2)
+  obj_array[2] = (x - 2)^2
 
-    sum_1 = sum_1 + (parameter_array[index] - 1/sqrt(number_of_parameters))^2
-    sum_2 = sum_2 + (parameter_array[index] + 1/sqrt(number_of_parameters))^2
-  end
+  # Constraints are implemented as a penaltly on obj value
+  lambda_value = 100.0
 
-  # objectives -
-  obj_array[1] = 1 - exp(-1*sum_1)
-  obj_array[2] = 1 - exp(-1*sum_2)
+  # How much do we violate the constraints?
+  penaltly_array = zeros(2)
 
-  # return -
-  return obj_array
+  # return the obj_array -
+  return obj_array+penaltly_array
 end
 
 # Generates new parameter array, given current array -
@@ -34,23 +30,21 @@ function neighbor_function(parameter_array)
   SIGMA = 0.05
   number_of_parameters = length(parameter_array)
 
-  # correct for zero -
-  idx_zero = find(parameter_array.==0.0)
-  parameter_array[idx_zero] = randn(length(idx_zero))
-
   # calculate new parameters -
   new_parameter_array = parameter_array.*(1+SIGMA*randn(number_of_parameters))
 
   # Check the bound constraints -
-  LOWER_BOUND = -4.0
-  UPPER_BOUND = 4.0
+  LOWER_BOUND = -10
+  UPPER_BOUND = 10
 
   # return the corrected parameter arrays -
-  return parameter_bounds_function(new_parameter_array,LOWER_BOUND*ones(number_of_parameters),UPPER_BOUND*ones(number_of_parameters))
+  return parameter_bounds_function(new_parameter_array,LOWER_BOUND,UPPER_BOUND)
 end
 
 function acceptance_probability_function(rank_array,temperature)
+
   return (exp(-rank_array[end]/temperature))
+
 end
 
 # Helper functions -
[…]

### 2016-05-13 data/test_fonseca_fleming.jl

@@ -1,26 +1,31 @@
 # Script to generate test data from the bihn and korn function
 include("fonseca_fleming_lib.jl")
 
-number_of_subdivisions = 10
-number_of_states = 4
-initial_state_array = zeros(number_of_states)
-x_initial_array = linspace(-4,4,number_of_subdivisions)
-iz = find(x_initial_array==0.0)
-x_initial_array[iz] = randn(length(iz))
+function test_fonseca_fleming()
 
+  number_of_subdivisions = 10
+  number_of_states = 3
+  initial_state_array = zeros(number_of_states)
+  x_initial_array = linspace(-4,4,number_of_subdivisions)
+  iz = find(x_initial_array==0.0)
+  x_initial_array[iz] = randn(length(iz))
 
-ec_array = zeros(2)
-pc_array = zeros(number_of_states)
-for index in collect(1:number_of_subdivisions)
+  ec_array = zeros(2)
+  pc_array = zeros(number_of_states)
+  ra_array = zeros(1)
+  for index in collect(1:number_of_subdivisions)
 
-  initial_state_array[1] = float(x_initial_array[index])
-  initial_state_array[2] = float(x_initial_array[index])
-  (EC,PC,RA) = estimate_ensemble(objective_function,neighbor_function,acceptance_probability_function,initial_state_array;rank_cutoff=4,maximum_number_of_iterations=12)
+    initial_state_array[1] = float(x_initial_array[index])
+    initial_state_array[2] = float(x_initial_array[index])
+    initial_state_array[3] = float(x_initial_array[index])
+    (EC,PC,RA) = estimate_ensemble(objective_function,neighbor_function,acceptance_probability_function,initial_state_array;rank_cutoff=4,maximum_number_of_iterations=10,show_trace=false)
 
-  ec_array = [ec_array EC]
-  pc_array = [pc_array PC]
+    ec_array = [ec_array EC]
+    pc_array = [pc_array PC]
 
+  end
+  return (ec_array,pc_array,ra_array)
 end
 
 # rerank -
-final_rank_array = rank_function(ec_array)
+#final_rank_array = rank_function(ec_array)

### 2016-05-13 data/test_schaffer.jl

@@ -1,19 +1,17 @@
 # Script to generate test data from the bihn and korn function
-include("binh_korn_lib.jl")
+include("schaffer_lib.jl")
 
-function test_binh_korn()
+function test_schaffer()
 
   number_of_subdivisions = 10
-  initial_state_array = zeros(2)
-  x_initial_array = linspace(0.01,5,number_of_subdivisions)
-  y_initial_array = linspace(0.01,3,number_of_subdivisions)
+  initial_state_array = zeros(1)
+  x_initial_array = linspace(-10,10,number_of_subdivisions)
   ec_array = zeros(2)
-  pc_array = zeros(2)
+  pc_array = zeros(1)
   ra_array = zeros(1)
   for index in collect(1:number_of_subdivisions)
 
     initial_state_array[1] = float(x_initial_array[index])
-    initial_state_array[2] = float(y_initial_array[index])
     (EC,PC,RA) = estimate_ensemble(objective_function,neighbor_function,acceptance_probability_function,initial_state_array;rank_cutoff=4,maximum_number_of_iterations=10,show_trace=false)
 
     #@show size(RA)

### 2016-05-13 octave_poets/fonseca_fleming_neighbor.m

@@ -1,0 +1,39 @@
+function [new_parameter_array] = fonseca_fleming_neighbor(parameter_array)
+
+  SIGMA = 0.05;
+  number_of_parameters = length(parameter_array);
+
+  % calculate new parameters -
+  new_parameter_array = parameter_array.*(1+SIGMA*randn(number_of_parameters));
+
+  % Check the bound constraints -
+  LOWER_BOUND = [-4.0,-4.0,-4.0];
+  UPPER_BOUND = [4.0,4.0,4.0];
+
+  % return the corrected parameter arrays -
+  new_parameter_array =  parameter_bounds_function(new_parameter_array,LOWER_BOUND,UPPER_BOUND);
+return
+
+function [new_parameter_array] = parameter_bounds_function(parameter_array,lower_bound_array,upper_bound_array)
+
+  % reflection_factor -
+  epsilon = 0.01;
+
+  % iterate through and fix the parameters -
+  new_parameter_array = parameter_array;
+  number_of_parameters = length(new_parameter_array);
+  for index = 1:number_of_parameters
+
+    value = parameter_array(index,1);
+
+    lower_bound = lower_bound_array(index);
+    upper_bound = upper_bound_array(index);
+
+    if (value<lower_bound)
+      new_parameter_array(index,1) = lower_bound;
+    elseif (value>upper_bound)
+      new_parameter_array(index,1) = upper_bound;
+    end
+  end
+
+return;

### 2016-05-13 octave_poets/fonseca_fleming_objective.m

@@ -1,0 +1,26 @@
+function [obj_array] = fonseca_fleming_objective(parameter_array)
+
+
+  BIG = 1e10;
+
+  # Alias the species -
+  number_of_parameters = length(parameter_array);
+
+  # Calculate the objective function array -
+  obj_array = BIG*ones(2,1);
+
+  # calculate the objectibe array -
+  # calculate the sums
+  sum_1 = 0.0;
+  sum_2 = 0.0;
+  for index = 1:number_of_parameters
+
+    sum_1 = sum_1 + (parameter_array(index) - 1/sqrt(number_of_parameters))^2;
+    sum_2 = sum_2 + (parameter_array(index) + 1/sqrt(number_of_parameters))^2;
+  end
+
+  # objectives -
+  obj_array(1,1) = 1 - exp(-1*sum_1);
+  obj_array(2,1) = 1 - exp(-1*sum_2);
+
+return

### 2016-05-13 octave_poets/fonseca_fleming_test.m

@@ -1,16 +1,17 @@
-function [rank_archive,error_cache,parameter_cache,elapsed_time] = binh_korn_test()
+function [rank_archive,error_cache,parameter_cache,elapsed_time] = fonseca_fleming_test()
 
   % start -
   id = tic();
 
   % Ok, we need to split the domain into a number of subdivisions -
   number_of_subdivisions = 10;
-  x_initial_array = linspace(0.01,5,number_of_subdivisions);
-  y_initial_array = linspace(0.01,3,number_of_subdivisions);
+  x_initial_array = linspace(-4.0,4.0,number_of_subdivisions);
+  y_initial_array = linspace(-4.0,4.0,number_of_subdivisions);
+  z_initial_array = linspace(-4.0,4.0,number_of_subdivisions);
 
   % Setup the function pointers -
-  pObjectiveFunction = @binh_korn_objective;
-  pNeighborFunction = @neighbor;
+  pObjectiveFunction = @fonseca_fleming_objective;
+  pNeighborFunction = @fonseca_fleming_neighbor;
   pCoolingFunction = @cooling;
   pAcceptanceFunction = @acceptance;
 
@@ -29,9 +30,13 @@ function [rank_archive,error_cache,parameter_cache,elapsed_time] = binh_korn_tes
     %domain_index
 
     % setup inital guess -
-    initial_state = zeros(2,1);
+    initial_state = zeros(3,1);
     initial_state(1,1) = x_initial_array(domain_index);
     initial_state(2,1) = y_initial_array(domain_index);
+    initial_state(3,1) = y_initial_array(domain_index);
+
+    iz = find(initial_state==0.0);
+    initial_state(iz,1) = randn(length(iz),1);
 
     % call POETs
     [RA,EC,PC] = POETs(pObjectiveFunction,pNeighborFunction,pAcceptanceFunction,initial_state,maximum_number_of_iterations,rank_cutoff,show_trace);

### 2016-05-13 octave_poets/run_binh_korn_test.m

@@ -1,0 +1,31 @@
+etime_array = []
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = binh_korn_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = binh_korn_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = binh_korn_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = binh_korn_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = binh_korn_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = binh_korn_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = binh_korn_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = binh_korn_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = binh_korn_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = binh_korn_test();
+etime_array = [etime_array elapsed_time];

### 2016-05-13 octave_poets/run_fonseca_fleming_test.m

@@ -1,0 +1,31 @@
+etime_array = [];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = fonseca_fleming_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = fonseca_fleming_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = fonseca_fleming_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = fonseca_fleming_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = fonseca_fleming_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = fonseca_fleming_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = fonseca_fleming_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = fonseca_fleming_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = fonseca_fleming_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = fonseca_fleming_test();
+etime_array = [etime_array elapsed_time];

### 2016-05-13 octave_poets/run_schaffer_test.m

@@ -1,0 +1,31 @@
+etime_array = [];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = schaffer_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = schaffer_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = schaffer_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = schaffer_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = schaffer_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = schaffer_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = schaffer_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = schaffer_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = schaffer_test();
+etime_array = [etime_array elapsed_time];
+
+[rank_archive,error_cache,parameter_cache,elapsed_time] = schaffer_test();
+etime_array = [etime_array elapsed_time];

### 2016-05-13 octave_poets/schaffer_neighbor.m

@@ -1,0 +1,39 @@
+function [new_parameter_array] = schaffer_neighbor(parameter_array)
+
+  SIGMA = 0.05;
+  number_of_parameters = length(parameter_array);
+
+  % calculate new parameters -
+  new_parameter_array = parameter_array.*(1+SIGMA*randn(number_of_parameters));
+
+  % Check the bound constraints -
+  LOWER_BOUND = [-10];
+  UPPER_BOUND = [10];
+
+  % return the corrected parameter arrays -
+  new_parameter_array =  parameter_bounds_function(new_parameter_array,LOWER_BOUND,UPPER_BOUND);
+return
+
+function [new_parameter_array] = parameter_bounds_function(parameter_array,lower_bound_array,upper_bound_array)
+
+  % reflection_factor -
+  epsilon = 0.01;
+
+  % iterate through and fix the parameters -
+  new_parameter_array = parameter_array;
+  number_of_parameters = length(new_parameter_array);
+  for index = 1:number_of_parameters
+
+    value = parameter_array(index,1);
+
+    lower_bound = lower_bound_array(index);
+    upper_bound = upper_bound_array(index);
+
+    if (value<lower_bound)
+      new_parameter_array(index,1) = lower_bound;
+    elseif (value>upper_bound)
+      new_parameter_array(index,1) = upper_bound;
+    end
+  end
+
+return;

### 2016-05-13 octave_poets/schaffer_objective.m

@@ -1,0 +1,23 @@
+function [total_objective_array] = schaffer_objective(parameter_array)
+
+  BIG = 1e10;
+
+  % Alias the species -
+  x = parameter_array(1,1);
+
+  % Calculate the objective function array -
+  obj_array = BIG*ones(2,1);
+
+  % f1 and f2 -
+  obj_array(1,1) = (x^2);
+  obj_array(2,1) = (x - 2)^2;
+
+  % Constraints are implemented as a penaltly on obj value
+  lambda_value = 100.0;
+
+  % How much do we violate the constraints?
+  penaltly_array = zeros(2,1);
+
+  % return the obj_array -
+  total_objective_array = obj_array+penaltly_array;
+return;

### 2016-05-13 octave_poets/schaffer_test.m

@@ -1,16 +1,15 @@
-function [rank_archive,error_cache,parameter_cache,elapsed_time] = binh_korn_test()
+function [rank_archive,error_cache,parameter_cache,elapsed_time] = schaffer_test()
 
-  % start -
+  # start -
   id = tic();
 
   % Ok, we need to split the domain into a number of subdivisions -
   number_of_subdivisions = 10;
-  x_initial_array = linspace(0.01,5,number_of_subdivisions);
-  y_initial_array = linspace(0.01,3,number_of_subdivisions);
+  x_initial_array = linspace(-10,10,number_of_subdivisions);
 
   % Setup the function pointers -
-  pObjectiveFunction = @binh_korn_objective;
-  pNeighborFunction = @neighbor;
+  pObjectiveFunction = @schaffer_objective;
+  pNeighborFunction = @schaffer_neighbor;
   pCoolingFunction = @cooling;
   pAcceptanceFunction = @acceptance;
 
@@ -29,9 +28,8 @@ function [rank_archive,error_cache,parameter_cache,elapsed_time] = binh_korn_tes
     %domain_index
 
     % setup inital guess -
-    initial_state = zeros(2,1);
+    initial_state = zeros(1,1);
     initial_state(1,1) = x_initial_array(domain_index);
-    initial_state(2,1) = y_initial_array(domain_index);
 
     % call POETs
     [RA,EC,PC] = POETs(pObjectiveFunction,pNeighborFunction,pAcceptanceFunction,initial_state,maximum_number_of_iterations,rank_cutoff,show_trace);

### 2016-05-17 data/binh_korn_lib.jl
New figures in figs dir

@@ -20,7 +20,7 @@ function objective_function(parameter_array)
 
   # How much do we violate the constraints?
   violation_constraint_1 = 25 - (x-5.0)^2 - y^2
-  violation_constraint_2 = (x-8.0)^2 + (y-3.0)^2 - 7.7
+  violation_constraint_2 = (x-8.0)^2 + (y+3.0)^2 - 7.7
   penaltly_array = zeros(2)
   penaltly_array[1] = lambda_value*(min(0,violation_constraint_1))^2
   penaltly_array[2] = lambda_value*(min(0,violation_constraint_2))^2
@@ -49,7 +49,13 @@ end
 function acceptance_probability_function(rank_array,temperature)
 
   return (exp(-rank_array[end]/temperature))
+end
 
+function cooling_function(temperature)
+
+  # define my new temperature -
+  alpha = 0.9
+  return 0.9*temperature
 end
 
 # Helper functions -

### 2016-05-17 data/fonseca_fleming_lib.jl

@@ -53,6 +53,14 @@ function acceptance_probability_function(rank_array,temperature)
   return (exp(-rank_array[end]/temperature))
 end
 
+function cooling_function(temperature)
+
+  # define my new temperature -
+  alpha = 0.9
+  return alpha*temperature
+end
+
+
 # Helper functions -
 function parameter_bounds_function(parameter_array,lower_bound_array,upper_bound_array)
 

### 2016-05-17 data/plot_binh_korn.jl

@@ -1,0 +1,16 @@
+using PyPlot
+using POETs
+
+# load the data -
+ec_array = readdlm("./ec_array_binh_korn.dat")
+
+# re-rank the solutions -
+rank_array = rank_function(ec_array[:,2:end])
+
+
+idx_rank_1 = find(rank_array.>=1.0)
+plot(ec_array[1,idx_rank_1],ec_array[2,idx_rank_1],".",color="0.75")
+
+# Plot rank 0 in black -
+idx_rank_0 = find(rank_array.==0.0)
+plot(ec_array[1,idx_rank_0],ec_array[2,idx_rank_0],"k.")

### 2016-05-17 data/plot_fonseca_fleming.jl

@@ -1,0 +1,16 @@
+using PyPlot
+using POETs
+
+# load the data -
+ec_array = readdlm("./ec_array_ff.dat")
+
+# re-rank the solutions -
+rank_array = rank_function(ec_array[:,2:end])
+
+
+idx_rank_1 = find(rank_array.>=1.0)
+plot(ec_array[1,idx_rank_1],ec_array[2,idx_rank_1],".",color="0.75")
+
+# Plot rank 0 in black -
+idx_rank_0 = find(rank_array.==0.0)
+plot(ec_array[1,idx_rank_0],ec_array[2,idx_rank_0],"k.")

### 2016-05-17 data/plot_schaffer.jl

@@ -1,0 +1,16 @@
+using PyPlot
+using POETs
+
+# load the data -
+ec_array = readdlm("./ec_array_schaffer.dat")
+
+# re-rank the solutions -
+rank_array = rank_function(ec_array[:,2:end])
+
+
+idx_rank_1 = find(rank_array.>=1.0)
+plot(ec_array[1,idx_rank_1],ec_array[2,idx_rank_1],".",color="0.75")
+
+# Plot rank 0 in black -
+idx_rank_0 = find(rank_array.==0.0)
+plot(ec_array[1,idx_rank_0],ec_array[2,idx_rank_0],"k.")

### 2016-05-17 data/schaffer_lib.jl

@@ -47,6 +47,14 @@ function acceptance_probability_function(rank_array,temperature)
 
 end
 
+function cooling_function(temperature)
+
+  # define my new temperature -
+  alpha = 0.9
+  return 0.9*temperature
+end
+
+
 # Helper functions -
 function parameter_bounds_function(parameter_array,lower_bound_array,upper_bound_array)
 

### 2016-05-17 data/test_binh_korn.jl

@@ -9,19 +9,17 @@ function test_binh_korn()
   y_initial_array = linspace(0.01,3,number_of_subdivisions)
   ec_array = zeros(2)
   pc_array = zeros(2)
-  ra_array = zeros(1)
   for index in collect(1:number_of_subdivisions)
 
     initial_state_array[1] = float(x_initial_array[index])
     initial_state_array[2] = float(y_initial_array[index])
-    (EC,PC,RA) = estimate_ensemble(objective_function,neighbor_function,acceptance_probability_function,initial_state_array;rank_cutoff=4,maximum_number_of_iterations=10,show_trace=false)
+    (EC,PC,RA) = estimate_ensemble(objective_function,neighbor_function,acceptance_probability_function,cooling_function,initial_state_array;rank_cutoff=4,maximum_number_of_iterations=20,show_trace=false)
 
     #@show size(RA)
 
     ec_array = [ec_array EC]
     pc_array = [pc_array PC]
-    #ra_array = [ra_array RA]
   end
 
-  return (ec_array,pc_array,ra_array)
+  return (ec_array,pc_array)
 end

### 2016-05-17 data/test_fonseca_fleming.jl

@@ -12,20 +12,16 @@ function test_fonseca_fleming()
 
   ec_array = zeros(2)
   pc_array = zeros(number_of_states)
-  ra_array = zeros(1)
   for index in collect(1:number_of_subdivisions)
 
     initial_state_array[1] = float(x_initial_array[index])
     initial_state_array[2] = float(x_initial_array[index])
     initial_state_array[3] = float(x_initial_array[index])
-    (EC,PC,RA) = estimate_ensemble(objective_function,neighbor_function,acceptance_probability_function,initial_state_array;rank_cutoff=4,maximum_number_of_iterations=10,show_trace=false)
+    (EC,PC,RA) = estimate_ensemble(objective_function,neighbor_function,acceptance_probability_function,cooling_function,initial_state_array;rank_cutoff=4,maximum_number_of_iterations=20,show_trace=false)
 
     ec_array = [ec_array EC]
     pc_array = [pc_array PC]
 
   end
-  return (ec_array,pc_array,ra_array)
+  return (ec_array,pc_array)
 end
-
-# rerank -
-#final_rank_array = rank_function(ec_array)

### 2016-05-17 data/test_schaffer.jl

@@ -8,18 +8,16 @@ function test_schaffer()
   x_initial_array = linspace(-10,10,number_of_subdivisions)
   ec_array = zeros(2)
   pc_array = zeros(1)
-  ra_array = zeros(1)
   for index in collect(1:number_of_subdivisions)
 
     initial_state_array[1] = float(x_initial_array[index])
-    (EC,PC,RA) = estimate_ensemble(objective_function,neighbor_function,acceptance_probability_function,initial_state_array;rank_cutoff=4,maximum_number_of_iterations=10,show_trace=false)
+    (EC,PC,RA) = estimate_ensemble(objective_function,neighbor_function,acceptance_probability_function,cooling_function,initial_state_array;rank_cutoff=4,maximum_number_of_iterations=20,show_trace=false)
 
     #@show size(RA)
 
     ec_array = [ec_array EC]
     pc_array = [pc_array PC]
-    #ra_array = [ra_array RA]
   end
 
-  return (ec_array,pc_array,ra_array)
+  return (ec_array,pc_array)
 end

### 2016-05-18 data/plot_binh_korn.jl
New figures

@@ -3,14 +3,24 @@ using POETs
 
 # load the data -
 ec_array = readdlm("./ec_array_binh_korn.dat")
+pc_array = readdlm("./pc_array_binh_korn.dat")
 
 # re-rank the solutions -
 rank_array = rank_function(ec_array[:,2:end])
 
 
+# idx_rank_1 = find(rank_array.>=1.0)
+# plot(ec_array[1,idx_rank_1],ec_array[2,idx_rank_1],".",color="0.75")
+#
+# # Plot rank 0 in black -
+# idx_rank_0 = find(rank_array.==0.0)
+# plot(ec_array[1,idx_rank_0],ec_array[2,idx_rank_0],"k.")
+
+figure()
+
 idx_rank_1 = find(rank_array.>=1.0)
-plot(ec_array[1,idx_rank_1],ec_array[2,idx_rank_1],".",color="0.75")
+plot(pc_array[1,idx_rank_1],pc_array[2,idx_rank_1],".",color="0.75")
 
 # Plot rank 0 in black -
 idx_rank_0 = find(rank_array.==0.0)
-plot(ec_array[1,idx_rank_0],ec_array[2,idx_rank_0],"k.")
+plot(pc_array[1,idx_rank_0],pc_array[2,idx_rank_0],"k.")

### 2016-05-18 data/plot_fonseca_fleming.jl

@@ -3,14 +3,24 @@ using POETs
 
 # load the data -
 ec_array = readdlm("./ec_array_ff.dat")
+pc_array = readdlm("./pc_array_ff.dat")
 
 # re-rank the solutions -
 rank_array = rank_function(ec_array[:,2:end])
 
 
+#idx_rank_1 = find(rank_array.>=1.0)
+#plot(ec_array[1,idx_rank_1],ec_array[2,idx_rank_1],".",color="0.75")
+
+# Plot rank 0 in black -
+#idx_rank_0 = find(rank_array.==0.0)
+#plot(ec_array[1,idx_rank_0],ec_array[2,idx_rank_0],"k.")
+
+figure()
+
 idx_rank_1 = find(rank_array.>=1.0)
-plot(ec_array[1,idx_rank_1],ec_array[2,idx_rank_1],".",color="0.75")
+plot3D(pc_array[1,idx_rank_1],pc_array[2,idx_rank_1],pc_array[3,idx_rank_1],".",color="0.75")
 
 # Plot rank 0 in black -
 idx_rank_0 = find(rank_array.==0.0)
-plot(ec_array[1,idx_rank_0],ec_array[2,idx_rank_0],"k.")
+plot3D(pc_array[1,idx_rank_0],pc_array[2,idx_rank_0],pc_array[3,idx_rank_0],"k.")

### 2016-05-18 data/plot_schaffer.jl

@@ -3,14 +3,24 @@ using POETs
 
 # load the data -
 ec_array = readdlm("./ec_array_schaffer.dat")
+pc_array = readdlm("./pc_array_schaffer.dat")
 
 # re-rank the solutions -
 rank_array = rank_function(ec_array[:,2:end])
 
 
 idx_rank_1 = find(rank_array.>=1.0)
-plot(ec_array[1,idx_rank_1],ec_array[2,idx_rank_1],".",color="0.75")
+#plot(ec_array[1,idx_rank_1],ec_array[2,idx_rank_1],".",color="0.75")
 
 # Plot rank 0 in black -
 idx_rank_0 = find(rank_array.==0.0)
-plot(ec_array[1,idx_rank_0],ec_array[2,idx_rank_0],"k.")
+#plot(ec_array[1,idx_rank_0],ec_array[2,idx_rank_0],"k.")
+
+figure()
+
+idx_rank_1 = find(rank_array.>=1.0)
+semilogx(transpose(1:length(idx_rank_1)),pc_array[1,idx_rank_1],".",color="0.75")
+
+# Plot rank 0 in black -
+idx_rank_0 = find(rank_array.==0.0)
+semilogx(transpose(1:length(idx_rank_0)),pc_array[1,idx_rank_0],"k.")

### 2016-05-18 Paper_v1.tex
More progess

@@
   O_{1}\left(\mathbf{x}(t,\mathbf{p}),\mathbf{p}\right) \\
-  O_{2}\left(\mathbf{x}(t,\mathbf{p}),\mathbf{p}\right) \\
   \vdots & \\
@@
 \begin{eqnarray}\nonumber
-  \frac{d\mathbf{x}}{dt} & = & \mathbf{f}(t,\mathbf{x}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p}) \\\nonumber
+  \mathbf{f}(t,\mathbf{x}(t,\mathbf{p}),\dot{\mathbf{x}}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p}) &=&\mathbf{0} \\\nonumber
   g_{1}\left(t,\mathbf{x}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p}\right) &\geq& 0 \\\nonumber
-  g_{2}\left(t,\mathbf{x}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p}\right) &\geq& 0 \\\nonumber
   &\vdots& \\\nonumber
@@
 
-POETs integrate Simulated Annealing (SA) with Pareto optimality to estimate parameter sets on or near the optimal tradeoff surface between competing training objectives (Fig. \ref{fig-MOSAalgorithm}).
+POETs integrate Simulated Annealing (SA) with Pareto optimality to estimate parameter sets on or near the optimal tradeoff surface between competing training objectives
+(Fig. \ref{fig:fig-MOSAalgorithm}).
 The central idea underlying POETs is a mapping between the value of the objective functions evaluated at $\mathbf{p}_{i+1}$ (parameter guess at iteration $i+1$) and the Pareto rank.
@@
 \section*{Results and Discussion}
-We compared the performance of JuPOETs versus POETs implemented in Octave for a suite of multiobjective test functions (Table ZZ).
-In all cases, the same cooling, neighbor, acceptance, search and objective logic was employed, with all other parameters held constant.
-For each test function, the search domain was partitioned into a fixed number of segments, where an initial parameter guess was drawn from each the partitions.
+
+JuPOETs identified solutions on or near the optimal tradeoff surface significantly faster than Octave-POETs for the suite of test problems
+(Fig. \ref{fig:fig-timing} and Fig. \ref{fig:fig-surfaces}).
+We compared the performance of JuPOETs versus POETs implemented in Octave for a suite of multiobjective test functions (Table \ref{table:objective-table}).
+The execution time (wall-clock) for JuPOETs and Octave-POETs was measured for 10 independent trials for the suite of test problems.
+The same cooling, neighbor, acceptance, search and objective logic was employed between the implementations, with all other parameters held constant.
+For each test function, the search domain was partitioned into 10 segments, where initial parameter guesses were drawn from each partition.
+The number of search steps for each temperate $\mathcal{I}$ = 10 for all cases, and the cooling parameter was $\alpha$ = 0.9.
+On average, JuPOETs identified optimal or near optimal solutions for the suite of test problems six-fold faster (60s versus 400s)
+than Octave-POETs (Fig. \ref{fig:fig-timing}). In addition, JuPOETs produced the characteristic tradeoff curves for each test problem, given both parameter bound and
+problem constraints (Fig.
+\ref{fig:fig-surfaces}).
+Taken together, these results suggested that JuPOETs was able to estimate an ensemble of solutions to constrained multiobjective optimization problems significantly faster than the current Octave implementation.
+However, we have not demonstrated JuPOETs on a biochemically relevant test problem.
+
+[JuPOETs and a proof-of-principle HCM-EM model goes here]
 
@@
 \clearpage
+
 \section*{Figures}
 
-\begin{figure}[h]
-  \caption{Schematic of multiobjective parameter mapping. }
+\begin{figure}
+  \caption{Schematic of multiobjective parameter mapping.
+The performance of any given parameter set is mapped into an objective space using a ranking function which quantifies the quality of the parameters.
+The distance away from the optimal tradeoff surface is quantified using the Pareto ranking scheme of Fonseca and Fleming in JuPOETs.}\label{fig:fig-MOSAalgorithm}
+\end{figure}
+
+\begin{figure}
+  \caption{The performance of JuPOETs on the multi-objective test suite.
+  The execution time (wall-clock) for JuPOETs and POETs implemented in Octave was measured for 10 independent trials for the suite of test problems.
+  The number of steps per temperature $\mathcal{I}$ = 10, and the cooling parameter $\alpha$ = 0.9 for all cases.
+  The problem domain was partitioned into 10 equal segments, an initial guess was drawn from each segment.
+  For each of the test functions, JuPOETs estimated solutions on (rank zero solutions, black) or near (gray) the optimal tradeoff surface, subject to bounds and problem constraints.
+}\label{fig:fig-timing}
+\end{figure}
+
+\begin{figure}
+  \caption{Representative JuPOETs solutions for problems in the multi-objective test suite.
+  The number of steps per temperature $\mathcal{I}$ = 10, and the cooling parameter $\alpha$ = 0.9 for all cases.
+  The problem domain was partitioned into 10 equal segments, an initial guess was drawn from each segment.
+  For each of the test functions, JuPOETs estimated solutions on (rank zero solutions, black) or near (gray) the optimal tradeoff surface, subject to bounds and problem constraints.
+}\label{fig:fig-surfaces}
 \end{figure}
@@
 \section*{Tables}
+
+\begin{table}
+  \caption{Multi-objective optimization test problems. We tested the JuPOETs implementation on three two-dimensional test problems, with one-, two- and three-dimensional parameter vectors.
+  Each problem had parameter bounds constraints, however, on the Binh and Korn function had additional non-linear problem constraints. For the Fonesca and Fleming problem, N = 3.}
+  \label{table:objective-table}
+\end{table}
 

### 2016-05-19 biochemical/Balances.jl
Biochemical model

@@ -1,0 +1,51 @@
+include("Kinetics.jl")
+include("Control.jl")
+
+function Balances(t,x,DF)
+
+  idx = find(x.<0)
+  x[idx] = 0.0
+
+  # Alias the species vector -
+  e1 = x[1];
+  e2 = x[2];
+  e3 = x[3];
+  e4 = x[4];
+  e5 = x[5];
+  e6 = x[6];
+  Ax = x[7];
+  Bx = x[8];
+  Cx = x[9];
+  Biox = x[10];
+
+  #Define rate vector
+  kinetics_dict = Kinetics(t,x,DF);
+  rM = kinetics_dict["rM_vector"] #Metabolite Rate Vector
+  rE = kinetics_dict["rE_vector"] #Enzyme Rate Vector
+  rG = kinetics_dict["rG_vector"] #Growth Rate
+
+  #Get Parameters from DF
+  alpha = DF["EnzymeSynthesis"]
+  beta = DF["Degradation"]
+  Z = DF["ModeMatrix"]
+  S = DF["MetaboliteMatrix"]
+  (num_reactions,num_modes) = size(Z)
+
+  #Define Control Vector
+  (u,v) = Control(t,x,rM,rE,DF)
+
+  #Redine rates (rate*control)
+  Enzyme_rate = rE.*u;
+  Growth_rate = rG.*v;
+  Reaction_rate = rM.*v;
+  mu = sum(Growth_rate)
+
+  # Enzyme Balances
+  dxdt_vector = zeros(10)
+  dxdt_vector[1:num_modes] = alpha + Enzyme_rate - (beta+mu)*x[1:num_modes]
+
+  # Metabolite Balances
+  dxdt_vector[7:10] = S[1:4,:]*Z*Reaction_rate*Biox
+
+  return dxdt_vector
+end

### 2016-05-19 biochemical/Data_Cybernetic.jl

@@ -1,0 +1,64 @@
+function Data_Cybernetic(tStart,tStop,tStep)
+
+#Load Modes
+mode_matrix = float(readdlm("./data/EM_modes.csv",','));
+
+#Load Stoichiometric Matrix
+stm_matrix = float(readdlm("./data/Network.csv",','));
+
+#============================#
+# EnzymeRateParameters
+  ke = .5
+  alpha = 0.004
+  beta = 0.05
+
+#============================#
+#kcat
+kmax = [
+  0.565;
+  0.337;
+  0.282;
+  0.353;
+  0.276;
+  0.0955;
+  ];
+
+#============================#
+#Ksat
+K = [
+  1.552;
+  1.612;
+  1.431;
+  1.587;
+  1.563;
+  1.42;
+  ];
+
+#=============================#
+#Initial Conditions
+  x0 = [
+    0.5; # em1
+    0.5; # em2
+    0.5; # em3
+    0.5; # em4
+    0.5; # em5
+    0.5; # em6
+    2; # Ax
+    0; # Bx
+    0; # Cx
+    0.01; # Biox
+  ];
+
+#=======================================#
+  #Parametrs from Data_dict
+  Data_dict = Dict()
+  Data_dict["ReactionRateVector"] = kmax
+  Data_dict["SaturationConstantVector"] = K
+  Data_dict["EnzymeRate"] = ke
+  Data_dict["EnzymeSynthesis"] = alpha
+  Data_dict["Degradation"] = beta
+  Data_dict["ModeMatrix"] = mode_matrix
+  Data_dict["MetaboliteMatrix"] = stm_matrix
+  Data_dict["InitialConditions"] = x0
+  return Data_dict
+end

### 2016-05-19 biochemical/SolveBalances.jl

@@ -1,0 +1,49 @@
+include("Data_Cybernetic.jl")
+include("Balances.jl")
+using ODE
+
+function SolveBalances(tStart,tStop,tStep,Param)
+
+  #Define time vector
+  t = collect(tStart:tStep:tStop)
+
+  # Load Initial Conditions
+  Data_dict = Data_Cybernetic(tStart,tStop,tStep)
+  IC = Data_dict["InitialConditions"]
+
+  #Define Parameter values
+  if isempty(Param)
+    DF = Data_dict
+  else
+
+    kmax = Param[1:6]
+    K = Param[7:12]
+    ke = Param[13]
+
+    # Rewrite Data Dictionary
+    Data_dict["ReactionRateVector"] = kmax
+    Data_dict["SaturationConstantVector"] = K
+    Data_dict["EnzymeRate"] = ke
+    DF = Data_dict
+  end
+
+  #RunSolver
+  f(t,x) = Balances(t,x,DF)
+  #x = Sundials.cvode(f,IC,t,reltol=1e-3,abstol=1e-6)
+  t,y = ode23(f,IC,t)
+
+  # compute x array -
+  e1 = map(y -> y[1], y);
+  e2 = map(y -> y[2], y);
+  e3 = map(y -> y[3], y);
+  e4 = map(y -> y[4], y);
+  e5 = map(y -> y[5], y);
+  e6 = map(y -> y[6], y);
+  Ax = map(y -> y[7], y);
+  Bx = map(y -> y[8], y);
+  Cx = map(y -> y[9], y);
+  Biox = map(y -> y[10], y);
+
+  x = [e1 e2 e3 e4 e5 e6 Ax Bx Cx Biox]
+  return t,x
+end

### 2016-05-19 biochemical/make_corrupted_measurment_data.jl

@@ -1,0 +1,27 @@
+include("SolveBalances.jl")
+
+#DefineTime
+tStart = 0.0;
+tStop = 100.0;
+tStep = 0.1;
+
+#loadParameterValues
+Param = Float64[]
+
+#SolverResults
+(t,x) = SolveBalances(tStart,tStop,tStep,Param)
+
+#Define species from SolverResults
+Ax = x[:,7]
+Bx = x[:,8]
+Cx = x[:,9]
+Biox = x[:,10]
+
+# write data files -
+Ax = Ax.*(1.0)
+Bx = Bx.*(1.0)
+Cx = Cx.*(1+2.0)
+Biox = Biox.*(1+0.5)
+
+data = [t Ax Bx Cx Biox]
+writedlm("./data/CORRUPTED_MEASUREMENT_SET.dat",data)

### 2016-05-19 biochemical/sample_ensemble.jl

@@ -1,0 +1,7 @@
+# Run the model for the ensemble of solutions -
+
+# Load the ensemble files from disk -
+pc_array_full = load("./data/pc_array.dat")
+rank_array = load("./data/rank_array.dat")
+
+idx_rank = find(rank_array==0)

### 2016-05-19 biochemical/write_ensemble.jl

@@ -1,0 +1,13 @@
+using POETs
+
+# Script to write the model ensemble to disk -
+ec_array = ec_array[:,2:end]
+pc_array = pc_array[:,2:end]
+
+# Re-rank -
+rank_array = rank_function(ec_array)
+
+# Write -
+writedlm("./data/ec_array.dat",ec_array)
+writedlm("./data/pc_array.dat",pc_array)
+writedlm("./data/rank_array.dat",rank_array)

### 2016-05-19 jupoets/MakePerformancePlot.m

@@ -1,0 +1,15 @@
+% Octave script to nake performannce plot -
+
+% Load the data -
+data_array = load("performance.dat")
+
+# make groups -
+hold on
+bar(data_array(:,1))
+
+# Generate SF -
+SF = 2.678/sqrt(10)
+
+# make error bars -
+h = errorbar(data_array(:,1),SF*data_array(:,2),'k.')
+set(h,"linestyle","none")

### 2016-05-19 jupoets/binh_korn_lib.jl

@@ -1,0 +1,83 @@
+# some global parameters -
+BIG = 1e10
+
+# Evaluates the objective function values -
+function objective_function(parameter_array)
+
+  # Alias the species -
+  x = parameter_array[1]
+  y = parameter_array[2]
+
+  # Calculate the objective function array -
+  obj_array = BIG*ones(2,1)
+
+  # f1 and f2 -
+  obj_array[1] = 4.0*(x^2)+4.0*(y^2)
+  obj_array[2] = (x - 5)^2 + (y - 5)^2
+
+  # Constraints are implemented as a penaltly on obj value
+  lambda_value = 100.0
+
+  # How much do we violate the constraints?
+  violation_constraint_1 = 25 - (x-5.0)^2 - y^2
+  violation_constraint_2 = (x-8.0)^2 + (y+3.0)^2 - 7.7
+  penaltly_array = zeros(2)
+  penaltly_array[1] = lambda_value*(min(0,violation_constraint_1))^2
+  penaltly_array[2] = lambda_value*(min(0,violation_constraint_2))^2
+
+  # return the obj_array -
+  return obj_array+penaltly_array
+end
+
+# Generates new parameter array, given current array -
+function neighbor_function(parameter_array)
+
+  SIGMA = 0.05
+  number_of_parameters = length(parameter_array)
+
+  # calculate new parameters -
+  new_parameter_array = parameter_array.*(1+SIGMA*randn(number_of_parameters))
+
+  # Check the bound constraints -
+  LOWER_BOUND = [0,0]
+  UPPER_BOUND = [5,3]
+
+  # return the corrected parameter arrays -
+  return parameter_bounds_function(new_parameter_array,LOWER_BOUND,UPPER_BOUND)
+end
+
+function acceptance_probability_function(rank_array,temperature)
+
+  return (exp(-rank_array[end]/temperature))
+end
+
+function cooling_function(temperature)
+
+  # define my new temperature -
+  alpha = 0.9
+  return 0.9*temperature
+end
+
+# Helper functions -
+function parameter_bounds_function(parameter_array,lower_bound_array,upper_bound_array)
+
+  # reflection_factor -
+  epsilon = 0.01
+
+  # iterate through and fix the parameters -
+  new_parameter_array = copy(parameter_array)
+  for (index,value) in enumerate(parameter_array)
+
+    lower_bound = lower_bound_array[index]
+    upper_bound = upper_bound_array[index]
+
+    if (value<lower_bound)
[…]

### 2016-05-19 jupoets/fonseca_fleming_lib.jl

@@ -1,0 +1,85 @@
+# some global parameters -
+BIG = 1e10
+
+# Evaluates the objective function values -
+function objective_function(parameter_array)
+
+  # Alias the species -
+  number_of_parameters = length(parameter_array)
+
+  # Calculate the objective function array -
+  obj_array = BIG*ones(2,1)
+
+  # calculate the objectibe array -
+  # calculate the sums
+  sum_1 = 0.0
+  sum_2 = 0.0
+  for index in collect(1:number_of_parameters)
+
+    sum_1 = sum_1 + (parameter_array[index] - 1/sqrt(number_of_parameters))^2
+    sum_2 = sum_2 + (parameter_array[index] + 1/sqrt(number_of_parameters))^2
+  end
+
+  # objectives -
+  obj_array[1] = 1 - exp(-1*sum_1)
+  obj_array[2] = 1 - exp(-1*sum_2)
+
+  # return -
+  return obj_array
+end
+
+# Generates new parameter array, given current array -
+function neighbor_function(parameter_array)
+
+  SIGMA = 0.05
+  number_of_parameters = length(parameter_array)
+
+  # correct for zero -
+  idx_zero = find(parameter_array.==0.0)
+  parameter_array[idx_zero] = randn(length(idx_zero))
+
+  # calculate new parameters -
+  new_parameter_array = parameter_array.*(1+SIGMA*randn(number_of_parameters))
+
+  # Check the bound constraints -
+  LOWER_BOUND = -4.0
+  UPPER_BOUND = 4.0
+
+  # return the corrected parameter arrays -
+  return parameter_bounds_function(new_parameter_array,LOWER_BOUND*ones(number_of_parameters),UPPER_BOUND*ones(number_of_parameters))
+end
+
+function acceptance_probability_function(rank_array,temperature)
+  return (exp(-rank_array[end]/temperature))
+end
+
+function cooling_function(temperature)
+
+  # define my new temperature -
+  alpha = 0.9
+  return alpha*temperature
+end
+
+
+# Helper functions -
+function parameter_bounds_function(parameter_array,lower_bound_array,upper_bound_array)
+
+  # reflection_factor -
+  epsilon = 0.01
+
+  # iterate through and fix the parameters -
+  new_parameter_array = copy(parameter_array)
+  for (index,value) in enumerate(parameter_array)
+
+    lower_bound = lower_bound_array[index]
+    upper_bound = upper_bound_array[index]
+
[…]

### 2016-05-19 jupoets/plot_binh_korn.jl

@@ -1,0 +1,26 @@
+using PyPlot
+using POETs
+
+# load the data -
+ec_array = readdlm("./ec_array_binh_korn.dat")
+pc_array = readdlm("./pc_array_binh_korn.dat")
+
+# re-rank the solutions -
+rank_array = rank_function(ec_array[:,2:end])
+
+
+# idx_rank_1 = find(rank_array.>=1.0)
+# plot(ec_array[1,idx_rank_1],ec_array[2,idx_rank_1],".",color="0.75")
+#
+# # Plot rank 0 in black -
+# idx_rank_0 = find(rank_array.==0.0)
+# plot(ec_array[1,idx_rank_0],ec_array[2,idx_rank_0],"k.")
+
+figure()
+
+idx_rank_1 = find(rank_array.>=1.0)
+plot(pc_array[1,idx_rank_1],pc_array[2,idx_rank_1],".",color="0.75")
+
+# Plot rank 0 in black -
+idx_rank_0 = find(rank_array.==0.0)
+plot(pc_array[1,idx_rank_0],pc_array[2,idx_rank_0],"k.")

### 2016-05-19 jupoets/plot_fonseca_fleming.jl

@@ -1,0 +1,26 @@
+using PyPlot
+using POETs
+
+# load the data -
+ec_array = readdlm("./ec_array_ff.dat")
+pc_array = readdlm("./pc_array_ff.dat")
+
+# re-rank the solutions -
+rank_array = rank_function(ec_array[:,2:end])
+
+
+#idx_rank_1 = find(rank_array.>=1.0)
+#plot(ec_array[1,idx_rank_1],ec_array[2,idx_rank_1],".",color="0.75")
+
+# Plot rank 0 in black -
+#idx_rank_0 = find(rank_array.==0.0)
+#plot(ec_array[1,idx_rank_0],ec_array[2,idx_rank_0],"k.")
+
+figure()
+
+idx_rank_1 = find(rank_array.>=1.0)
+plot3D(pc_array[1,idx_rank_1],pc_array[2,idx_rank_1],pc_array[3,idx_rank_1],".",color="0.75")
+
+# Plot rank 0 in black -
+idx_rank_0 = find(rank_array.==0.0)
+plot3D(pc_array[1,idx_rank_0],pc_array[2,idx_rank_0],pc_array[3,idx_rank_0],"k.")

### 2016-05-19 jupoets/plot_schaffer.jl

@@ -1,0 +1,26 @@
+using PyPlot
+using POETs
+
+# load the data -
+ec_array = readdlm("./ec_array_schaffer.dat")
+pc_array = readdlm("./pc_array_schaffer.dat")
+
+# re-rank the solutions -
+rank_array = rank_function(ec_array[:,2:end])
+
+
+idx_rank_1 = find(rank_array.>=1.0)
+#plot(ec_array[1,idx_rank_1],ec_array[2,idx_rank_1],".",color="0.75")
+
+# Plot rank 0 in black -
+idx_rank_0 = find(rank_array.==0.0)
+#plot(ec_array[1,idx_rank_0],ec_array[2,idx_rank_0],"k.")
+
+figure()
+
+idx_rank_1 = find(rank_array.>=1.0)
+semilogx(transpose(1:length(idx_rank_1)),pc_array[1,idx_rank_1],".",color="0.75")
+
+# Plot rank 0 in black -
+idx_rank_0 = find(rank_array.==0.0)
+semilogx(transpose(1:length(idx_rank_0)),pc_array[1,idx_rank_0],"k.")

### 2016-05-19 jupoets/run_fonseca_fleming_test.jl

@@ -1,0 +1,13 @@
+include("test_fonseca_fleming.jl")
+
+
+number_of_trials = 10
+etime_array = zeros(number_of_trials)
+for index in collect(1:number_of_trials)
+
+  tic()
+  test_fonseca_fleming()
+  elapsed_time = toc()
+  etime_array[index] = elapsed_time
+
+end

### 2016-05-19 jupoets/run_schaffer_test.jl

@@ -1,0 +1,13 @@
+include("test_schaffer.jl")
+
+
+number_of_trials = 10
+etime_array = zeros(number_of_trials)
+for index in collect(1:number_of_trials)
+
+  tic()
+  test_schaffer()
+  elapsed_time = toc()
+  etime_array[index] = elapsed_time
+
+end

### 2016-05-19 jupoets/schaffer_lib.jl

@@ -1,0 +1,80 @@
+# some global parameters -
+BIG = 1e10
+
+# Evaluates the objective function values -
+function objective_function(parameter_array)
+
+  # Alias the species -
+  x = parameter_array[1]
+
+  # Calculate the objective function array -
+  obj_array = BIG*ones(2,1)
+
+  # f1 and f2 -
+  obj_array[1] = (x^2)
+  obj_array[2] = (x - 2)^2
+
+  # Constraints are implemented as a penaltly on obj value
+  lambda_value = 100.0
+
+  # How much do we violate the constraints?
+  penaltly_array = zeros(2)
+
+  # return the obj_array -
+  return obj_array+penaltly_array
+end
+
+# Generates new parameter array, given current array -
+function neighbor_function(parameter_array)
+
+  SIGMA = 0.05
+  number_of_parameters = length(parameter_array)
+
+  # calculate new parameters -
+  new_parameter_array = parameter_array.*(1+SIGMA*randn(number_of_parameters))
+
+  # Check the bound constraints -
+  LOWER_BOUND = -10
+  UPPER_BOUND = 10
+
+  # return the corrected parameter arrays -
+  return parameter_bounds_function(new_parameter_array,LOWER_BOUND,UPPER_BOUND)
+end
+
+function acceptance_probability_function(rank_array,temperature)
+
+  return (exp(-rank_array[end]/temperature))
+
+end
+
+function cooling_function(temperature)
+
+  # define my new temperature -
+  alpha = 0.9
+  return 0.9*temperature
+end
+
+
+# Helper functions -
+function parameter_bounds_function(parameter_array,lower_bound_array,upper_bound_array)
+
+  # reflection_factor -
+  epsilon = 0.01
+
+  # iterate through and fix the parameters -
+  new_parameter_array = copy(parameter_array)
+  for (index,value) in enumerate(parameter_array)
+
+    lower_bound = lower_bound_array[index]
+    upper_bound = upper_bound_array[index]
+
+    if (value<lower_bound)
+      new_parameter_array[index] = lower_bound
+    elseif (value>upper_bound)
+      new_parameter_array[index] = upper_bound
+    end
+  end
+
+  return new_parameter_array
+
+end

### 2016-05-19 jupoets/test_binh_korn.jl

@@ -1,0 +1,25 @@
+# Script to generate test data from the bihn and korn function
+include("binh_korn_lib.jl")
+
+function test_binh_korn()
+
+  number_of_subdivisions = 10
+  initial_state_array = zeros(2)
+  x_initial_array = linspace(0.01,5,number_of_subdivisions)
+  y_initial_array = linspace(0.01,3,number_of_subdivisions)
+  ec_array = zeros(2)
+  pc_array = zeros(2)
+  for index in collect(1:number_of_subdivisions)
+
+    initial_state_array[1] = float(x_initial_array[index])
+    initial_state_array[2] = float(y_initial_array[index])
+    (EC,PC,RA) = estimate_ensemble(objective_function,neighbor_function,acceptance_probability_function,cooling_function,initial_state_array;rank_cutoff=4,maximum_number_of_iterations=20,show_trace=false)
+
+    #@show size(RA)
+
+    ec_array = [ec_array EC]
+    pc_array = [pc_array PC]
+  end
+
+  return (ec_array,pc_array)
+end

### 2016-05-19 jupoets/test_fonseca_fleming.jl

@@ -1,0 +1,27 @@
+# Script to generate test data from the bihn and korn function
+include("fonseca_fleming_lib.jl")
+
+function test_fonseca_fleming()
+
+  number_of_subdivisions = 10
+  number_of_states = 3
+  initial_state_array = zeros(number_of_states)
+  x_initial_array = linspace(-4,4,number_of_subdivisions)
+  iz = find(x_initial_array==0.0)
+  x_initial_array[iz] = randn(length(iz))
+
+  ec_array = zeros(2)
+  pc_array = zeros(number_of_states)
+  for index in collect(1:number_of_subdivisions)
+
+    initial_state_array[1] = float(x_initial_array[index])
+    initial_state_array[2] = float(x_initial_array[index])
+    initial_state_array[3] = float(x_initial_array[index])
+    (EC,PC,RA) = estimate_ensemble(objective_function,neighbor_function,acceptance_probability_function,cooling_function,initial_state_array;rank_cutoff=4,maximum_number_of_iterations=20,show_trace=false)
+
+    ec_array = [ec_array EC]
+    pc_array = [pc_array PC]
+
+  end
+  return (ec_array,pc_array)
+end

### 2016-05-19 jupoets/test_schaffer.jl

@@ -1,0 +1,23 @@
+# Script to generate test data from the bihn and korn function
+include("schaffer_lib.jl")
+
+function test_schaffer()
+
+  number_of_subdivisions = 10
+  initial_state_array = zeros(1)
+  x_initial_array = linspace(-10,10,number_of_subdivisions)
+  ec_array = zeros(2)
+  pc_array = zeros(1)
+  for index in collect(1:number_of_subdivisions)
+
+    initial_state_array[1] = float(x_initial_array[index])
+    (EC,PC,RA) = estimate_ensemble(objective_function,neighbor_function,acceptance_probability_function,cooling_function,initial_state_array;rank_cutoff=4,maximum_number_of_iterations=20,show_trace=false)
+
+    #@show size(RA)
+
+    ec_array = [ec_array EC]
+    pc_array = [pc_array PC]
+  end
+
+  return (ec_array,pc_array)
+end

### 2016-05-21 biochemical/Balances.jl
Added biochemical proof of concept

@@ -1,10 +1,10 @@
 include("Kinetics.jl")
 include("Control.jl")
 
-function Balances(t,x,DF)
+function Balances(t,x,dxdt_vector,DF)
 
-  idx = find(x.<0)
-  x[idx] = 0.0
+  idx = find(x.<1e-9)
+  x[idx] = 1e-9
 
   # Alias the species vector -
   e1 = x[1];
@@ -41,11 +41,11 @@ function Balances(t,x,DF)
   mu = sum(Growth_rate)
 
   # Enzyme Balances
-  dxdt_vector = zeros(10)
+  #dxdt_vector = zeros(10)
   dxdt_vector[1:num_modes] = alpha + Enzyme_rate - (beta+mu)*x[1:num_modes]
 
   # Metabolite Balances
   dxdt_vector[7:10] = S[1:4,:]*Z*Reaction_rate*Biox
 
-  return dxdt_vector
+  #return dxdt_vector
 end

### 2016-05-21 biochemical/Data_Cybernetic.jl

@@ -8,7 +8,7 @@ stm_matrix = float(readdlm("./data/Network.csv",','));
 
 #============================#
 # EnzymeRateParameters
-  ke = .5
+  ke = 0.5
   alpha = 0.004
   beta = 0.05
 
@@ -22,6 +22,8 @@ kmax = [
   0.276;
   0.0955;
   ];
+
+#kmax = kmax.*(1+0.5*randn(length(kmax)))
 
 #============================#
 #Ksat
@@ -37,16 +39,16 @@ K = [
 #=============================#
 #Initial Conditions
   x0 = [
-    0.5; # em1
-    0.5; # em2
-    0.5; # em3
-    0.5; # em4
-    0.5; # em5
-    0.5; # em6
-    2; # Ax
-    0; # Bx
-    0; # Cx
-    0.01; # Biox
+    0.5;  # 1 em1
+    0.5;  # 2 em2
+    0.5;  # 3 em3
+    0.5;  # 4 em4
+    0.5;  # 5 em5
+    0.5;  # 6 em6
+    2;    # 7 Ax
+    0;    # 8 Bx
+    0;    # 9 Cx
+    0.01; # 10 Biox
   ];
 
 #=======================================#

### 2016-05-21 biochemical/Kinetics.jl

@@ -18,11 +18,11 @@ function Kinetics(t,x,DF)
   Biox = x[10];
 
   #Parametrs from DF
-  kmax = DF["ReactionRateVector"]
-  K = DF["SaturationConstantVector"]
-  ke = DF["EnzymeRate"]
-  alpha = DF["EnzymeSynthesis"]
-  beta = DF["Degradation"]
+  kmax = abs(DF["ReactionRateVector"])
+  K = abs(DF["SaturationConstantVector"])
+  ke = abs(DF["EnzymeRate"])
+  alpha = abs(DF["EnzymeSynthesis"])
+  beta = abs(DF["Degradation"])
   Z = DF["ModeMatrix"]
   S = DF["MetaboliteMatrix"]
 

### 2016-05-21 biochemical/SolveBalances.jl

@@ -1,6 +1,6 @@
 include("Data_Cybernetic.jl")
 include("Balances.jl")
-using ODE
+using Sundials
 
 function SolveBalances(tStart,tStop,tStep,Param)
 
@@ -28,22 +28,22 @@ function SolveBalances(tStart,tStop,tStep,Param)
   end
 
   #RunSolver
-  f(t,x) = Balances(t,x,DF)
-  #x = Sundials.cvode(f,IC,t,reltol=1e-3,abstol=1e-6)
-  t,y = ode23(f,IC,t)
+  f(t,x,dxdt) = Balances(t,x,dxdt,DF)
+  x = Sundials.cvode(f,IC,t,reltol=1e-3,abstol=1e-6)
+  #t,y = ode23(f,IC,t;reltol=1e-3,abstol=1e-6)
 
-  # compute x array -
-  e1 = map(y -> y[1], y);
-  e2 = map(y -> y[2], y);
-  e3 = map(y -> y[3], y);
-  e4 = map(y -> y[4], y);
-  e5 = map(y -> y[5], y);
-  e6 = map(y -> y[6], y);
-  Ax = map(y -> y[7], y);
-  Bx = map(y -> y[8], y);
-  Cx = map(y -> y[9], y);
-  Biox = map(y -> y[10], y);
-
-  x = [e1 e2 e3 e4 e5 e6 Ax Bx Cx Biox]
+  # # compute x array -
+  # e1 = map(y -> y[1], y);
+  # e2 = map(y -> y[2], y);
+  # e3 = map(y -> y[3], y);
+  # e4 = map(y -> y[4], y);
+  # e5 = map(y -> y[5], y);
+  # e6 = map(y -> y[6], y);
+  # Ax = map(y -> y[7], y);
+  # Bx = map(y -> y[8], y);
+  # Cx = map(y -> y[9], y);
+  # Biox = map(y -> y[10], y);
+  #
+  # x = [e1 e2 e3 e4 e5 e6 Ax Bx Cx Biox]
   return t,x
 end

### 2016-05-21 biochemical/hcmem_lib.jl

@@ -12,10 +12,243 @@ BIG = 1e10
 SMALL = 1e-6
 
 # Globally load data so we don't have load on each iteration -
-MEASURED_ARRAY = readdlm("./data/CORRUPTED_MEASUREMENT_SET.dat")
+MEASURED_ARRAY_1 = readdlm("./data/MEASUREMENT_SET_1.dat")
+MEASURED_ARRAY_2 = readdlm("./data/MEASUREMENT_SET_2.dat")
+MEASURED_ARRAY_3 = readdlm("./data/MEASUREMENT_SET_3.dat")
+MEASURED_ARRAY_4 = readdlm("./data/MEASUREMENT_SET_4.dat")
 
 # Evaluates the objective function values -
-@debug function objective_function(parameter_array)
+function local_refienment_step(parameter_array)
+
+  SIGMA = 0.05
+
+  # initialize -
+  number_of_parameters = length(parameter_array)
+
+  # Setup the bound constraints -
+  LOWER_BOUND = SMALL
+  UPPER_BOUND = [
+    1.00  ; # 1 k1
+    1.00  ; # 2 k2
+    1.00  ; # 3 k3
+    1.00  ; # 4 k4
+    1.00  ; # 5 k5
+    1.00  ; # 6 k6
+    3.00  ; # 7 K1
+    3.00  ; # 8 K2
+    3.00  ; # 9 K3
+    3.00  ; # 10 K4
+    3.00  ; # 11 K5
+    3.00  ; # 12 K6
+    1.00  ; # 13 ke
+  ]
+
+  # calculate the starting error -
+  parameter_array_best = parameter_array
+  error_array = BIG*ones(4)
+  error_array[1] = sum(objective_function(parameter_array_best))
+
+  # main refinement loop -
+  iteration_counter = 1
+  iteration_max = 1000
+  while (iteration_counter<iteration_max)
+
+    # take a step up -
+    parameter_up = parameter_array_best.*(1+SIGMA*rand(number_of_parameters))
+    parameter_up = parameter_bounds_function(parameter_up,LOWER_BOUND*ones(number_of_parameters),UPPER_BOUND)
+
+    # take a step down -
+    parameter_down = parameter_array_best.*(1-SIGMA*rand(number_of_parameters))
+    parameter_down = parameter_bounds_function(parameter_down,LOWER_BOUND*ones(number_of_parameters),UPPER_BOUND)
+
+    # Evaluate the obj function -
[…]

### 2016-05-21 biochemical/make_corrupted_measurment_data.jl

@@ -20,8 +20,8 @@ Biox = x[:,10]
 # write data files -
 Ax = Ax.*(1.0)
 Bx = Bx.*(1.0)
-Cx = Cx.*(1+2.0)
-Biox = Biox.*(1+0.5)
+Cx = Cx.*(1.0)
+Biox = Biox.*(1.0)
 
 data = [t Ax Bx Cx Biox]
-writedlm("./data/CORRUPTED_MEASUREMENT_SET.dat",data)
+writedlm("./data/MEASUREMENT_SET_4.dat",data)

### 2016-05-21 biochemical/run.jl

@@ -21,9 +21,9 @@ Biox = x[:,10]
 using PyPlot
 figure(1)
 plot(t,Ax,color="k",label="HCM EM",linewidth=2)
-plot(t,Bx, color="k",linewidth=2)
-plot(t,Cx, color="k",linewidth=2)
-plot(t,Biox, color="k",linewidth=2)
+plot(t,Bx, color="r",linewidth=2)
+plot(t,Cx, color="b",linewidth=2)
+plot(t,Biox, color="g",linewidth=2)
 legend(fontsize=18)
 xlabel("Time (hr)",fontsize=20)
 ylabel("Abundance (mM)",fontsize=20)

### 2016-05-21 biochemical/run_biochemical_test.jl

@@ -8,22 +8,34 @@ function test_biochemical_model()
   number_of_parameters = 13
   number_of_objectives = 4
 
+  # generate parameter guess -
+  #initial_parameter_array = readdlm("./data/initial_parameter_array.dat")
   initial_parameter_array = zeros(13)
+  initial_parameter_array[1:6]  = ones(6).*(1+0.25*randn(6))
+  initial_parameter_array[7:12] = 2*ones(6).*(1+0.25*randn(6))
+  initial_parameter_array[13] = (1+0.25*randn())
+
   ec_array = zeros(number_of_objectives)
   pc_array = zeros(number_of_parameters)
   for index in collect(1:number_of_subdivisions)
 
-    # Grab a starting point -
-    initial_parameter_array[1:6]  = ones(6).*(1+0.25*randn(6))
-    initial_parameter_array[7:12] = 2*ones(6).*(1+0.25*randn(6))
-    initial_parameter_array[13] = (1+0.25*randn())
-
     # Run JuPOETs -
-    (EC,PC,RA) = estimate_ensemble(objective_function,neighbor_function,acceptance_probability_function,cooling_function,initial_parameter_array;rank_cutoff=4,maximum_number_of_iterations=10,show_trace=true)
+    (EC,PC,RA) = estimate_ensemble(objective_function,neighbor_function,acceptance_probability_function,cooling_function,initial_parameter_array;rank_cutoff=4,maximum_number_of_iterations=20,show_trace=true)
 
     # Package -
     ec_array = [ec_array EC]
     pc_array = [pc_array PC]
+
+    # Take the *best* value from the current ensemble, refine it and go around again -
+    total_error = sum(ec_array[:,2:end],1)
+
+    # Which col is the min error?
+    min_index = indmin(total_error)
+    @show index,total_error[min_index]
+
+    # Refine the best solution -
+    initial_parameter_array = pc_array[:,min_index].*(1+0.15*randn(number_of_parameters))
+    initial_parameter_array = local_refienment_step(initial_parameter_array)
   end
 
   return (ec_array,pc_array)

### 2016-05-21 biochemical/sample_ensemble.jl

@@ -1,7 +1,63 @@
 # Run the model for the ensemble of solutions -
+include("SolveBalances.jl")
+using PyPlot
+using PyCall
+@pyimport numpy as np
+
 
 # Load the ensemble files from disk -
-pc_array_full = load("./data/pc_array.dat")
-rank_array = load("./data/rank_array.dat")
+pc_array_full = readdlm("./data/pc_array.dat")
+rank_array = readdlm("./data/rank_array.dat")
 
-idx_rank = find(rank_array==0)
+# Select the desired rank -
+idx_rank = find(rank_array .<= 0.0)
+
+# Setup time scale -
+tStart = 0.0;
+tStop = 100.0;
+tStep = 0.1;
+number_of_timesteps = 200
+time_experimental = linspace(tStart,tStop,number_of_timesteps)
+
+# initialize data array -
+data_array = zeros(length(time_experimental),1)
+
+# Run the simulation for this rank -
+number_of_samples = 1
+for (index,rank_index_value) in enumerate(idx_rank)
+
+  if (mod(index,20)==0)
+
+    # Grab the parameter set from the cache -
+    parameter_array = pc_array_full[:,rank_index_value]
+
+    # Run the model -
+    (t,x) = SolveBalances(tStart,tStop,tStep,parameter_array)
+
+    # Need to interpolate the simulation onto the experimental time scale -
+    AI = np.interp(time_experimental[:],t,x[:,7])
+    BI = np.interp(time_experimental[:],t,x[:,8])
+    CI = np.interp(time_experimental[:],t,x[:,9])
+    XI = np.interp(time_experimental[:],t,x[:,10])
+
+    # grab -
+    data_array = [data_array CI]
+
+    # update the sample count -
+    number_of_samples = number_of_samples + 1
+  end
+end
+
+# calculate mean, and std
+mean_value = mean(data_array,2)
+std_value = std(data_array,2)
+SF = (2.58/sqrt(number_of_samples))
+
+UB = mean_value + (SF)*std_value
+LB = mean_value - (SF)*std_value
+idx_z = find(LB.<0)
+LB[idx_z] = 0.0
+
+# Make the plot -
+#plot(time_experimental,mean_value,"k")
+fill_between(vec(time_experimental),vec(LB),vec(UB),color="gray",lw=2)

### 2016-05-21 biochemical/write_ensemble.jl

@@ -11,3 +11,11 @@ rank_array = rank_function(ec_array)
 writedlm("./data/ec_array.dat",ec_array)
 writedlm("./data/pc_array.dat",pc_array)
 writedlm("./data/rank_array.dat",rank_array)
+
+# Take the *best* value from the current ensemble, refine it and go around again -
+total_error = sum(ec_array[:,2:end],1)
+
+# Which col is the min error?
+min_index = indmin(total_error)
+initial_parameter_array = pc_array[:,min_index]
+writedlm("./data/initial_parameter_array.dat",initial_parameter_array)

### 2016-05-21 Paper_v1.tex
Working on Biochemical results

@@
 ]{\inits{DB}\fnm{David} \snm{Bassen}}
+\author[
+  addressref={aff1}
+]{\inits{MV}\fnm{Michael} \snm{Vilkhovoy}}
+\author[
+  addressref={aff1}
+]{\inits{MM}\fnm{Mason} \snm{Minot}}
 \author[
@@
 The central idea underlying POETs is a mapping between the value of the objective functions evaluated at $\mathbf{p}_{i+1}$ (parameter guess at iteration $i+1$) and the Pareto rank.
-We computed the Pareto rank of $\mathbf{p}_{i+1}$ by comparing the simulation error at iteration $i+1$ against an archive of simulation errors compiled up to iteration $i$ $\mathbf{E}_{i}$.
+Pareto rank is a measure of distance from the trade-off surface;
+Parameter sets on or near the optimal trade-off surface between the objectives have a rank equal to $0$ (no other current parameter sets are better).
+Sets with increasing non-zero rank are progressively further away from the optimal trade-off surface.
+Thus, a parameter set with a $rank = 0$ is \emph{better} in a trade-off sense than $rank>0$.
+We computed the Pareto rank of $\mathbf{p}_{i+1}$ by comparing the simulation error at iteration $i+1$ against an archive of simulation errors compiled up to iteration $i$ $\mathcal{O}_{i}$.
 We used the Fonseca and Fleming scheme to compute the Pareto rank, $r$ \cite{RANKING}:
@@
 \end{equation} where $r$ denotes the rank, or the number of parameter sets that dominate parameter set $\mathbf{p}_{i+1}$.
-Parameter sets on the optimal trade-off surface between the objectives have a rank equal to $0$ (no other current parameter sets are better).
-Sets with increasing non-zero rank are progressively further away from the optimal trade-off surface.
-Parameter sets on the optimal trade-off surface have a rank equal to $0$ (no other current parameter sets are better).
-Sets with increasing non-zero rank are progressively farther away from the optimal trade-off surface.
-Thus, a parameter set with a $rank = 0$ is \emph{better} in a trade-off sense than $rank>0$.
 We used the Pareto rank to inform the SA calculation.
@@
 We compared the performance of JuPOETs versus POETs implemented in Octave for a suite of multiobjective test functions (Table \ref{table:objective-table}).
-The execution time (wall-clock) for JuPOETs and Octave-POETs was measured for 10 independent trials for the suite of test problems.
-The same cooling, neighbor, acceptance, search and objective logic was employed between the implementations, with all other parameters held constant.
-For each test function, the search domain was partitioned into 10 segments, where initial parameter guesses were drawn from each partition.
+The wall-clock time for JuPOETs and Octave-POETs was measured for 10 independent trials for the test problems.
+The same cooling, neighbor, acceptance, search and objective logic was employed between the implementations, and all other parameters were held constant.
+For each test function, the search domain was partitioned into 10 segments, where an initial parameter guess was drawn from each partition.
 The number of search steps for each temperate $\mathcal{I}$ = 10 for all cases, and the cooling parameter was $\alpha$ = 0.9.
@@
 Taken together, these results suggested that JuPOETs was able to estimate an ensemble of solutions to constrained multiobjective optimization problems significantly faster than the current Octave implementation.
-However, we have not demonstrated JuPOETs on a biochemically relevant test problem.
+Next, we tested JuPOETs on a proof-of-concept biochemical model identification problem.
 
-[JuPOETs and a proof-of-principle HCM-EM model goes here]
+JuPOETs estimated an ensemble of biochemical models that was consistent with the mean of synthetic training data (Fig. \ref{fig:fig-biochemical}).
+Four synthetic data sets were generated from a prototypical biochemical network consisting of 6 metabolites and 7 reactions (Fig. \ref{fig:fig-biochemical}, inset right).
+We considered a typical case in which the same measurements were made on four different hypothetical cell types, each having the same biological connectivity but different performance.
+Network dynamics were modeled using the hybrid cybernetic model with elementary modes (HCM-EM) approach of Ramkrishna and coworkers \cite{2008_kim_varner_ramkrishna_BiotechProg}.
+In the HCM-EM approach, metabolic networks are first decomposed into a set of elementary modes (EMs) (chemically balanced steady-state pathways, see \cite{Schuster:2000aa}).
+Combinations of elementary modes are then used to characterize network behavior.
+Each elementary mode is catalyzed by a pseudo enzyme; thus, each mode has both kinetic and enzyme synthesis parameters.
+The proof of concept network generated 6 EMs, resulting in 13 model parameters.
+The synthetic data was generated by randomly varying these parameters.
+JuPOETs produced an ensemble of models that captured the mean of the measured data sets for extracellular metabolites and cellmass
+(Fig. \ref{fig:fig-biochemical}A and B). The 95\% confidence estimate produced by the ensemble was consistent with mean of the measured data, despite having significant uncertainty in the data.
 
@@
 \section*{Author's contributions}
-    J.V developed the software presented in this study. The manuscript was prepared and edited for publication by D.B, J.B. and J.V.
+    J.V developed the software presented in this study. M.M and M.V developed the proof-of-concept biochemical model.
+    The manuscript was prepared and edited for publication by D.B, J.B. and J.V.
 
@@
 
-\begin{figure}
+\begin{figure}[h]
   \caption{Schematic of multiobjective parameter mapping.
@@
 
-\begin{figure}
+\begin{figure}[h]
   \caption{The performance of JuPOETs on the multi-objective test suite.
@@
 
-\begin{figure}
+\begin{figure}[h]
   \caption{Representative JuPOETs solutions for problems in the multi-objective test suite.
@@
 }\label{fig:fig-surfaces}
+\end{figure}
+
+\begin{figure}[h]
+  \caption{Proof of concept biochemical network study.
+  Inset: Prototypical biochemical network with six metabolites and seven reactions modeled using the hybrid cybernetic approach (HCM).
+  Intracellular cellmass precursors $A,B$, and $C$ are balanced (no accumulation) while the extracellular metabolites $A_{e},B_{e}$, and $C_{e}$ are dynamic.
+  The oval denotes the cell boundary, $q_{j}$ is the $j$th flux across the boundary, and $v_{k}$ denotes the $k$th intracellular flux.
+  Four data sets (each with $A_{e},B_{e}$,$C_{e}$ and cellmass measurements) were generated by varying the kinetic constant for each biochemical modes.
+  Each data set was a single objective.
+  A: Ensemble simulation of extracellular substrate $A_{e}$ and cellmass versus time.
+  B: Ensemble simulation of extracellular substrate $B_{e}$ and $C_{e}$ versus time.
+  The gray region denotes the 95\% confidence estimate of the mean ensemble simulation.
+  The data points denote mean synthetic measurements, while the error bars denote the 95\% confidence estimate of the measurement computed over the four training data sets.
+  C: Trade-off plots between the four training objectives. The quantity $O_{j}$ denotes the jth training objective.
+  Each point represents a member of the parameter ensemble, where gray denotes rank 0 sets, while black denotes rank 1 sets.
+   }\label{fig:fig-biochemical}
 \end{figure}

### 2016-05-22 Paper_v1.tex
Last paragraph in RD

@@
   \street{Cornell University},                     %
-  \postcode{14806}                              % post or zip code
+  \postcode{14853}                              % post or zip code
   \city{Ithaca NY},                             % city
@@
 
-POETs integrate Simulated Annealing (SA) with Pareto optimality to estimate parameter sets on or near the optimal tradeoff surface between competing training objectives
+POETs integrates simulated annealing (SA) with Pareto optimality to estimate parameter sets on or near the optimal tradeoff surface between competing training objectives
 (Fig. \ref{fig:fig-MOSAalgorithm}).
@@
 Only parameter sets with rank less than or equal to a user defined threshold (rank$\leq$4 by default) were included in $\mathbf{K}_{i}$ to ensure that we characterized the neighborhood near the trade-off surface (Algorithm \ref{algo:JuPOETs-pcode}).
+Parameter bounds constraints were implemented in the neighbor function while problem specific constraints were treated using the penalty method where each objective function was
+augmented by the total of the constraint violation:
+\begin{equation}
+  O_{i}+\sum_{j=1}^{\mathcal{C}}\lambda_{j}\min\left(0,g_j\left(t,\mathbf{x}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p}\right)\right)\qquad~i=1,\hdots,\mathcal{K}
+\end{equation}
 
@@
 JuPOETs produced an ensemble of models that captured the mean of the measured data sets for extracellular metabolites and cellmass
-(Fig. \ref{fig:fig-biochemical}A and B). The 95\% confidence estimate produced by the ensemble was consistent with mean of the measured data, despite having significant uncertainty in the data.
+(Fig.
+\ref{fig:fig-biochemical}A and B).
+The 95\% confidence estimate produced by the ensemble was consistent with mean of the measured data, despite having significant uncertainty in the data.
+JuPOETs produced a consensus estimate of the synthetic data by calculating optimal trade-offs between the synthetic data sets (Fig.
+\ref{fig:fig-biochemical}C).
+Thus, JuPOETs produced an ensemble of parameter sets that gives the mean of the training data for conflicting data sets.
+
+JuPOETs is a significant advancement over Octave-POETs, both in terms of performance and capability.
+However, there are several areas that could be explored further.
+First, JuPOETs should be compared with other multiobjective evolutionary algorithms (MOEAs) to determine its relative performance on test and real world problems.
+Many evolutionary approaches e.g., the nondominated sorting genetic algorithm (NSGA) family of algorithms, have been adapted to solve multiobjective optimization problems [REF].
+It is unclear if JuPOETs will perform as well as these other approaches;
+one potential advantage that JuPOETs may have is the local refinement step which temporarily reduces the problem to a single objective formulation.
+Previously, this hybrid approach led to better convergence on a proof-of-concept signal transduction model [REF].
+Next, JuPOETs should take advantage of the native parallel execution capabilities of Julia to accelerate the evaluation of the objective functions.
+For many real world parameter estimation problems, the bulk of the execution time is spent evaluating the objective functions.
+One strategy to improve performance could be to optimize surrogates [REF], while another would be parallel execution of the objective functions.
+Currently, JuPOETs serially evaluate the components of the objective function vector.
+However, because of the flexible function pointer architecture of JuPOETs, a shift to parallel evaluation of the objective functions requires changes to only the user defined objective function and not the main run loop.
+Thus, parallel evaluation of objective functions could be easily implemented using a variety of techniques without changing to JuPOETs.
 

### 2016-05-22 jupoets/binh_korn_lib.jl

@@ -24,9 +24,10 @@ function objective_function(parameter_array)
   penaltly_array = zeros(2)
   penaltly_array[1] = lambda_value*(min(0,violation_constraint_1))^2
   penaltly_array[2] = lambda_value*(min(0,violation_constraint_2))^2
+  penaltly_total = sum(penaltly_array)
 
   # return the obj_array -
-  return obj_array+penaltly_array
+  return obj_array+penaltly_total*ones(2)
 end
 
 # Generates new parameter array, given current array -

### 2016-05-23 Paper_v1.tex
Almost there ...

@@
 \usepackage[linesnumbered]{algorithm2e}
+\usepackage{courier}
 %\usepackage{algorithm}
-%\RequirePackage{natbib}
+\RequirePackage{natbib}
 %\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
-%\RequirePackage{hyperref}
+\RequirePackage{hyperref}
 \usepackage[utf8]{inputenc} %unicode support
@@
 \begin{abstract} % abstract
-\parttitle{First part title} %if any
-Text for this section.
+Ensemble modeling is a well established approach for dealing with parameter uncertainty in large-scale deterministic models.
+In this study, we present an open-source implementation of the Pareto Optimal Ensemble Technique in the Julia programming language (JuPOETs).
+JuPOETs integrates simulated annealing with Pareto optimality to estimate parameter sets on or near the optimal tradeoff surface between competing training objectives.
+JuPOETs can be installed using the Julia package manager from the JuPOETs GitHub repository at https://github.com/varnerlab/POETs.jl.
 
@@
 \begin{keyword}
-\kwd{sample}
-\kwd{article}
-\kwd{author}
+\kwd{Ensemble~modeling}
+\kwd{Multiobjective~optimization}
+\kwd{Julia}
 \end{keyword}
@@
 \section*{Background}
-Models of signal transduction networks may exhibit complex relationships between model performance and parameter values \cite{Brown:2004}.
-It is often not possible to uniquely identify parameters from noisy experimental measurements, even when given extensive training data \cite{Gadkar:2005ad}.
-Thus, despite significant advances in standardizing model identification \cite{Gennemark:2009on},
-the core problem of estimating model parameters from noisy data remains.
+Ensemble modeling is a well established approach for dealing with parameter uncertainty in large-scale deterministic models.
+It is often not possible to uniquely identify biochemical model parameters from noisy experimental measurements, even when given extensive training data \cite{Gadkar:2005ad}.
+Thus, despite significant advances in standardizing biochemical model identification \cite{Gennemark:2009on},
+the problem of estimating model parameters from experimental data remains challenging.
 Ensemble approaches have been used to address parameter uncertainty in systems biology and other fields like weather prediction \cite{Battogtokh:2002,Kuepfer:2007, Brown:2003,Palmer:2005}.
-In an ensemble approach, a \emph{family} of approximate parameter sets is identified instead of a single best-fit parameter set.
-Parameter families can be selected based on simulation error, along with other criterion such as diversity.
+In an ensemble approach, a family of parameter sets is identified instead of a single best-fit parameter set.
+Parameter families can be selected based upon simulation error, along with other criterion such as diversity.
 Simulations using parameter ensembles can estimate confidence intervals on model variables, and robustly constrain model predictions,
@@
 
-Identification of large models with hundreds or even thousands of states and parameters may not be tractable as a single objective formulation.
-Models at these scales require significant training data perhaps taken from diverse sources, for example different laboratories or cell-lines.
-These diverse training data are likely heterogenous, and contain intrinsic conflicts.
-On the other hand, large models can also be constructed from many smaller modules.
-For example, Diamond and coworkers developed a model of phosphoinositide and calcium signaling in human platelets composed of many
+Identification of large biochemical models with hundreds or even thousands of states and parameters may not be tractable as a single objective formulation.
+Large models require significant training data perhaps taken from diverse sources, for example different laboratories or cell-lines.
+These diverse training data are likely heterogenous, and contain intrinsic conflicts which complicate parameter estimation.
+On the other hand, large models can also be constructed from many smaller modules, each of which is independently identified.
+For example, Diamond and coworkers developed a model of calcium signaling in human platelets composed of many
 separate submodels identified using different types of training data \cite{Purvis:2008}.
-Taken together, the problem of identifying parameter ensembles for large-scale models with many parameters posses a unique challenge.
 Parameter ensemble estimation techniques which optimally balance tradeoffs between submodels or conflicts in training data
-might lead to more robust model performance. One class of such techniques is multiobjective optimization.
+lead to robust model performance. One class of such techniques is multiobjective optimization.
 Previously, we developed the Pareto Optimal Ensemble Technique (POETs) algorithm to address the challenge of competing or conflicting objectives.
-POETs, which integrates simulated annealing and multiobjective optimization through the notion of Pareto rank, finds a family of parameter estimates which optimally trade-off between
+POETs, which integrates simulated annealing (SA) and multiobjective optimization through the notion of Pareto rank, finds a family of parameter estimates which optimally trade-off between
 competing (and potentially conflicting) experimental objectives \cite{Song:2010aa}.
@@
 However, the previous implementation of POETs, in the Octave programming language \cite{Octave}, suffered from poor performance, and did not offer user definable
-functions such as custom cooling schedules, parameter constraints, or custom search logic. It was also not well integrated into a modern package or
-source code management (SCM) system.
+functions to accommodate custom cooling schedules, parameter constraints, or custom search logic. Octave-POETs was also not well integrated into a modern package or
+source code management (SCM) system, thus upgrades containing new features, or bug fixes were not centrally managed.
 
 \section*{Implementation}
-
-In this study, we present JuPOETs, an open-source implementation of the Pareto Optimal Ensemble Technique (POETs) in the Julia programming language.
-JuPOETs offers many advantages and improvements compared to the previous implementation of POETs.
+In this study, we present an open-source implementation of the Pareto optimal ensemble technique in the Julia programming language (JuPOETs).
+JuPOETs offers many advantages and improvements when compared to Octave-POETs.
 First, JuPOETs takes advantage of the unique performance features of the Julia programming language.
@@
   \mathcal{L} \leq \mathbf{p} \leq \mathcal{U}
-\end{equation}
-The term t denotes time, $g_i\left(t,\mathbf{x}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p}\right)$ denotes the ith constraint on the search, where
+\end{equation}using a modified simulated annealing algorithm.
+The quantity t denotes time, $g_i\left(t,\mathbf{x}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p}\right)$ denotes the ith constraint on the search, where
 $\mathbf{x}\left(t,\mathbf{p}\right)$ denotes the model state variable vector with an initial state $\mathbf{x}_{0}$, and $\mathbf{u}(t)$ is an input vector.
@@
 
-POETs integrates simulated annealing (SA) with Pareto optimality to estimate parameter sets on or near the optimal tradeoff surface between competing training objectives
-(Fig. \ref{fig:fig-MOSAalgorithm}).
-The central idea underlying POETs is a mapping between the value of the objective functions evaluated at $\mathbf{p}_{i+1}$ (parameter guess at iteration $i+1$) and the Pareto rank.
+JuPOETs integrates simulated annealing with Pareto optimality to estimate parameter sets on or near the optimal tradeoff surface between competing training objectives
+(Fig. \ref{fig:fig-MOSAalgorithm} and Algorithm \ref{algo:JuPOETs-pcode}).
+The central idea underlying POETs is a mapping between the value of the objective functions evaluated at $\mathbf{p}_{i+1}$ (parameter guess at iteration $i+1$) and Pareto rank.
+JuPOETs calculates the performance of a candidate parameter set $\mathbf{p}_{i+1}$ by calling the user defined objective function, \texttt{objective}. The function \texttt{objective} takes
+a parameter set as an input argument and returns a $\mathcal{K}\times{1}$ error vector. Candidate parameter sets are generated by the user supplied \texttt{neighbor} function.
+The error vector associated with $\mathbf{p}_{i+1}$ is ranked using the builtin Pareto \texttt{rank} function, by comparing the current error at iteration i+1 to the error archive $\mathcal{O}_{i}$.
 Pareto rank is a measure of distance from the trade-off surface;
-Parameter sets on or near the optimal trade-off surface between the objectives have a rank equal to $0$ (no other current parameter sets are better).
+parameter sets on or near the optimal trade-off surface between the objectives have a rank equal to $0$ (no other current parameter sets are better).
 Sets with increasing non-zero rank are progressively further away from the optimal trade-off surface.
 Thus, a parameter set with a $rank = 0$ is \emph{better} in a trade-off sense than $rank>0$.
-We computed the Pareto rank of $\mathbf{p}_{i+1}$ by comparing the simulation error at iteration $i+1$ against an archive of simulation errors compiled up to iteration $i$ $\mathcal{O}_{i}$.
-We used the Fonseca and Fleming scheme to compute the Pareto rank, $r$ \cite{RANKING}:
+We implemented the Fonseca and Fleming ranking scheme in the \texttt{rank} function which computes the Pareto rank, $r$ \cite{RANKING}:
 \begin{equation}\label{eqn_rank}
-rank\left(\mathbf{p}_{i+1}\mid \mathbf{E}_{i}\right)=r
-\end{equation} where $r$ denotes the rank, or the number of parameter sets that dominate parameter set $\mathbf{p}_{i+1}$.
+\texttt{rank}\left(\mathbf{p}_{i+1}\mid \mathcal{O}_{i}\right)=r
+\end{equation} where rank is the number of parameter sets that dominate (are better than) parameter set $\mathbf{p}_{i+1}$.
 We used the Pareto rank to inform the SA calculation.
-The parameter set $\mathbf{k}_{i+1}$ was accepted or rejected by the SA using the acceptance probability $\mathcal{P}\left(\mathbf{k}_{i+1}\right)$:
+The parameter set $\mathbf{p}_{i+1}$ was accepted or rejected by the SA, by calculating an \texttt{acceptance} probability $\mathcal{P}\left(\mathbf{p}_{i+1}\right)$:
 \begin{equation}\label{eqn_costMOSA}
-\mathcal{P}(\mathbf{k}_{i+1}) \equiv \exp{\{-rank\left(\mathbf{k}_{i+1} \mid \mathbf{K}_{i} \right)/T\}}
+\mathcal{P}(\mathbf{p}_{i+1}) \equiv \exp{\{-\texttt{rank}\left(\mathbf{p}_{i+1} \mid \mathcal{S}_{i} \right)/T\}}
 \end{equation}
-where $T$ is the computational annealing temperature. As $rank\left(\mathbf{k}_{i+1}\mid \mathbf{K}_{i}\right)\rightarrow{0}$, the acceptance probability moved toward one,
+where $T$ is the computational annealing temperature. As $\texttt{rank}\left(\mathbf{p}_{i+1}\mid \mathcal{O}_{i}\right)\rightarrow{0}$, the acceptance probability moved toward one,
 ensuring that we explored parameter sets along the Pareto surface.
@@
 However, as $T$ was reduced, the probability of this occurring decreased.
-Parameter sets could be accepted by the SA and $\emph{not}$ archived in $\mathbf{K}_{i}$.
-Only parameter sets with rank less than or equal to a user defined threshold (rank$\leq$4 by default) were included in $\mathbf{K}_{i}$ to ensure that we characterized the neighborhood near the trade-off surface (Algorithm \ref{algo:JuPOETs-pcode}).
-Parameter bounds constraints were implemented in the neighbor function while problem specific constraints were treated using the penalty method where each objective function was
-augmented by the total of the constraint violation:
+Parameter sets could be accepted by the SA and $\emph{not}$ archived in $\mathcal{S}_{i}$.
+Only parameter sets with rank less than or equal to threshold (rank$\leq$4 by default) were included in $\mathcal{S}_{i}$
+to ensure that we characterized the neighborhood near the trade-off surface.
+Parameter bounds were implemented in the \texttt{neighbor} function as box constraints, while problem specific constraints were implemented in \texttt{objective} using a penalty method:
 \begin{equation}
-  O_{i}+\sum_{j=1}^{\mathcal{C}}\lambda_{j}\min\left(0,g_j\left(t,\mathbf{x}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p}\right)\right)\qquad~i=1,\hdots,\mathcal{K}
-\end{equation}
+  O_{i}+\lambda\sum_{j=1}^{\mathcal{C}}\min\left(0,g_j\left(t,\mathbf{x}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p}\right)\right)\qquad~i=1,\hdots,\mathcal{K}
+\end{equation}where $\lambda$ denotes the penalty parameter ($\lambda$ = 100 by default).
+However, because both the neighbor and objective functions are user defined, different constraint implementations are easily defined.
+JuPOETs can be installed using the Julia package manager, see instructions at the JuPOETs GitHub repository (https://github.com/varnerlab/POETs.jl).
 

### 2016-05-23 Paper_v1.tex
Dinner break ...

@@
 
-\subsection*{Optimization problem formulation.}
+\subsection*{JuPOETs optimization problem formulation.}
 JuPOETs solves the $\mathcal{K}-$dimensional constrained multiobjective optimization problem:
@@
   \mathcal{L} \leq \mathbf{p} \leq \mathcal{U}
-\end{equation}using a modified simulated annealing algorithm.
-The quantity t denotes time, $g_i\left(t,\mathbf{x}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p}\right)$ denotes the ith constraint on the search, where
-$\mathbf{x}\left(t,\mathbf{p}\right)$ denotes the model state variable vector with an initial state $\mathbf{x}_{0}$, and $\mathbf{u}(t)$ is an input vector.
-The terms $\mathbf{f}(t,\mathbf{x}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p})$ denote the system of model equations (e.g., differential equations,
-differential algebraic equations or algebraic constraints) and $\mathbf{p}$ denotes the unknown parameter vector ($\mathcal{P}\times~1$).
-The parameter search can be subject to parameter bound constraints where $\mathcal{L}$ and $\mathcal{U}$ denote the lower and upper parameter bounds, respectively.
+\end{equation}using a modified simulated annealing approach.
+The quantity $t$ denotes time, $\mathbf{x}\left(t,\mathbf{p}\right)$ denotes the model state (with an initial state $\mathbf{x}_{0}$), and $\mathbf{u}(t)$ denotes an input vector.
+The terms $\mathbf{f}(t,\mathbf{x}(t,\mathbf{p}),\dot{\mathbf{x}}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p})$ denote the system of model equations (e.g., differential equations,
+differential algebraic equations or linear/non-linear algebraic equations) where $\mathbf{p}$ denotes the unknown parameter vector ($\mathcal{D}\times~1$).
+The parameter search can be subject to parameter bound constraints, where $\mathcal{L}$ and $\mathcal{U}$ denote the lower and upper parameter bounds, respectively
+as well as $\mathcal{C}$ problem specific constraints $g_i\left(t,\mathbf{x}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p}\right),i=1,\hdots,\mathcal{C}$.
 
@@
 (Fig. \ref{fig:fig-MOSAalgorithm} and Algorithm \ref{algo:JuPOETs-pcode}).
-The central idea underlying POETs is a mapping between the value of the objective functions evaluated at $\mathbf{p}_{i+1}$ (parameter guess at iteration $i+1$) and Pareto rank.
-JuPOETs calculates the performance of a candidate parameter set $\mathbf{p}_{i+1}$ by calling the user defined objective function, \texttt{objective}. The function \texttt{objective} takes
-a parameter set as an input argument and returns a $\mathcal{K}\times{1}$ error vector. Candidate parameter sets are generated by the user supplied \texttt{neighbor} function.
-The error vector associated with $\mathbf{p}_{i+1}$ is ranked using the builtin Pareto \texttt{rank} function, by comparing the current error at iteration i+1 to the error archive $\mathcal{O}_{i}$.
+The central idea of POETs is a mapping between the value of the objective vector evaluated at $\mathbf{p}_{i+1}$ (parameter guess at iteration $i+1$) and Pareto rank.
+JuPOETs calculates the performance of a candidate parameter set $\mathbf{p}_{i+1}$ by calling the user defined \texttt{objective} function; \texttt{objective} takes
+a parameter set as an input and returns a $\mathcal{K}\times{1}$ objective vector. Candidate parameter sets are generated by the user supplied \texttt{neighbor} function.
+The error vector associated with $\mathbf{p}_{i+1}$ is ranked using the builtin Pareto \texttt{rank} function, by comparing the current error at iteration $i$+1 to the error archive $\mathcal{O}_{i}$.
 Pareto rank is a measure of distance from the trade-off surface;
@@
 Sets with increasing non-zero rank are progressively further away from the optimal trade-off surface.
-Thus, a parameter set with a $rank = 0$ is \emph{better} in a trade-off sense than $rank>0$.
-We implemented the Fonseca and Fleming ranking scheme in the \texttt{rank} function which computes the Pareto rank, $r$ \cite{RANKING}:
+Thus, a parameter set with a rank$=0$ is \emph{better} in a trade-off sense than rank$>0$.
+We implemented the Fonseca and Fleming ranking scheme \cite{RANKING} in the builtin \texttt{rank} function:
 \begin{equation}\label{eqn_rank}
 \texttt{rank}\left(\mathbf{p}_{i+1}\mid \mathcal{O}_{i}\right)=r
-\end{equation} where rank is the number of parameter sets that dominate (are better than) parameter set $\mathbf{p}_{i+1}$.
+\end{equation} where rank $r$ is the number of parameter sets that dominate (are better than) parameter set $\mathbf{p}_{i+1}$.
 We used the Pareto rank to inform the SA calculation.
@@
 \end{equation}
-where $T$ is the computational annealing temperature. As $\texttt{rank}\left(\mathbf{p}_{i+1}\mid \mathcal{O}_{i}\right)\rightarrow{0}$, the acceptance probability moved toward one,
-ensuring that we explored parameter sets along the Pareto surface.
+where $T$ is the computational annealing temperature.
+As $\texttt{rank}\left(\mathbf{p}_{i+1}\mid \mathcal{O}_{i}\right)\rightarrow{0}$, the acceptance probability moves toward one,
+ensuring that we explore parameter sets along the Pareto surface.
 Occasionally (depending upon $T$) a parameter set with a high Pareto rank was accepted by the SA allowing a more diverse search of the parameter space.
-However, as $T$ was reduced, the probability of this occurring decreased.
-Parameter sets could be accepted by the SA and $\emph{not}$ archived in $\mathcal{S}_{i}$.
-Only parameter sets with rank less than or equal to threshold (rank$\leq$4 by default) were included in $\mathcal{S}_{i}$
-to ensure that we characterized the neighborhood near the trade-off surface.
+However, as $T$ is reduced, the probability of accepting a high-rank set occurring decreases.
+Parameter sets could also be accepted by the SA but $\emph{not}$ permanently archived in $\mathcal{S}_{i}$.
+Only parameter sets with rank less than or equal to threshold (rank$\leq$4 by default) were included in $\mathcal{S}_{i}$, where the archive was re-ranked and filtered after
+every new parameter set was accepted.
 Parameter bounds were implemented in the \texttt{neighbor} function as box constraints, while problem specific constraints were implemented in \texttt{objective} using a penalty method:
 \begin{equation}
-  O_{i}+\lambda\sum_{j=1}^{\mathcal{C}}\min\left(0,g_j\left(t,\mathbf{x}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p}\right)\right)\qquad~i=1,\hdots,\mathcal{K}
+  O_{i}+\lambda\sum_{j=1}^{\mathcal{C}}\min\Bigl\{0,g_j\left(t,\mathbf{x}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p}\right)\Bigr\}\qquad~i=1,\hdots,\mathcal{K}
 \end{equation}where $\lambda$ denotes the penalty parameter ($\lambda$ = 100 by default).
 However, because both the neighbor and objective functions are user defined, different constraint implementations are easily defined.
-JuPOETs can be installed using the Julia package manager, see instructions at the JuPOETs GitHub repository (https://github.com/varnerlab/POETs.jl).
+JuPOETs can be installed using the Julia package manager from the JuPOETs repository at https://github.com/varnerlab/POETs.jl.
 
@@
 \vspace{0.05in}
-\caption{Pseudo-code for the run-loop of the JuPOETs algorithm. The user specifies the neighbor, acceptance, cooling and objective function pointers along with a initial parameter guess.
+\caption{Pseudo-code for the main run-loop of JuPOETs.
+The user specifies the \texttt{neighbor}, \texttt{acceptance}, \texttt{cooling} and \texttt{objective} functions along with an initial parameter guess.
 The rank archive $\mathcal{R}$, solution archive $\mathcal{S}$ and objective archive $\mathcal{O}$ are initialized from the initial guess.
-The initial guess is then perturbed in the neighbor function, which generates a new solution whose performance is evaluated using the user supplied objective function.
-The new solution and objective values are then added to the respective archives and ranked using the builtin rank function.
-If the new solution accepted (based upon a probability calculated with the user supplied acceptance function) it is added to the solution and objective archive.
+The initial guess is perturbed in the \texttt{neighbor} function, which generates a new solution whose performance is evaluated using the user supplied \texttt{objective} function.
+The new solution and objective values are then added to the respective archives and ranked using the builtin \texttt{rank} function.
+If the new solution is accepted (based upon a probability calculated with the user supplied \texttt{acceptance} function) it is added to the solution and objective archive.
 This solution is then perturbed during the next iteration of the algorithm.
 However, if the solution is not accepted, it is removed from the archive and discarded.
-The computational temperature is adjusted using the user supplied cooling function after each $\mathcal{I}$ iterations.
+The computational temperature is adjusted using the user supplied \texttt{cooling} function after each $\mathcal{I}$ iterations.
 }\label{algo:JuPOETs-pcode}
 \end{algorithm}
+
+%(Fig. \ref{fig:fig-timing} and Fig. \ref{fig:fig-surfaces}).
 
@@
 
-JuPOETs identified solutions on or near the optimal tradeoff surface significantly faster than Octave-POETs for the suite of test problems
-(Fig. \ref{fig:fig-timing} and Fig. \ref{fig:fig-surfaces}).
-We compared the performance of JuPOETs versus POETs implemented in Octave for a suite of multiobjective test functions (Table \ref{table:objective-table}).
-The wall-clock time for JuPOETs and Octave-POETs was measured for 10 independent trials for the test problems.
-The same cooling, neighbor, acceptance, search and objective logic was employed between the implementations, and all other parameters were held constant.
+JuPOETs identified optimal or nearly optimal solutions significantly faster than Octave-POETs for a suite of test problems (Table \ref{table:objective-table}).
+The wall-clock time for JuPOETs and Octave-POETs was measured for 10 independent trials for each of the test problems.
+The same \texttt{cooling}, \texttt{neighbor}, \texttt{acceptance}, and \texttt{objective} logic was employed between the implementations, and all other parameters were held constant.
 For each test function, the search domain was partitioned into 10 segments, where an initial parameter guess was drawn from each partition.
-The number of search steps for each temperate $\mathcal{I}$ = 10 for all cases, and the cooling parameter was $\alpha$ = 0.9.
+The number of search steps for each temperate was $\mathcal{I}$ = 10 for all cases, and the cooling parameter was $\alpha$ = 0.9.
 On average, JuPOETs identified optimal or near optimal solutions for the suite of test problems six-fold faster (60s versus 400s)
-than Octave-POETs (Fig. \ref{fig:fig-timing}). In addition, JuPOETs produced the characteristic tradeoff curves for each test problem, given both parameter bound and
-problem constraints (Fig.
-\ref{fig:fig-surfaces}).
-Taken together, these results suggested that JuPOETs was able to estimate an ensemble of solutions to constrained multiobjective optimization problems significantly faster than the current Octave implementation.
+than Octave-POETs (Fig. \ref{fig:fig-timing}). JuPOETs produced the characteristic tradeoff curves for each test problem, given both parameter bound and
+problem constraints (Fig. \ref{fig:fig-surfaces}).
+Thus, JuPOETs estimated an ensemble of solutions to constrained multiobjective optimization test problems significantly faster than the current Octave implementation.
 Next, we tested JuPOETs on a proof-of-concept biochemical model identification problem.
@@
 JuPOETs estimated an ensemble of biochemical models that was consistent with the mean of synthetic training data (Fig. \ref{fig:fig-biochemical}).
-Four synthetic data sets were generated from a prototypical biochemical network consisting of 6 metabolites and 7 reactions (Fig. \ref{fig:fig-biochemical}, inset right).
-We considered a typical case in which the same measurements were made on four different hypothetical cell types, each having the same biological connectivity but different performance.
+Four synthetic training data sets were generated from a prototypical biochemical network consisting of 6 metabolites and 7 reactions (Fig. \ref{fig:fig-biochemical}, inset right).
+We considered a common case in which the same measurements were made on four hypothetical cell types, each having the same biological connectivity but different performance.
 Network dynamics were modeled using the hybrid cybernetic model with elementary modes (HCM-EM) approach of Ramkrishna and coworkers \cite{2008_kim_varner_ramkrishna_BiotechProg}.
 In the HCM-EM approach, metabolic networks are first decomposed into a set of elementary modes (EMs) (chemically balanced steady-state pathways, see \cite{Schuster:2000aa}).
-Combinations of elementary modes are then used to characterize network behavior.
+Dynamic combinations of elementary modes are then used to characterize network behavior.
 Each elementary mode is catalyzed by a pseudo enzyme; thus, each mode has both kinetic and enzyme synthesis parameters.
@@
 JuPOETs produced an ensemble of models that captured the mean of the measured data sets for extracellular metabolites and cellmass
-(Fig.
-\ref{fig:fig-biochemical}A and B).
-The 95\% confidence estimate produced by the ensemble was consistent with mean of the measured data, despite having significant uncertainty in the data.
-JuPOETs produced a consensus estimate of the synthetic data by calculating optimal trade-offs between the synthetic data sets (Fig.
+(Fig. \ref{fig:fig-biochemical}A and B). The 95\% confidence estimate produced by the ensemble was consistent with mean of the measured data, despite having significant
+uncertainty in the training data.
+JuPOETs produced a consensus estimate of the synthetic data by calculating optimal trade-offs between the training data sets (Fig.
 \ref{fig:fig-biochemical}C).
-Thus, JuPOETs produced an ensemble of parameter sets that gives the mean of the training data for conflicting data sets.
+Thus, JuPOETs produced an ensemble of parameter sets that gave the mean of the training data for conflicting data sets.
 

### 2016-05-24 Paper_v1.tex
First draft

@@
 \begin{abstract} % abstract
-Ensemble modeling is a well established approach for dealing with parameter uncertainty in large-scale deterministic models.
-In this study, we present an open-source implementation of the Pareto Optimal Ensemble Technique in the Julia programming language (JuPOETs).
-JuPOETs integrates simulated annealing with Pareto optimality to estimate parameter sets on or near the optimal tradeoff surface between competing training objectives.
+Ensemble modeling is a well established approach for obtaining robust predictions and course graining population behavior in large deterministic mathematical models.
+In this study, we present the Pareto Optimal Ensemble Technique in the Julia programming language (JuPOETs).
+JuPOETs integrates simulated annealing with Pareto optimality to estimate an ensemble of parameter sets on or near the optimal tradeoff surface between competing training objectives.
+We demonstrated JuPOETs on a suite of multiobjective test problems, with both bounds and problem constraints as well as a proof-of-concept biochemical model identification problem.
+JuPOETs identified optimal or near optimal solutions approximately six-fold faster than a corresponding implementation in Octave. 
 JuPOETs can be installed using the Julia package manager from the JuPOETs GitHub repository at https://github.com/varnerlab/POETs.jl.
-
-\parttitle{Second part title} %if any
-Text for this section.
 \end{abstract}
@@
 %%
+
+%Sethna and coworkers later showed that model performance is often controlled by only a few parameter combinations,
+%a characteristic seemingly universal to multi-parameter models referred to as \textit{sloppiness} \citep{Machta:2013aa}.
+%n the other hand, large models can also be constructed from many smaller modules, each of which is independently identified.
+%For example, Diamond and coworkers developed a model of calcium signaling in human platelets composed of many
+%separate submodels identified using different types of training data \cite{Purvis:2008}.
+
 \section*{Background}
-Ensemble modeling is a well established approach for dealing with parameter uncertainty in large-scale deterministic models.
-It is often not possible to uniquely identify biochemical model parameters from noisy experimental measurements, even when given extensive training data \cite{Gadkar:2005ad}.
+Ensemble modeling is a well established approach for obtaining robust predictions and course graining population behavior in large deterministic mathematical models.
+It is often not possible to uniquely identify all the parameters in biochemical models, even when given extensive training data \cite{Gadkar:2005ad}.
 Thus, despite significant advances in standardizing biochemical model identification \cite{Gennemark:2009on},
 the problem of estimating model parameters from experimental data remains challenging.
-Ensemble approaches have been used to address parameter uncertainty in systems biology and other fields like weather prediction \cite{Battogtokh:2002,Kuepfer:2007, Brown:2003,Palmer:2005}.
-In an ensemble approach, a family of parameter sets is identified instead of a single best-fit parameter set.
-Parameter families can be selected based upon simulation error, along with other criterion such as diversity.
+Ensemble approaches address parameter uncertainty in systems biology and other fields like weather prediction \cite{Battogtokh:2002,Kuepfer:2007, Brown:2003,Palmer:2005}
+by using parameter families instead of single best-fit parameter sets.
+Parameter families can be selected based upon simulation error, along with other criterion such as diversity or steady-state peformance.
 Simulations using parameter ensembles can estimate confidence intervals on model variables, and robustly constrain model predictions,
 despite having many poorly constrained parameters \cite{Gutenkunst2007,Song:2009b}.
-There are several techniques to generate parameter ensembles.
+There are many techniques to generate parameter ensembles.
 Battogtokh et al., Brown et al., and later Tasseff et al. generated experimentally constrained parameter ensembles using a
-Metropolis-type random walk through parameter space \cite{Battogtokh:2002,Brown:2003,Tasseff:2010aa,Tasseff:2011aa}.
-Other strategies could also be adapted to generate parameter ensembles \cite{Moles:2003}.
-However, the unifying component of all of these previous strategies was the minimization of a single objective function.
+Metropolis-type random walk \cite{Battogtokh:2002,Brown:2003,Tasseff:2010aa,Tasseff:2011aa}.
+Liao and coworkers developed methods that generate ensembles that all approach the same steady-state, for example one determined by fluxomics measurements \cite{Tran:2008aa}.
+They have used this approach for model reduction \cite{Tan:2011aa}, strain engineering \cite{Contador2009221,Tan:2012aa} and to study the robustness of non-native pathways and network failure \cite{Lee:2014aa}.
+Maranas and coworkers have also applied this method to develop a comprehensive kinetic model of bacterial central carbon metabolism,
+including mutant data \cite{Khodayari:2014aa}.
+We and others have used ensemble approaches, generated using optimization and sampling approaches, to robustly simulate a wide variety of signal transduction processes \citep{Luan:2007aa,Song:2009aa,Tasseff:2010aa,Tasseff:2011aa,Nayak:2011aa}, neutrophil trafficking in sepsis \cite{Song:2012aa},
+patient specific coagulation behavior \cite{Luan:2010aa} and to capture cell to cell variation \cite{Lequieu:2011aa}.
+Thus, ensemble approaches are widely used to robustly simulate a variety of biochemical systems.
 
-Identification of large biochemical models with hundreds or even thousands of states and parameters may not be tractable as a single objective formulation.
+Identification of biochemical models with hundreds or even thousands of states and parameters may not be tractable as a single objective optimization problem.
 Large models require significant training data perhaps taken from diverse sources, for example different laboratories or cell-lines.
-These diverse training data are likely heterogenous, and contain intrinsic conflicts which complicate parameter estimation.
-On the other hand, large models can also be constructed from many smaller modules, each of which is independently identified.
-For example, Diamond and coworkers developed a model of calcium signaling in human platelets composed of many
-separate submodels identified using different types of training data \cite{Purvis:2008}.
-Parameter ensemble estimation techniques which optimally balance tradeoffs between submodels or conflicts in training data
-lead to robust model performance. One class of such techniques is multiobjective optimization.
+These data are often heterogenous, and contain intrinsic conflicts which complicate parameter estimation.
+Parameter ensembles which optimally balance tradeoffs between submodels or conflicts in training data can
+lead to robust model performance. One class of ensemble generation technique that naturally balances conflicting training data is multiobjective optimization.
 Previously, we developed the Pareto Optimal Ensemble Technique (POETs) algorithm to address the challenge of competing or conflicting objectives.
-POETs, which integrates simulated annealing (SA) and multiobjective optimization through the notion of Pareto rank, finds a family of parameter estimates which optimally trade-off between
+POETs, which integrates simulated annealing (SA) and multiobjective optimization through the notion of Pareto rank, estimates parameter ensembles which optimally trade-off between
 competing (and potentially conflicting) experimental objectives \cite{Song:2010aa}.
-POETs has been used in several biochemical and signal transduction studies to estimate parameter ensembles \cite{Song:2009aa,Lequieu:2011aa}.
-However, the previous implementation of POETs, in the Octave programming language \cite{Octave}, suffered from poor performance, and did not offer user definable
-functions to accommodate custom cooling schedules, parameter constraints, or custom search logic. Octave-POETs was also not well integrated into a modern package or
-source code management (SCM) system, thus upgrades containing new features, or bug fixes were not centrally managed.
+However, the previous implementation of POETs, in the Octave programming language \cite{CITE_Octave}, suffered from poor performance and was not configurable.
+For example, Octave-POETs does not accommodate user definable objective functions, bounds and problem constraints, cooling schedules, different variable types e.g., a mixture of binary and continuous design variables or custom diversity generation routines.
+Octave-POETs was also not well integrated into a package or
+source code management (SCM) system. Thus, upgrades to the approach containing new features, or bug fixes were not centrally managed.
 
@@
 In this study, we present an open-source implementation of the Pareto optimal ensemble technique in the Julia programming language (JuPOETs).
-JuPOETs offers many advantages and improvements when compared to Octave-POETs.
-First, JuPOETs takes advantage of the unique performance features of the Julia programming language.
-Julia, which has performance comparable to C but with syntax similar to MATLAB/Octave and Python, is a cross-platform, high-performance programming language for technical computing \cite{Julia}.
-Julia offers a sophisticated compiler, distributed parallel execution, numerical accuracy, and an extensive mathematical function library.
-Additionally, Julia offers a built-in package manager which is directly integrated with GitHub, a popular web-based Git repository hosting service which offers distributed revision control and source code management.
+JuPOETs offers many advantages and improvements compared to Octave-POETs.
+JuPOETs takes advantage of the unique features of Julia.
+Julia is a cross-platform, high-performance programming language for technical computing that has performance comparable to C but with syntax similar to MATLAB/Octave and Python \cite{Julia}.
+Julia also offers a sophisticated compiler, distributed parallel execution, numerical accuracy, and an extensive function library.
+Additionally, Julia offers a built-in package manager which is directly integrated with GitHub,
+a popular web-based Git repository hosting service offering distributed revision control and source code management.
 Next, because Julia can natively call other languages such as Python or C, JuPOETs can be used with models implemented in a variety of languages on many platforms.
-Lastly, the architecture of JuPOETs takes advantage of the first-class function type in Julia allowing user definable behavior for all key aspects of the algorithm, including objective functions, custom search logic, linear/non-linear parameter constraints (and parameter bounds constraints) as well as custom cooling schedule functions.
-Thus, JuPOETs can easily be adapted to solve many problem types, including mixed binary and continuous variable types, without the need to change the base algorithm (which was not true of the previous POETs implementation).
+Lastly, the architecture of JuPOETs takes advantage of the first-class function type in Julia allowing user definable behavior for all key aspects of the algorithm, including objective functions, custom diversity generation logic, linear/non-linear parameter constraints (and parameter bounds constraints) as well as custom cooling schedules.
+Thus, JuPOETs can be adapted to many problem types, including mixed binary and continuous variable types, bilevel problems and constrained problems without changing the base algorithm (which was not true of the previous POETs implementation).
 
@@
 
-JuPOETs identified optimal or nearly optimal solutions significantly faster than Octave-POETs for a suite of test problems (Table \ref{table:objective-table}).
+JuPOETs identified optimal or nearly optimal solutions significantly faster than Octave-POETs for a suite of multiobjective test problems (Table \ref{table:objective-table}).
 The wall-clock time for JuPOETs and Octave-POETs was measured for 10 independent trials for each of the test problems.
@@
 The synthetic data was generated by randomly varying these parameters.
-JuPOETs produced an ensemble of models that captured the mean of the measured data sets for extracellular metabolites and cellmass
-(Fig. \ref{fig:fig-biochemical}A and B). The 95\% confidence estimate produced by the ensemble was consistent with mean of the measured data, despite having significant
+JuPOETs produced an ensemble of parameters that captured the mean of the measured data sets for extracellular metabolites and cellmass
+(Fig. \ref{fig:fig-biochemical}A and B). The 95\% confidence estimate produced by the ensemble was consistent with the mean of the measured data, despite having significant
 uncertainty in the training data.
@@
 \ref{fig:fig-biochemical}C).
-Thus, JuPOETs produced an ensemble of parameter sets that gave the mean of the training data for conflicting data sets.
+Thus, JuPOETs produced an ensemble of parameters that gave the mean of the training data for conflicting data sets.
 
-JuPOETs is a significant advancement over Octave-POETs, both in terms of performance and capability.
-However, there are several areas that could be explored further.
+\section*{Conclusions}
+
+JuPOETs is a significant advance over the previous POETs implementation.
+It offers improved performance and is highly adaptable to different problem types.
+We demonstrated JuPOETs on a suite of test problems, and a proof-of-concept biochemical model.
+However, there are several areas that could be explored further to improve JuPOETs.
 First, JuPOETs should be compared with other multiobjective evolutionary algorithms (MOEAs) to determine its relative performance on test and real world problems.
-Many evolutionary approaches e.g., the nondominated sorting genetic algorithm (NSGA) family of algorithms, have been adapted to solve multiobjective optimization problems [REF].
+Many evolutionary approaches e.g., the nondominated sorting genetic algorithm (NSGA) family of algorithms, have been adapted to solve multiobjective optimization problems \cite{DEB2002,HUBAND2006}.
 It is unclear if JuPOETs will perform as well as these other approaches;
 one potential advantage that JuPOETs may have is the local refinement step which temporarily reduces the problem to a single objective formulation.
-Previously, this hybrid approach led to better convergence on a proof-of-concept signal transduction model [REF].
+Previously, this hybrid approach led to better convergence on a proof-of-concept signal transduction model \cite{Song:2010aa}.
 Next, JuPOETs should take advantage of the native parallel execution capabilities of Julia to accelerate the evaluation of the objective functions.
 For many real world parameter estimation problems, the bulk of the execution time is spent evaluating the objective functions.
-One strategy to improve performance could be to optimize surrogates [REF], while another would be parallel execution of the objective functions.
-Currently, JuPOETs serially evaluate the components of the objective function vector.
-However, because of the flexible function pointer architecture of JuPOETs, a shift to parallel evaluation of the objective functions requires changes to only the user defined objective function and not the main run loop.
-Thus, parallel evaluation of objective functions could be easily implemented using a variety of techniques without changing to JuPOETs.
-
-\section*{Conclusions}
+One strategy to improve performance could be to optimize surrogates \cite{SURROGATES}, while another would be parallel execution of the objective functions.
+Currently, JuPOETs serially evaluates the objective function vector.
+However, because of the flexible function pointer architecture of JuPOETs, a shift to parallel evaluation of the objectives requires changes to the user
+defined objective function and not the main run loop. Thus, parallel evaluation of objective functions could be easily implemented
+using a variety of techniques without changing to JuPOETs.
 

### 2016-05-25 Paper_v1.tex
Name update / NSFGRF correction

@@
    email={dmb457@cornell.edu}
-]{\inits{DB}\fnm{David} \snm{Bassen}}
+]{\inits{DB}\fnm{David M} \snm{Bassen}}
 \author[
@@
 This study was supported by an award from the National Science Foundation (NSF CBET-0955172) and the National Institutes of Health (NIH HL110328) to J.B, and by
-a National Science Foundation Graduate Research Fellowship (NSGRF) to D.B.
+a National Science Foundation Graduate Research Fellowship (NSFGRF) to D.B.
 

### 2016-05-27 Paper_v1.tex
grantnum

@@
 This study was supported by an award from the National Science Foundation (NSF CBET-0955172) and the National Institutes of Health (NIH HL110328) to J.B, and by
-a National Science Foundation Graduate Research Fellowship (NSFGRF) to D.B.
+a National Science Foundation Graduate Research Fellowship (DGE-1144153) to D.B.
 

### 2016-05-27 Paper_v1.tex
DB_proofs

@@
 \begin{abstract} % abstract
-Ensemble modeling is a well established approach for obtaining robust predictions and course graining population behavior in large deterministic mathematical models.
+Ensemble modeling is a well established approach for obtaining robust predictions and course grained population behavior in large deterministic mathematical models.
 In this study, we present the Pareto Optimal Ensemble Technique in the Julia programming language (JuPOETs).
 JuPOETs integrates simulated annealing with Pareto optimality to estimate an ensemble of parameter sets on or near the optimal tradeoff surface between competing training objectives.
-We demonstrated JuPOETs on a suite of multiobjective test problems, with both bounds and problem constraints as well as a proof-of-concept biochemical model identification problem.
+We demonstrated JuPOETs on a suite of multiobjective problems, including test functions with parameter bounds and system constraints as well as identification of a proof-of-concept biochemical model.
 JuPOETs identified optimal or near optimal solutions approximately six-fold faster than a corresponding implementation in Octave. 
@@
 \section*{Background}
-Ensemble modeling is a well established approach for obtaining robust predictions and course graining population behavior in large deterministic mathematical models.
+Ensemble modeling is a well established approach for obtaining robust predictions and course grained population behavior in large deterministic mathematical models.
 It is often not possible to uniquely identify all the parameters in biochemical models, even when given extensive training data \cite{Gadkar:2005ad}.
@@
 Identification of biochemical models with hundreds or even thousands of states and parameters may not be tractable as a single objective optimization problem.
-Large models require significant training data perhaps taken from diverse sources, for example different laboratories or cell-lines.
-These data are often heterogenous, and contain intrinsic conflicts which complicate parameter estimation.
+Further, large models require significant training data perhaps taken from diverse sources, for example different laboratories or cell-lines.
+These data are often heterogenous, and contain intrinsic conflicts that complicate parameter estimation.
 Parameter ensembles which optimally balance tradeoffs between submodels or conflicts in training data can
-lead to robust model performance. One class of ensemble generation technique that naturally balances conflicting training data is multiobjective optimization.
+lead to robust model performance. Multiobjective optimization is a class of ensemble generation techniques that naturally balances conflicting training data.
 Previously, we developed the Pareto Optimal Ensemble Technique (POETs) algorithm to address the challenge of competing or conflicting objectives.
@@
 Julia also offers a sophisticated compiler, distributed parallel execution, numerical accuracy, and an extensive function library.
+Further, the architecture of JuPOETs takes advantage of the first-class function type in Julia allowing user definable behavior for all key aspects of the algorithm, including objective functions, custom diversity generation logic, linear/non-linear parameter constraints (and parameter bounds constraints) as well as custom cooling schedules.
+Julia's ability to naturally call other languages such as Python or C also allows JuPOETS to be used with models implemented in a variety of languages across many platforms.
 Additionally, Julia offers a built-in package manager which is directly integrated with GitHub,
 a popular web-based Git repository hosting service offering distributed revision control and source code management.
-Next, because Julia can natively call other languages such as Python or C, JuPOETs can be used with models implemented in a variety of languages on many platforms.
-Lastly, the architecture of JuPOETs takes advantage of the first-class function type in Julia allowing user definable behavior for all key aspects of the algorithm, including objective functions, custom diversity generation logic, linear/non-linear parameter constraints (and parameter bounds constraints) as well as custom cooling schedules.
-Thus, JuPOETs can be adapted to many problem types, including mixed binary and continuous variable types, bilevel problems and constrained problems without changing the base algorithm (which was not true of the previous POETs implementation).
+Thus, JuPOETs can be adapted to many problem types, including mixed binary and continuous variable types, bilevel problems and constrained problems without altering the base algorithm, as was required in the previous POETs implementation.
 
@@
 Previously, this hybrid approach led to better convergence on a proof-of-concept signal transduction model \cite{Song:2010aa}.
-Next, JuPOETs should take advantage of the native parallel execution capabilities of Julia to accelerate the evaluation of the objective functions.
 For many real world parameter estimation problems, the bulk of the execution time is spent evaluating the objective functions.
@@
 defined objective function and not the main run loop. Thus, parallel evaluation of objective functions could be easily implemented
-using a variety of techniques without changing to JuPOETs.
+using a variety of techniques without alterating JuPOETs. Finally, JuPOETs has the potential for widespread use due to the high level syntax and distribution tools native to Julia.
 

### 2016-05-29 Paper_v1.tex
Updated with MVs comments

@@
 These data are often heterogenous, and contain intrinsic conflicts that complicate parameter estimation.
-Parameter ensembles which optimally balance tradeoffs between submodels or conflicts in training data can
-lead to robust model performance. Multiobjective optimization is a class of ensemble generation techniques that naturally balances conflicting training data.
+Parameter ensembles which optimally balance tradeoffs between submodels and conflicts in training data can
+lead to robust model performance. Multiobjective optimization is an ensemble generation technique that naturally balances conflicting training data.
 Previously, we developed the Pareto Optimal Ensemble Technique (POETs) algorithm to address the challenge of competing or conflicting objectives.
@@
 a parameter set as an input and returns a $\mathcal{K}\times{1}$ objective vector. Candidate parameter sets are generated by the user supplied \texttt{neighbor} function.
-The error vector associated with $\mathbf{p}_{i+1}$ is ranked using the builtin Pareto \texttt{rank} function, by comparing the current error at iteration $i$+1 to the error archive $\mathcal{O}_{i}$.
+The error vector associated with $\mathbf{p}_{i+1}$ is ranked using the builtin Pareto \texttt{rank} function, by comparing the current error at iteration $i$+1 to the error archive $\mathcal{O}_{i}$ (all error vectors up to iteration $i$-1 that meeting a ranking criteria).
 Pareto rank is a measure of distance from the trade-off surface;
@@
 Sets with increasing non-zero rank are progressively further away from the optimal trade-off surface.
-Thus, a parameter set with a rank$=0$ is \emph{better} in a trade-off sense than rank$>0$.
+Thus, a parameter set with a rank $=0$ is \emph{better} in a trade-off sense than rank $>0$.
 We implemented the Fonseca and Fleming ranking scheme \cite{RANKING} in the builtin \texttt{rank} function:
@@
 ensuring that we explore parameter sets along the Pareto surface.
-Occasionally (depending upon $T$) a parameter set with a high Pareto rank was accepted by the SA allowing a more diverse search of the parameter space.
+Occasionally, (depending upon $T$) a parameter set with a high Pareto rank was accepted by the SA allowing a more diverse search of the parameter space.
 However, as $T$ is reduced, the probability of accepting a high-rank set occurring decreases.
 Parameter sets could also be accepted by the SA but $\emph{not}$ permanently archived in $\mathcal{S}_{i}$.
-Only parameter sets with rank less than or equal to threshold (rank$\leq$4 by default) were included in $\mathcal{S}_{i}$, where the archive was re-ranked and filtered after
+Only parameter sets with rank less than or equal to threshold (rank $\leq$4 by default) were included in $\mathcal{S}_{i}$, where the archive was re-ranked and filtered after
 every new parameter set was accepted.
@@
 Currently, JuPOETs serially evaluates the objective function vector.
-However, because of the flexible function pointer architecture of JuPOETs, a shift to parallel evaluation of the objectives requires changes to the user
-defined objective function and not the main run loop. Thus, parallel evaluation of objective functions could be easily implemented
-using a variety of techniques without alterating JuPOETs. Finally, JuPOETs has the potential for widespread use due to the high level syntax and distribution tools native to Julia.
+However, parallel evaluation of the objective functions could be easily implemented using a variety of techniques without changing the main run loop of JuPOETs.
+Because of the flexible function pointer architecture of JuPOETs, the only changes required are in the user defined objective function.
+Taken together, JuPOETs has demonstrated improved flexibility, and performance over POETs in parameter identification and ensemble generation for multiple objectives.
+JuPOETs has the potential for widespread use due to the flexibility of the implementation, and the high level syntax and distribution tools native to Julia.
+
+%However, because of the flexible function pointer architecture of JuPOETs, a shift to parallel evaluation of the objectives requires changes to the user
+%defined objective function and not the main run loop. Thus, parallel evaluation of objective functions could be easily implemented
+%using a variety of techniques without alterating JuPOETs.
 

### 2016-05-29 Paper_v1.tex
Almost submitted

@@
 including mutant data \cite{Khodayari:2014aa}.
-We and others have used ensemble approaches, generated using optimization and sampling approaches, to robustly simulate a wide variety of signal transduction processes \citep{Luan:2007aa,Song:2009aa,Tasseff:2010aa,Tasseff:2011aa,Nayak:2011aa}, neutrophil trafficking in sepsis \cite{Song:2012aa},
+We and others have used ensemble approaches, generated using both sampling and optimization techniques, to robustly simulate a wide variety of signal transduction processes \citep{Luan:2007aa,Song:2009aa,Tasseff:2010aa,Tasseff:2011aa,Nayak:2011aa}, neutrophil trafficking in sepsis \cite{Song:2012aa},
 patient specific coagulation behavior \cite{Luan:2010aa} and to capture cell to cell variation \cite{Lequieu:2011aa}.
@@
 JuPOETs can be installed using the Julia package manager from the JuPOETs repository at https://github.com/varnerlab/POETs.jl.
+Sample code is included in the \texttt{sample/biochemical} subdirectory of the JuPOETs repository to help users get started using JuPOETs in their projects.
 
@@
 
-  \Input{User specified neighbor, objective, acceptance and cooling functions. Initial parameter guess ($\mathcal{P}\times~1$)}
-  \Output{Rank archive $\mathcal{R}$, solution archive $\mathcal{S}$ and objective archive $\mathcal{O}$}
+  \Input{User specified neighbor, objective, acceptance and cooling functions. Initial parameter guess ($\mathcal{D}\times~1$)}
+  \Output{Rank archive $\mathcal{R}$, parameter solution archive $\mathcal{S}$ and objective archive $\mathcal{O}$}
 
@@
 The synthetic data was generated by randomly varying these parameters.
-JuPOETs produced an ensemble of parameters that captured the mean of the measured data sets for extracellular metabolites and cellmass
-(Fig. \ref{fig:fig-biochemical}A and B). The 95\% confidence estimate produced by the ensemble was consistent with the mean of the measured data, despite having significant
+JuPOETs produced an ensemble of approximately $\dim{\mathcal{S}}\simeq$ 13,000 parameters that captured the mean of the measured data sets for extracellular metabolites and cellmass
+(Fig. \ref{fig:fig-biochemical}A and B). JuPOETs minimized the difference between the simulated and measured values for A$_{e}$, B$_{e}$, C$_{e}$ and cellmass,
+where the residual for each data set was treated as a single objective (leading to four objectives).
+The 95\% confidence estimate produced by the ensemble was consistent with the mean of the measured data, despite having significant
 uncertainty in the training data.
@@
 \ref{fig:fig-biochemical}C).
-Thus, JuPOETs produced an ensemble of parameters that gave the mean of the training data for conflicting data sets.
+Multiple trade-off fronts were visible in the objective plots, for example between data set 3 (O$_{3}$) and data set 2 (O$_{2}$).
+Thus, without a multiobjective approach, it would be challenging to capture these data sets as fitting one leads to decreased performance on the other.
+However, the ensemble contained parameter sets that described each data set independently (Fig. \ref{fig:fig-experiment-variation}).
+Thus, JuPOETs produced an ensemble of parameters that gave the mean of the training data for conflicting data sets,
+while simultaneously estimating parameter sets that performed well on each individual objective function.
 
@@
   \caption{Proof of concept biochemical network study.
-  Inset: Prototypical biochemical network with six metabolites and seven reactions modeled using the hybrid cybernetic approach (HCM).
+  Inset right: Prototypical biochemical network with six metabolites and seven reactions modeled using the hybrid cybernetic approach (HCM).
   Intracellular cellmass precursors $A,B$, and $C$ are balanced (no accumulation) while the extracellular metabolites $A_{e},B_{e}$, and $C_{e}$ are dynamic.
   The oval denotes the cell boundary, $q_{j}$ is the $j$th flux across the boundary, and $v_{k}$ denotes the $k$th intracellular flux.
-  Four data sets (each with $A_{e},B_{e}$,$C_{e}$ and cellmass measurements) were generated by varying the kinetic constant for each biochemical modes.
-  Each data set was a single objective.
+  Four data sets (each with $A_{e},B_{e}$,$C_{e}$ and cellmass measurements) were generated by varying the kinetic constants for each biochemical mode.
+  Each data set was a single objective in the JuPOETs procedure.
   A: Ensemble simulation of extracellular substrate $A_{e}$ and cellmass versus time.
@@
    }\label{fig:fig-biochemical}
+\end{figure}
+
+\begin{figure}[h]
+  \caption{Experiment to experiment variation captured by the ensemble.
+  Cellmass measurements (points) versus time for experiment 2 and 3 were compared with ensemble simulations.
+  The full ensemble was sorted by simultaneously selecting the top 25\% of solutions for each objective with rank $\leq$~1.
+  The best fit solution for each objective (line) $\pm$ 1-standard deviation (gray region) for experiment 2 and 3 brackets the training data despite significant differences the training values between the two data sets.}\label{fig:fig-experiment-variation}
 \end{figure}

### 2016-05-29 biochemical/plot_indivdual_experiment.jl

@@ -1,0 +1,99 @@
+# Script to extract parameter solutions from the ensemble that fit individual experiments -
+
+include("SolveBalances.jl")
+using PyPlot
+using PyCall
+@pyimport numpy as np
+
+# What experiment do we want to see?
+experiment_index = 3
+MEASURED_ARRAY = readdlm("./data/MEASUREMENT_SET_3.dat")
+
+# Load the ensemble files from disk -
+pc_array_full = readdlm("./data/pc_array.dat")
+ec_array_full = readdlm("./data/ec_array.dat")
+rank_array = readdlm("./data/rank_array.dat")
+
+# Select the desired rank -
+idx_rank = find(rank_array .<= 1.0)
+
+# Remove solutions that are above the rank threshold -
+pc_array = pc_array_full[:,idx_rank]
+ec_array = ec_array_full[:,idx_rank]
+
+# Sort the remaining solutions for experiment_index from lowest to highest error -
+idx_sort = sortperm(vec(ec_array[experiment_index,:]))
+
+
+# Setup time scale -
+tStart = 0.0;
+tStop = 100.0;
+tStep = 0.1;
+number_of_timesteps = 400
+time_experimental = linspace(tStart,tStop,number_of_timesteps)
+
+# initialize data array -
+data_array = zeros(length(time_experimental),1)
+
+# Run the simulation for this rank -
+number_of_samples_max = floor(0.25*length(idx_sort))
+number_of_samples = 1
+error_index_value = 1
+while (number_of_samples<number_of_samples_max)
+
+  # Grab the parameter set from the cache -
+  parameter_array = pc_array[:,idx_sort[error_index_value]]
+
+  # Run the model -
+  (t,x) = SolveBalances(tStart,tStop,tStep,parameter_array)
+
+  # Need to interpolate the simulation onto the experimental time scale -
+  AI = np.interp(time_experimental[:],t,x[:,7])
+  BI = np.interp(time_experimental[:],t,x[:,8])
+  CI = np.interp(time_experimental[:],t,x[:,9])
+  XI = np.interp(time_experimental[:],t,x[:,10])
+
+  # grab -
+  data_array = [data_array XI]
+
+  # update the sample count -
+  number_of_samples = number_of_samples + 1
+
+  # Update the index for the next solution -
[…]

### 2016-05-29 Paper_BioArXiv_Submitted_v1.tex
Formatting for BioArXiv submission

--- 2016-05-29 Paper_v1.tex
@@
-%% BioMed_Central_Tex_Template_v1.06
-%%                                      %
-%  bmc_article.tex            ver: 1.06 %
-%                                       %
-
-%%IMPORTANT: do not delete the first line of this template
-%%It must be present to enable the BMC Submission system to
-%%recognise this template!!
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                     %%
-%%  LaTeX template for BioMed Central  %%
-%%     journal article submissions     %%
-%%                                     %%
-%%          <8 June 2012>              %%
-%%                                     %%
-%%                                     %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                                                 %%
-%% For instructions on how to fill out this Tex template           %%
-%% document please refer to Readme.html and the instructions for   %%
-%% authors page on the biomed central website                      %%
-%% http://www.biomedcentral.com/info/authors/                      %%
-%%                                                                 %%
-%% Please do not use \input{...} to include other tex files.       %%
-%% Submit your LaTeX manuscript as one .tex document.              %%
-%%                                                                 %%
-%% All additional figures and files should be attached             %%
-%% separately and not embedded in the \TeX\ document itself.       %%
-%%                                                                 %%
-%% BioMed Central currently use the MikTex distribution of         %%
-%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
-%% http://www.miktex.org                                           %%
-%%                                                                 %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%% additional documentclass options:
-%  [doublespacing]
-%  [linenumbers]   - put the line numbers on margins
-
-%%% loading packages, author definitions
-
-%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
-\documentclass{bmcart}
-
-%%% Load packages
-\usepackage{amsthm,amsmath}
+\documentclass[12pt]{article}
+% Load packages
+\usepackage{url}  % Formatting web addresses
+\usepackage{ifthen}  % Conditional
+\usepackage{multicol}   %Columns
+\usepackage[utf8]{inputenc} %unicode support
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{epsfig}
+\usepackage{epstopdf}
 \usepackage{graphicx}
 \usepackage[margin=0.1pt,font=footnotesize,labelfont=bf]{caption}
-
-% Imports required for p-code of JuPOETs
-\usepackage[linesnumbered]{algorithm2e}
-\usepackage{courier}
-%\usepackage{algorithm}
-\RequirePackage{natbib}
-%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
-\RequirePackage{hyperref}
-\usepackage[utf8]{inputenc} %unicode support
+\usepackage{setspace}
+%\usepackage{longtable}
+\usepackage{colortbl}
+%\usepackage{palatino,lettrine}
+%\usepackage{times}
 %\usepackage[applemac]{inputenc} %applemac support if unicode package fails
 %\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
+\usepackage[wide]{sidecap}
+%\usepackage[authoryear,round,comma,sort&compress]{natbib}
+\usepackage[square,sort,comma,numbers,sort&compress]{natbib}
+%\usepackage[authoryear,round]{natbib}
+\usepackage{supertabular}
+\usepackage{simplemargins}
+\usepackage{fullpage}
+\usepackage{comment}
+\usepackage{lineno}
+%\usepackage{chicago}
+\usepackage{textcomp}
+\usepackage{multirow}
+\usepackage{amsmath}
+\usepackage[linesnumbered,lined,boxed,commentsnumbered]{algorithm2e}
+\DeclareMathOperator*{\argmin}{\arg\!\min}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                             %%
-%%  If you wish to display your graphics for   %%
-%%  your own use using includegraphic or       %%
-%%  includegraphics, then comment out the      %%
-%%  following two lines of code.               %%
-%%  NB: These line *must* be included when     %%
-%%  submitting to BMC.                         %%
-%%  All figure files must be submitted as      %%
-%%  separate graphics through the BMC          %%
-%%  submission process, not included in the    %%
-%%  submitted article.                         %%
-%%                                             %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\usepackage{algorithm2e}
+\usepackage{algpseudocode}
+%\usepackage[space]{cite}
+\urlstyle{rm}
 
-\def\includegraphic{}
-\def\includegraphics{}
+%\textwidth = 6.50 in
+%\textheight = 9.5 in
+%\oddsidemargin =  0.0 in
+%\evensidemargin = 0.0 in
+%\topmargin = -0.50 in
+%\headheight = 0.0 in
+%\headsep = 0.25 in
+%\parskip = 0.15in
+%\linespread{1.75}
+\doublespace
 
-%%% Put your definitions there:
-\startlocaldefs
-\endlocaldefs
+%\bibliographystyle{chicago}
+\bibliographystyle{plos2009}
 
-%%% Begin ...
+\makeatletter
+\renewcommand\subsection{\@startsection
+	{subsection}{2}{0mm}
+	{-0.05in}
+	{-0.5\baselineskip}
+	{\normalfont\normalsize\bfseries}}
+\renewcommand\subsubsection{\@startsection
+	{subsubsection}{2}{0mm}
+	{-0.05in}
+	{-0.5\baselineskip}
+	{\normalfont\normalsize\itshape}}
+\renewcommand\section{\@startsection
+	{subsection}{2}{0mm}
+	{-0.2in}
+	{0.05\baselineskip}
+	{\normalfont\large\bfseries}}
+\renewcommand\paragraph{\@startsection
+	{paragraph}{2}{0mm}
+	{-0.05in}
+	{-0.5\baselineskip}
+	{\normalfont\normalsize\itshape}}
+\makeatother
+
+%Review style settings
+%\newenvironment{bmcformat}{\begin{raggedright}\baselineskip20pt\sloppy\setboolean{publ}{false}}{\end{raggedright}\baselineskip20pt\sloppy}
+
+%Publication style settings
+
+% Single space'd bib -
+\setlength\bibsep{0pt}
+
+\renewcommand{\rmdefault}{phv}\renewcommand{\sfdefault}{phv}
+\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
+
+% Change the number format in the ref list -
+\renewcommand{\bibnumfmt}[1]{#1.}
+
+% Change Figure to Fig.
+\renewcommand{\figurename}{Fig.}
+
+% Begin ...
 \begin{document}
-
-%%% Start of article front matter
-\begin{frontmatter}
-
-\begin{fmbox}
-\dochead{Software}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Enter the title of your article here     %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\title{JuPOETs: A Constrained Multiobjective Optimization Approach to Estimate Biochemical Model Ensembles in the Julia Programming Language}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Enter the authors here                   %%
-%%                                          %%
-%% Specify information, if available,       %%
-%% in the form:                             %%
-%%   <key>={<id1>,<id2>}                    %%
-%%   <key>=                                 %%
-%% Comment or delete the keys which are     %%
-%% not used. Repeat \author command as much %%
-%% as required.                             %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\author[
-   addressref={aff2},
-   email={dmb457@cornell.edu}
-]{\inits{DB}\fnm{David M} \snm{Bassen}}
-\author[
-  addressref={aff1}
-]{\inits{MV}\fnm{Michael} \snm{Vilkhovoy}}
-\author[
-  addressref={aff1}
-]{\inits{MM}\fnm{Mason} \snm{Minot}}
-\author[
-   addressref={aff2},
-   email={jtb47@cornell.edu}
-]{\inits{JB}\fnm{Jonathan T} \snm{Butcher}}
-\author[
-   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
-   corref={aff1},                       % id of corresponding address, if any
-   %noteref={n1},                        % id's of article notes, if any
-   email={jdv27@cornell.edu}   % email address
-]{\inits{JV}\fnm{Jeffrey D} \snm{Varner}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Enter the authors' addresses here        %%
-%%                                          %%
-%% Repeat \address commands as much as      %%
-%% required.                                %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\address[id=aff1]{%                           % unique id
-  \orgname{Department of Chemical and Biomolecular Engineering}, % university, etc
-  \street{Cornell University},                     %
-  \postcode{14853}                              % post or zip code
-  \city{Ithaca NY},                             % city
-  \cny{USA}                                    % country
-}
-\address[id=aff2]{%
-  \orgname{Department of Biomedical Engineering},
-  \street{Cornell University},                     %
-  \postcode{14853}                              % post or zip code
-  \city{Ithaca NY},                             % city
-  \cny{USA}
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Enter short notes here                   %%
-%%                                          %%
-%% Short notes will be after addresses      %%
-%% on first page.                           %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{artnotes}
-%\note{Sample of title note}     % note to the article
-%\note[id=n1]{Equal contributor} % note, connected to author
-\end{artnotes}
-
-\end{fmbox}% comment this for two column layout
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% The Abstract begins here                 %%
-%%                                          %%
-%% Please refer to the Instructions for     %%
-%% authors on http://www.biomedcentral.com  %%
-%% and include the section headings         %%
-%% accordingly for your article type.       %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{abstractbox}
-
-\begin{abstract} % abstract
-Ensemble modeling is a well established approach for obtaining robust predictions and course grained population behavior in large deterministic mathematical models.
-In this study, we present the Pareto Optimal Ensemble Technique in the Julia programming language (JuPOETs).
-JuPOETs integrates simulated annealing with Pareto optimality to estimate an ensemble of parameter sets on or near the optimal tradeoff surface between competing training objectives.
-We demonstrated JuPOETs on a suite of multiobjective problems, including test functions with parameter bounds and system constraints as well as identification of a proof-of-concept biochemical model.
-JuPOETs identified optimal or near optimal solutions approximately six-fold faster than a corresponding implementation in Octave.
-JuPOETs can be installed using the Julia package manager from the JuPOETs GitHub repository at https://github.com/varnerlab/POETs.jl.
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% The keywords begin here                  %%
-%%                                          %%
-%% Put each keyword in separate \kwd{}.     %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{keyword}
-\kwd{Ensemble~modeling}
-\kwd{Multiobjective~optimization}
-\kwd{Julia}
-\end{keyword}
-
-% MSC classifications codes, if any
-%\begin{keyword}[class=AMS]
-%\kwd[Primary ]{}
-%\kwd{}
-%\kwd[; secondary ]{}
-%\end{keyword}
-
-\end{abstractbox}
-%
-%\end{fmbox}% uncomment this for twcolumn layout
-
-\end{frontmatter}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% The Main Body begins here                %%
-%%                                          %%
-%% Please refer to the instructions for     %%
-%% authors on:                              %%
-%% http://www.biomedcentral.com/info/authors%%
-%% and include the section headings         %%
-%% accordingly for your article type.       %%
-%%                                          %%
-%% See the Results and Discussion section   %%
-%% for details on how to create sub-sections%%
-%%                                          %%
-%% use \cite{...} to cite references        %%
-%%  \cite{koon} and                         %%
-%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
-%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
-% <put your article body there>
-
-%%%%%%%%%%%%%%%%
-%% Background %%
-%%
-
-%Sethna and coworkers later showed that model performance is often controlled by only a few parameter combinations,
-%a characteristic seemingly universal to multi-parameter models referred to as \textit{sloppiness} \citep{Machta:2013aa}.
-%n the other hand, large models can also be constructed from many smaller modules, each of which is independently identified.
-%For example, Diamond and coworkers developed a model of calcium signaling in human platelets composed of many
-%separate submodels identified using different types of training data \cite{Purvis:2008}.
 
@@
 source code management (SCM) system. Thus, upgrades to the approach containing new features, or bug fixes were not centrally managed.
+
+\clearpage
 
@@
 \section*{Results and Discussion}
-
 JuPOETs identified optimal or nearly optimal solutions significantly faster than Octave-POETs for a suite of multiobjective test problems (Table \ref{table:objective-table}).
@@
 \section*{Conclusions}
-
 JuPOETs is a significant advance over the previous POETs implementation.
@@
 
-%However, because of the flexible function pointer architecture of JuPOETs, a shift to parallel evaluation of the objectives requires changes to the user
-%defined objective function and not the main run loop. Thus, parallel evaluation of objective functions could be easily implemented
-%using a variety of techniques without alterating JuPOETs.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Backmatter begins here                   %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{backmatter}
+\clearpage
 
@@
 
-%\section*{Acknowledgements}
-%  Text for this section \ldots
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                  The Bibliography                       %%
-%%                                                         %%
-%%  Bmc_mathpys.bst  will be used to                       %%
-%%  create a .BBL file for submission.                     %%
-%%  After submission of the .TEX file,                     %%
-%%  you will be prompted to submit your .BBL file.         %%
-%%                                                         %%
-%%                                                         %%
-%%  Note that the displayed Bibliography will not          %%
-%%  necessarily be rendered by Latex exactly as specified  %%
-%%  in the online Instructions for Authors.                %%
-%%                                                         %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% if your bibliography is in bibtex format, use those commands:
-\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
-\bibliography{References_v1}      % Bibliography file (usually '*.bib' )
-% for author-year bibliography (bmc-mathphys or spbasic)
-% a) write to bib file (bmc-mathphys only)
-% @settings{label, options="nameyear"}
-% b) uncomment next line
-%\nocite{label}
-
-% or include bibliography directly:
-% \begin{thebibliography}
-% \bibitem{b1}
-% \end{thebibliography}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                               %%
-%% Figures                       %%
-%%                               %%
-%% NB: this is for captions and  %%
-%% Titles. All graphics must be  %%
-%% submitted separately and NOT  %%
-%% included in the Tex document  %%
-%%                               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% Do not use \listoffigures as most will included as separate files
+\bibliography{References_v1}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                               %%
-%% Tables                        %%
-%%                               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\clearpage
 
-%% Use of \listoftables is discouraged.
-%%
-\clearpage
-\section*{Tables}
+% Supplemental figures -
+% Set the S-
+\renewcommand\thefigure{S\arabic{figure}}
+\renewcommand\thetable{T\arabic{table}}
+\renewcommand\thepage{S-\arabic{page}}
+\renewcommand\theequation{S\arabic{equation}}
 
-\begin{table}
-  \caption{Multi-objective optimization test problems. We tested the JuPOETs implementation on three two-dimensional test problems, with one-, two- and three-dimensional parameter vectors.
-  Each problem had parameter bounds constraints, however, on the Binh and Korn function had additional non-linear problem constraints. For the Fonesca and Fleming problem, N = 3.}
-  \label{table:objective-table}
-\end{table}
+% Reset the counters -
+\setcounter{equation}{0}
+\setcounter{table}{0}
+\setcounter{figure}{0}
+\setcounter{page}{1}
 
-% \begin{table}[h!]
-% \caption{Sample table title. This is where the description of the table should go.}
-%       \begin{tabular}{cccc}
-%         \hline
-%            & B1  &B2   & B3\\ \hline
-%         A1 & 0.1 & 0.2 & 0.3\\
-%         A2 & ... & ..  & .\\
-%         A3 & ..  & .   & .\\ \hline
-%       \end{tabular}
-% \end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                               %%
-%% Additional Files              %%
-%%                               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \section*{Additional Files}
-%   \subsection*{Additional file 1 --- Sample additional file title}
-%     Additional file descriptions text (including details of how to
-%     view the file, if it is in a non-standard format or the file extension).  This might
-%     refer to a multi-page table or a figure.
-%
-%   \subsection*{Additional file 2 --- Sample additional file title}
-%     Additional file descriptions text.
-
-\end{backmatter}
 \end{document}

### 2016-05-29 Paper_BioArXiv_Submitted_v1.tex
BioArXiv submission with figures

@@
 \usepackage{amsmath}
-\usepackage[linesnumbered,lined,boxed,commentsnumbered]{algorithm2e}
+\usepackage[linesnumbered,lined,commentsnumbered]{algorithm2e}
 \DeclareMathOperator*{\argmin}{\arg\!\min}
 
-\usepackage{algorithm2e}
 \usepackage{algpseudocode}
@@
 \begin{document}
+
+\begin{titlepage}
+{\par\centering\textbf{\Large {JuPOETs: A Constrained Multiobjective Optimization Approach to Estimate Biochemical Model Ensembles in the Julia Programming Language}}}
+\vspace{0.05in}
+{\par \centering \large{David Bassen$^{1}$, Michael Vilkhovoy, Mason Minot, Jonathan T Butcher$^{1}$ and Jeffrey D. Varner$^{*}$}}
+\vspace{0.10in}
+{\par \centering {School of Chemical and Biomolecular Engineering and}}
+{\par \centering {$^{1}$School of Biomedical Engineering, Cornell University, Ithaca NY 14853}}
+\vspace{0.1in}
+{\par \centering \textbf{Running Title:}~Multiobjective model ensemble generation}
+\vspace{0.1in}
+{\par \centering \textbf{To be submitted:}~\emph{BMC~Systems~Biology}}
+\vspace{0.1in}
+{\par \centering $^{*}$Corresponding author:}
+{\par \centering Jeffrey D. Varner,}
+{\par \centering Professor, School of Chemical and Biomolecular Engineering,}
+{\par \centering 244 Olin Hall, Cornell University, Ithaca NY, 14853}
+{\par \centering Email: jdv27@cornell.edu}
+{\par \centering Phone: (607) 255 - 4258}
+{\par \centering Fax: (607) 255 - 9166}
+\end{titlepage}
+\date{}
+\thispagestyle{empty}
+\pagebreak
+
+\section*{Abstract} % abstract
+Ensemble modeling is a well established approach for obtaining robust predictions and course grained population behavior in large deterministic mathematical models.
+In this study, we present the Pareto Optimal Ensemble Technique in the Julia programming language (JuPOETs).
+JuPOETs integrates simulated annealing with Pareto optimality to estimate an ensemble of parameter sets on or near the optimal tradeoff surface between competing training objectives.
+We demonstrated JuPOETs on a suite of multiobjective problems, including test functions with parameter bounds and system constraints as well as identification of a proof-of-concept biochemical model.
+JuPOETs identified optimal or near optimal solutions approximately six-fold faster than a corresponding implementation in Octave.
+JuPOETs can be installed using the Julia package manager from the JuPOETs GitHub repository at https://github.com/varnerlab/POETs.jl.
+
+\clearpage
+
+% Uncomment in production -
+\linenumbers
 
@@
 
-\begin{algorithm}[h!]
-
-  \SetKwFunction{objective}{objective}
-  \SetKwFunction{neighbor}{neighbor}
-  \SetKwFunction{acceptance}{acceptance}
-  \SetKwFunction{rank}{rank}
-
-  \SetKwInOut{Input}{input}
-  \SetKwInOut{Output}{Output}
-
-  \Input{User specified neighbor, objective, acceptance and cooling functions. Initial parameter guess ($\mathcal{D}\times~1$)}
-  \Output{Rank archive $\mathcal{R}$, parameter solution archive $\mathcal{S}$ and objective archive $\mathcal{O}$}
-
-  \BlankLine
-  \BlankLine
-  initialize: $\mathcal{R}$, $\mathcal{S}$ and $\mathcal{O}$ using initial guess\;
-  initialize: T $\leftarrow$1.0\;
-  initialize: $T_{min}\leftarrow$1/10000\;
-  initialize: Maximum number of steps per temperature $\mathcal{I}$\;
-
-  \BlankLine
-  \BlankLine
- \While{$T>T_{min}$}{
-    i $\leftarrow$  1\;
-    \While{i$<\mathcal{I}$}{
-      \BlankLine
-      \BlankLine
-      \tcp{Generate a new parameter solution using user neighbor function}
-      $\mathbf{p}_{i+1}$ $\leftarrow$ user-function::\neighbor{$\mathbf{p}^{*}$}\;
-      \BlankLine
-      \BlankLine
-      \tcp{Evaluate $\mathbf{p}_{i+1}$ using user objective function}
-      $\mathbf{o}_{i+1}$ $\leftarrow$ user-function::\objective{$\mathbf{p}_{i+1}$}\;
-      \BlankLine
-      \BlankLine
-      Add $\mathbf{p}_{i+1}$ to solution archive $\mathcal{S}$\;
-      Add $\mathbf{o}_{i+1}$ to objective archive $\mathcal{O}$\;
-      \BlankLine
-      \BlankLine
-      \tcp{Calculate Pareto rank of solutions in $\mathcal{O}$ using builtin rank function}
-      $\mathcal{R}\leftarrow$ builtin-function::\rank{$\mathcal{O}$}\;
-
-      \BlankLine
-      \tcp{Accept $\mathbf{p}_{i+1}$ into the archive with user defined probability}
-      $\mathcal{P}$ $\leftarrow$ user-function::\acceptance{$\mathcal{R}$,$T$}\;
-      \eIf {$\mathcal{P}>$rand}{
-
-          \BlankLine
-          \tcp{Update the best solution with $\mathbf{p}_{i+1}$}
-          $\mathbf{p}^{*}$ $\leftarrow$ $\mathbf{p}_{i+1}$\;
-          prune $\mathcal{S}$, $\mathcal{R}$ and $\mathcal{O}$ of all solutions above a rank threshold\;
-      }{
-        Remove $\mathbf{p}_{i+1}$ from solution archive $\mathcal{S}$\;
-        Remove $\mathbf{o}_{i+1}$ from error archive $\mathcal{O}$\;
-      }
-
-      \BlankLine
-      i $\leftarrow$  i + 1\;
-    }
-
-    \BlankLine
-    \tcp{Update $T$ using the user cooling function}
-    $T$ $\leftarrow$ user-function::cooling($T$)\;
- }
-\vspace{0.05in}
-\caption{Pseudo-code for the main run-loop of JuPOETs.
-The user specifies the \texttt{neighbor}, \texttt{acceptance}, \texttt{cooling} and \texttt{objective} functions along with an initial parameter guess.
-The rank archive $\mathcal{R}$, solution archive $\mathcal{S}$ and objective archive $\mathcal{O}$ are initialized from the initial guess.
-The initial guess is perturbed in the \texttt{neighbor} function, which generates a new solution whose performance is evaluated using the user supplied \texttt{objective} function.
-The new solution and objective values are then added to the respective archives and ranked using the builtin \texttt{rank} function.
-If the new solution is accepted (based upon a probability calculated with the user supplied \texttt{acceptance} function) it is added to the solution and objective archive.
-This solution is then perturbed during the next iteration of the algorithm.
-However, if the solution is not accepted, it is removed from the archive and discarded.
-The computational temperature is adjusted using the user supplied \texttt{cooling} function after each $\mathcal{I}$ iterations.
-}\label{algo:JuPOETs-pcode}
-\end{algorithm}
-
 %(Fig. \ref{fig:fig-timing} and Fig. \ref{fig:fig-surfaces}).
@@
 
+\begin{algorithm}[h!]
+
+\footnotesize{
+  \SetKwFunction{objective}{objective}
+  \SetKwFunction{neighbor}{neighbor}
+  \SetKwFunction{acceptance}{acceptance}
+  \SetKwFunction{rank}{rank}
+
+  \SetKwInOut{Input}{input}
+  \SetKwInOut{Output}{Output}
+
+  \Input{User specified neighbor, objective, acceptance and cooling functions. Initial parameter guess ($\mathcal{D}\times~1$)}
+  \Output{Rank archive $\mathcal{R}$, parameter solution archive $\mathcal{S}$ and objective archive $\mathcal{O}$}
+
+  \BlankLine
+  \BlankLine
+  initialize: $\mathcal{R}$, $\mathcal{S}$ and $\mathcal{O}$ using initial guess\;
+  initialize: T $\leftarrow$1.0\;
+  initialize: $T_{min}\leftarrow$1/10000\;
+  initialize: Maximum number of steps per temperature $\mathcal{I}$\;
+
+  \BlankLine
+  \BlankLine
+ \While{$T>T_{min}$}{
+    i $\leftarrow$  1\;
+    \While{i$<\mathcal{I}$}{
+      \BlankLine
+      \BlankLine
+      \tcp{Generate a new parameter solution using user neighbor function}
+      $\mathbf{p}_{i+1}$ $\leftarrow$ user-function::\neighbor{$\mathbf{p}^{*}$}\;
+      \BlankLine
+      \BlankLine
+      \tcp{Evaluate $\mathbf{p}_{i+1}$ using user objective function}
+      $\mathbf{o}_{i+1}$ $\leftarrow$ user-function::\objective{$\mathbf{p}_{i+1}$}\;
+      \BlankLine
+      \BlankLine
+      Add $\mathbf{p}_{i+1}$ to solution archive $\mathcal{S}$\;
+      Add $\mathbf{o}_{i+1}$ to objective archive $\mathcal{O}$\;
+      \BlankLine
+      \BlankLine
+      \tcp{Calculate Pareto rank of solutions in $\mathcal{O}$ using builtin rank function}
+      $\mathcal{R}\leftarrow$ builtin-function::\rank{$\mathcal{O}$}\;
+
+      \BlankLine
+      \tcp{Accept $\mathbf{p}_{i+1}$ into the archive with user defined probability}
+      $\mathcal{P}$ $\leftarrow$ user-function::\acceptance{$\mathcal{R}$,$T$}\;
+      \eIf {$\mathcal{P}>$rand}{
+
+          \BlankLine
+          \tcp{Update the best solution with $\mathbf{p}_{i+1}$}
+          $\mathbf{p}^{*}$ $\leftarrow$ $\mathbf{p}_{i+1}$\;
+          prune $\mathcal{S}$, $\mathcal{R}$ and $\mathcal{O}$ of all solutions above a rank threshold\;
+      }{
+        Remove $\mathbf{p}_{i+1}$ from solution archive $\mathcal{S}$\;
+        Remove $\mathbf{o}_{i+1}$ from error archive $\mathcal{O}$\;
+      }
+
+      \BlankLine
+      i $\leftarrow$  i + 1\;
+    }
+
+    \BlankLine
+    \tcp{Update $T$ using the user cooling function}
+    $T$ $\leftarrow$ user-function::cooling($T$)\;
+ }}
+
+\vspace{0.05in}
+\caption{\footnotesize{Pseudo-code for the main run-loop of JuPOETs.
+The user specifies the \texttt{neighbor}, \texttt{acceptance}, \texttt{cooling} and \texttt{objective} functions along with an initial parameter guess.
+The rank archive $\mathcal{R}$, solution archive $\mathcal{S}$ and objective archive $\mathcal{O}$ are initialized from the initial guess.
+The initial guess is perturbed in the \texttt{neighbor} function, which generates a new solution whose performance is evaluated using the user supplied \texttt{objective} function.
+The new solution and objective values are then added to the respective archives and ranked using the builtin \texttt{rank} function.
+If the new solution is accepted (based upon a probability calculated with the user supplied \texttt{acceptance} function) it is added to the solution and objective archive.
+This solution is then perturbed during the next iteration of the algorithm.
+However, if the solution is not accepted, it is removed from the archive and discarded.
+The computational temperature is adjusted using the user supplied \texttt{cooling} function after each $\mathcal{I}$ iterations.} }\label{algo:JuPOETs-pcode}
+\end{algorithm}
+
+\clearpage
+
 \section*{Figures}
@@
 \begin{figure}[h]
+  \centering
+  \includegraphics[width=1.0\textwidth]{./figs/Figure-1-Parameter-ObjMapping.pdf}
   \caption{Schematic of multiobjective parameter mapping.
@@
 \begin{figure}[h]
+  \includegraphics[width=1.0\textwidth]{./figs/Figure-2-ObjectiveDomain-Space.png}
   \caption{The performance of JuPOETs on the multi-objective test suite.
@@
 \begin{figure}[h]
+  \includegraphics[width=1.0\textwidth]{./figs/Figure-3-PerformanceComparison.pdf}
   \caption{Representative JuPOETs solutions for problems in the multi-objective test suite.
@@
 \begin{figure}[h]
+  \includegraphics[width=1.0\textwidth]{./figs/Figure-4-BiochemicalProofOfConcept.pdf}
   \caption{Proof of concept biochemical network study.
@@
 \begin{figure}[h]
+  \includegraphics[width=1.0\textwidth]{./figs/Figure-5-ExperimentToExperimentVariation.pdf}
   \caption{Experiment to experiment variation captured by the ensemble.

### 2016-05-29 Paper_BioArXiv_Submitted_v1.tex
Updated for BioArXiv submission

@@
 \section*{Abstract} % abstract
-Ensemble modeling is a well established approach for obtaining robust predictions and course grained population behavior in large deterministic mathematical models.
-In this study, we present the Pareto Optimal Ensemble Technique in the Julia programming language (JuPOETs).
-JuPOETs integrates simulated annealing with Pareto optimality to estimate an ensemble of parameter sets on or near the optimal tradeoff surface between competing training objectives.
-We demonstrated JuPOETs on a suite of multiobjective problems, including test functions with parameter bounds and system constraints as well as identification of a proof-of-concept biochemical model.
-JuPOETs identified optimal or near optimal solutions approximately six-fold faster than a corresponding implementation in Octave.
+Ensemble modeling is a well established approach for obtaining robust predictions and for simulating course grained population behavior in deterministic mathematical models.
+In this study, we present a multiobjective based technique to estimate model ensembles, the Pareto Optimal Ensemble Technique in the Julia programming language (JuPOETs).
+JuPOETs integrates simulated annealing with Pareto optimality to estimate parameter ensembles on or near the optimal tradeoff surface between competing training objectives.
+We demonstrated JuPOETs on a suite of multiobjective problems, including test functions with parameter bounds and system constraints as well as for the identification of a proof-of-concept biochemical model with four conflicting training objectives.
+JuPOETs identified optimal or near optimal solutions approximately six-fold faster than a corresponding implementation in Octave for the suite of test functions.
+For the proof-of-concept biochemical model, JuPOETs produced an ensemble of parameters that gave both the mean of the training data for conflicting data sets,
+while simultaneously estimating parameter sets that performed well on each of the individual objective functions.
+JuPOETs can be adapted to solve many problem types, including mixed binary and continuous variable types, bilevel optimization problems
+and constrained problems without altering the base algorithm.
 JuPOETs can be installed using the Julia package manager from the JuPOETs GitHub repository at https://github.com/varnerlab/POETs.jl.
@@
   \includegraphics[width=1.0\textwidth]{./figs/Figure-5-ExperimentToExperimentVariation.pdf}
-  \caption{Experiment to experiment variation captured by the ensemble.
+  \caption{Experiment to experiment variation is captured by a single ensemble.
   Cellmass measurements (points) versus time for experiment 2 and 3 were compared with ensemble simulations.

### 2016-05-29 Paper_BioArXiv_Submitted_v1.tex
Updated the BMC version

@@
 a parameter set as an input and returns a $\mathcal{K}\times{1}$ objective vector. Candidate parameter sets are generated by the user supplied \texttt{neighbor} function.
-The error vector associated with $\mathbf{p}_{i+1}$ is ranked using the builtin Pareto \texttt{rank} function, by comparing the current error at iteration $i$+1 to the error archive $\mathcal{O}_{i}$ (all error vectors up to iteration $i$-1 that meeting a ranking criteria).
+The error vector associated with $\mathbf{p}_{i+1}$ is ranked using the builtin Pareto \texttt{rank} function, by comparing the current error at iteration $i+1$ to the error archive $\mathcal{O}_{i}$ (all error vectors up to iteration $i-1$ meeting a ranking criteria).
 Pareto rank is a measure of distance from the trade-off surface;
@@
 Thus, a parameter set with a rank $=0$ is \emph{better} in a trade-off sense than rank $>0$.
-We implemented the Fonseca and Fleming ranking scheme \cite{RANKING} in the builtin \texttt{rank} function:
+We implemented the Fonseca and Fleming ranking scheme in the builtin \texttt{rank} function \cite{RANKING}:
 \begin{equation}\label{eqn_rank}
-\texttt{rank}\left(\mathbf{p}_{i+1}\mid \mathcal{O}_{i}\right)=r
-\end{equation} where rank $r$ is the number of parameter sets that dominate (are better than) parameter set $\mathbf{p}_{i+1}$.
+\texttt{rank}\left(\mathcal{O}_{i+1}\left(\mathbf{p}_{i+1}\right)\mid \mathcal{O}_{i}\right)=r
+\end{equation} where rank $r$ is the number of parameter sets that dominate (are better than) parameter set $\mathbf{p}_{i+1}$, and $\mathcal{O}_{i+1}\left(\mathbf{p}_{i+1}\right)$
+denotes the objective vector evaluated at $\mathbf{p}_{i+1}$.
 We used the Pareto rank to inform the SA calculation.
@@
 \begin{equation}\label{eqn_costMOSA}
-\mathcal{P}(\mathbf{p}_{i+1}) \equiv \exp{\{-\texttt{rank}\left(\mathbf{p}_{i+1} \mid \mathcal{S}_{i} \right)/T\}}
+\mathcal{P}(\mathbf{p}_{i+1}) \equiv \exp{\{-\texttt{rank}\left(\mathcal{O}_{i+1}\left(\mathbf{p}_{i+1}\right) \mid \mathcal{O}_{i} \right)/T\}}
 \end{equation}
 where $T$ is the computational annealing temperature.
-As $\texttt{rank}\left(\mathbf{p}_{i+1}\mid \mathcal{O}_{i}\right)\rightarrow{0}$, the acceptance probability moves toward one,
+As $\texttt{rank}\left(\mathcal{O}_{i+1}\left(\mathbf{p}_{i+1}\right)\mid \mathcal{O}_{i}\right)\rightarrow{0}$, the acceptance probability moves toward one,
 ensuring that we explore parameter sets along the Pareto surface.
@@
   C: Trade-off plots between the four training objectives. The quantity $O_{j}$ denotes the jth training objective.
-  Each point represents a member of the parameter ensemble, where gray denotes rank 0 sets, while black denotes rank 1 sets.
+  Each point represents a member of the parameter ensemble, where black denotes a rank 0 set, while gray denotes rank 1 set.
    }\label{fig:fig-biochemical}
@@
 \clearpage
+
+\begin{table}
+	\includegraphics[width=1.0\textwidth]{./figs/Table-1-Objectives.pdf}
+	\caption{Multi-objective optimization test problems. We tested the JuPOETs implementation on three two-dimensional test problems, with one-, two- and three-dimensional parameter vectors.
+  Each problem had parameter bounds constraints, however, on the Binh and Korn function had additional non-linear problem constraints. For the Fonesca and Fleming problem, N = 3.}
+  \label{table:objective-table}
+\end{table}
 

### 2016-05-29 Paper_v1.tex
Updated the BMC version

@@
 \begin{abstract} % abstract
-Ensemble modeling is a well established approach for obtaining robust predictions and course grained population behavior in large deterministic mathematical models.
-In this study, we present the Pareto Optimal Ensemble Technique in the Julia programming language (JuPOETs).
-JuPOETs integrates simulated annealing with Pareto optimality to estimate an ensemble of parameter sets on or near the optimal tradeoff surface between competing training objectives.
-We demonstrated JuPOETs on a suite of multiobjective problems, including test functions with parameter bounds and system constraints as well as identification of a proof-of-concept biochemical model.
-JuPOETs identified optimal or near optimal solutions approximately six-fold faster than a corresponding implementation in Octave.
+Ensemble modeling is a well established approach for obtaining robust predictions and for simulating course grained population behavior in deterministic mathematical models.
+In this study, we present a multiobjective based technique to estimate model ensembles, the Pareto Optimal Ensemble Technique in the Julia programming language (JuPOETs).
+JuPOETs integrates simulated annealing with Pareto optimality to estimate parameter ensembles on or near the optimal tradeoff surface between competing training objectives.
+We demonstrated JuPOETs on a suite of multiobjective problems, including test functions with parameter bounds and system constraints as well as for the identification of a proof-of-concept biochemical model with four conflicting training objectives.
+JuPOETs identified optimal or near optimal solutions approximately six-fold faster than a corresponding implementation in Octave for the suite of test functions.
+For the proof-of-concept biochemical model, JuPOETs produced an ensemble of parameters that gave both the mean of the training data for conflicting data sets,
+while simultaneously estimating parameter sets that performed well on each of the individual objective functions.
+JuPOETs can be adapted to solve many problem types, including mixed binary and continuous variable types, bilevel optimization problems
+and constrained problems without altering the base algorithm.
 JuPOETs can be installed using the Julia package manager from the JuPOETs GitHub repository at https://github.com/varnerlab/POETs.jl.

### 2016-05-31 Paper_v1.tex
Submitted

@@
 JuPOETs integrates simulated annealing with Pareto optimality to estimate parameter ensembles on or near the optimal tradeoff surface between competing training objectives.
-We demonstrated JuPOETs on a suite of multiobjective problems, including test functions with parameter bounds and system constraints as well as for the identification of a proof-of-concept biochemical model with four conflicting training objectives.
+We demonstrate JuPOETs on a suite of multiobjective problems, including test functions with parameter bounds and system constraints as well as for the identification of a proof-of-concept biochemical model with four conflicting training objectives.
 JuPOETs identified optimal or near optimal solutions approximately six-fold faster than a corresponding implementation in Octave for the suite of test functions.
@@
 including mutant data \cite{Khodayari:2014aa}.
-We and others have used ensemble approaches, generated using both sampling and optimization techniques, to robustly simulate a wide variety of signal transduction processes \citep{Luan:2007aa,Song:2009aa,Tasseff:2010aa,Tasseff:2011aa,Nayak:2011aa}, neutrophil trafficking in sepsis \cite{Song:2012aa},
-patient specific coagulation behavior \cite{Luan:2010aa} and to capture cell to cell variation \cite{Lequieu:2011aa}.
+We and others have used ensemble approaches, generated using both sampling and optimization techniques,
+to robustly simulate a wide variety of signal transduction processes \citep{Luan:2007aa,Song:2009aa,Tasseff:2010aa,Tasseff:2011aa,Nayak:2011aa}, neutrophil trafficking in sepsis \cite{Song:2012aa},
+patient specific coagulation behavior \cite{Luan:2010aa}, uncertainty quantification in metabolic kinetic models \cite{Andreozzi:2016aa} and to capture cell to cell variation \cite{Lequieu:2011aa}.
 Thus, ensemble approaches are widely used to robustly simulate a variety of biochemical systems.
@@
 This study was supported by an award from the National Science Foundation (NSF CBET-0955172) and the National Institutes of Health (NIH HL110328) to J.B, and by
-a National Science Foundation Graduate Research Fellowship (DGE-1144153) to D.B.
+a National Science Foundation Graduate Research Fellowship (DGE-1144153) to D.B. Lastly, J.V was supported by an award from
+the US Army and Systems Biology of Trauma Induced Coagulopathy (W911NF-10-1-0376).
 

### 2016-06-01 Paper_v1.tex
Added some formatting items

@@
 \begin{abstract} % abstract
+\parttitle{Background}
 Ensemble modeling is a well established approach for obtaining robust predictions and for simulating course grained population behavior in deterministic mathematical models.
@@
 JuPOETs integrates simulated annealing with Pareto optimality to estimate parameter ensembles on or near the optimal tradeoff surface between competing training objectives.
+
+\parttitle{Results}
 We demonstrate JuPOETs on a suite of multiobjective problems, including test functions with parameter bounds and system constraints as well as for the identification of a proof-of-concept biochemical model with four conflicting training objectives.
@@
 while simultaneously estimating parameter sets that performed well on each of the individual objective functions.
+
+\parttitle{Conclusions}
 JuPOETs can be adapted to solve many problem types, including mixed binary and continuous variable types, bilevel optimization problems
@@
 
+\section*{Ethics approval and consent to participate}
+Not applicable
+
+\section*{Consent to publish}
+Not applicable
+
 \section*{Competing interests}
@@
     The manuscript was prepared and edited for publication by D.B, J.B. and J.V.
+
+\section*{Acknowledgements}
+We gratefully acknowledge Ani Chakrabarti, Russell Gould and Kathy Rogers for their input and suggestions regarding new features to include into JuPOETs.
+We also gratefully acknowledge the suggestions from the anonymous reviewers to improve this manuscript and JuPOETs.
+
+\section*{Availability of data and materials}
+JuPOETs is open source, available under an MIT software license.
+The JuPOETs source code is freely available from the JuPOETs GitHub repository at https://github.com/varnerlab/POETs.jl.
+All samples used in this study are included in the \texttt{sample} subdirectory of the JuPOETs GitHub repository.
+
+\section*{List of abbreviations}
 

### 2016-09-06 Paper_v2.tex
edits_first_review

--- 2016-05-31 Paper_v1.tex
@@
 \begin{abstract} % abstract
-Ensemble modeling is a well established approach for obtaining robust predictions and for simulating course grained population behavior in deterministic mathematical models.
-In this study, we present a multiobjective based technique to estimate model ensembles, the Pareto Optimal Ensemble Technique in the Julia programming language (JuPOETs).
-JuPOETs integrates simulated annealing with Pareto optimality to estimate parameter ensembles on or near the optimal tradeoff surface between competing training objectives.
+\parttitle{Background}
+Ensemble modeling is a promising approach for obtaining robust predictions and course grained population behavior in deterministic mathematical models.
+Ensemble approaches address model uncertainty by using parameter or model families instead of single best-fit parameters or fixed model structures.
+Parameter ensembles can be selected based upon simulation error, along with other criterion such as diversity or steady-state performance.
+Simulations using parameter ensembles can estimate confidence intervals on model variables, and robustly constrain model predictions,
+despite having many poorly constrained parameters.
+
+\parttitle{Results}
+In this study, we present a multiobjective based technique to estimate parameter or models ensembles, the Pareto Optimal Ensemble Technique in the Julia programming language (JuPOETs).
+JuPOETs integrates simulated annealing with Pareto optimality to estimate ensembles on or near the optimal tradeoff surface between competing training objectives.
 We demonstrate JuPOETs on a suite of multiobjective problems, including test functions with parameter bounds and system constraints as well as for the identification of a proof-of-concept biochemical model with four conflicting training objectives.
@@
 while simultaneously estimating parameter sets that performed well on each of the individual objective functions.
+
+\parttitle{Conclusions}
+JuPOETs is a promising approach for the estimation of parameter and model ensembles using multiobjective optimization.
 JuPOETs can be adapted to solve many problem types, including mixed binary and continuous variable types, bilevel optimization problems
 and constrained problems without altering the base algorithm.
-JuPOETs can be installed using the Julia package manager from the JuPOETs GitHub repository at https://github.com/varnerlab/POETs.jl.
+JuPOETs is open source available under an MIT license, and can be installed using the Julia package manager from the JuPOETs GitHub repository
 \end{abstract}
@@
 \section*{Background}
-Ensemble modeling is a well established approach for obtaining robust predictions and course grained population behavior in large deterministic mathematical models.
+Ensemble modeling is a promising approach for obtaining robust predictions and course grained population behavior in deterministic mathematical models.
 It is often not possible to uniquely identify all the parameters in biochemical models, even when given extensive training data \cite{Gadkar:2005ad}.
@@
 by using parameter families instead of single best-fit parameter sets.
-Parameter families can be selected based upon simulation error, along with other criterion such as diversity or steady-state peformance.
+Parameter families can be selected based upon simulation error, along with other criterion such as diversity or steady-state performance.
 Simulations using parameter ensembles can estimate confidence intervals on model variables, and robustly constrain model predictions,
@@
 However, because both the neighbor and objective functions are user defined, different constraint implementations are easily defined.
-JuPOETs can be installed using the Julia package manager from the JuPOETs repository at https://github.com/varnerlab/POETs.jl.
-Sample code is included in the \texttt{sample/biochemical} subdirectory of the JuPOETs repository to help users get started using JuPOETs in their projects.
+
+\section*{Availability of data and materials}
+JuPOETs is open source, available under an MIT software license.
+The JuPOETs source code is freely available from the JuPOETs GitHub repository at https://github.com/varnerlab/POETs.jl.
+All samples used in this study are included in the \texttt{sample/biochemical} and \texttt{sample/test\_functions} subdirectories of the JuPOETs GitHub repository.
 
@@
 \section*{Conclusions}
+JuPOETs is a promising approach for the estimation of parameter and model ensembles using multiobjective optimization.
 JuPOETs is a significant advance over the previous POETs implementation.
@@
 First, JuPOETs should be compared with other multiobjective evolutionary algorithms (MOEAs) to determine its relative performance on test and real world problems.
-Many evolutionary approaches e.g., the nondominated sorting genetic algorithm (NSGA) family of algorithms, have been adapted to solve multiobjective optimization problems \cite{DEB2002,HUBAND2006}.
+Many evolutionary approaches e.g., the non-dominated sorting genetic algorithm (NSGA) family of algorithms, have been adapted to solve multiobjective optimization problems \cite{DEB2002,HUBAND2006}.
 It is unclear if JuPOETs will perform as well as these other approaches;
@@
 Currently, JuPOETs serially evaluates the objective function vector.
-However, parallel evaluation of the objective functions could be easily implemented using a variety of techniques without changing the main run loop of JuPOETs.
+However, parallel evaluation of the objective functions e.g., using the \texttt{\@parallel} Julia macro or other techniques,
+could be easily implemented without changing the main run loop of JuPOETs.
 Because of the flexible function pointer architecture of JuPOETs, the only changes required are in the user defined objective function.
@@
     The manuscript was prepared and edited for publication by D.B, J.B. and J.V.
+
+\section*{Acknowledgements}
+We gratefully acknowledge Ani Chakrabarti, Russell Gould and Kathy Rogers for their input and suggestions regarding new features to include into JuPOETs.
+We also gratefully acknowledge the suggestions from the anonymous reviewers to improve this manuscript and JuPOETs.
+
+\section*{Availability of data and materials}
+JuPOETs is open source, available under an MIT software license.
+The JuPOETs source code is freely available from the JuPOETs GitHub repository at https://github.com/varnerlab/POETs.jl.
+All samples used in this study are included in the \texttt{sample/biochemical} and \texttt{sample/test\_functions} subdirectories of the JuPOETs GitHub repository.
+
+\section*{List of abbreviations}
+Not applicable
 

### 2016-09-21 Paper_BioArXiv_Submitted_v1.tex
Starting on v2 with RR

@@
 a popular web-based Git repository hosting service offering distributed revision control and source code management.
-Thus, JuPOETs can be adapted to many problem types, including mixed binary and continuous variable types, bilevel problems and constrained problems without altering the base algorithm, as was required in the previous POETs implementation.
+Thus, JuPOETs can be adapted to many problem types, including mixed binary and continuous variable types,
+bilevel problems and constrained problems without altering the base algorithm, as was required in the previous POETs implementation.
 
@@
 a parameter set as an input and returns a $\mathcal{K}\times{1}$ objective vector. Candidate parameter sets are generated by the user supplied \texttt{neighbor} function.
-The error vector associated with $\mathbf{p}_{i+1}$ is ranked using the builtin Pareto \texttt{rank} function, by comparing the current error at iteration $i+1$ to the error archive $\mathcal{O}_{i}$ (all error vectors up to iteration $i-1$ meeting a ranking criteria).
-Pareto rank is a measure of distance from the trade-off surface;
+The error vector associated with $\mathbf{p}_{i+1}$ is ranked using the builtin Pareto \texttt{rank} function, by comparing the current error at iteration $i+1$
+to the error archive $\mathcal{O}_{i}$ (all error vectors up to iteration $i-1$ meeting a ranking criteria). Pareto rank is a measure of distance from the trade-off surface;
 parameter sets on or near the optimal trade-off surface between the objectives have a rank equal to $0$ (no other current parameter sets are better).

### 2016-09-21 response_to_reviewers/Paper_v2.tex
Starting on v2 with RR

--- 2016-09-06 Paper_v2.tex
@@
 \parttitle{Background}
-Ensemble modeling is a promising approach for obtaining robust predictions and course grained population behavior in deterministic mathematical models.
+Ensemble modeling is a promising approach for obtaining robust predictions and coarse grained population behavior in deterministic mathematical models.
 Ensemble approaches address model uncertainty by using parameter or model families instead of single best-fit parameters or fixed model structures.
-Parameter ensembles can be selected based upon simulation error, along with other criterion such as diversity or steady-state performance.
+Parameter ensembles can be selected based upon simulation error, along with other criteria such as diversity or steady-state performance.
 Simulations using parameter ensembles can estimate confidence intervals on model variables, and robustly constrain model predictions,
@@
 and constrained problems without altering the base algorithm.
-JuPOETs is open source available under an MIT license, and can be installed using the Julia package manager from the JuPOETs GitHub repository
+JuPOETs is open source, available under an MIT license, and can be installed using the Julia package manager from the JuPOETs GitHub repository
 \end{abstract}
@@
 \section*{Background}
-Ensemble modeling is a promising approach for obtaining robust predictions and course grained population behavior in deterministic mathematical models.
+Ensemble modeling is a promising approach for obtaining robust predictions and coarse grained population behavior in deterministic mathematical models.
 It is often not possible to uniquely identify all the parameters in biochemical models, even when given extensive training data \cite{Gadkar:2005ad}.
@@
 by using parameter families instead of single best-fit parameter sets.
-Parameter families can be selected based upon simulation error, along with other criterion such as diversity or steady-state performance.
+Parameter families can be selected based upon simulation error, along with other criteria such as diversity or steady-state performance.
 Simulations using parameter ensembles can estimate confidence intervals on model variables, and robustly constrain model predictions,
@@
 Metropolis-type random walk \cite{Battogtokh:2002,Brown:2003,Tasseff:2010aa,Tasseff:2011aa}.
-Liao and coworkers developed methods that generate ensembles that all approach the same steady-state, for example one determined by fluxomics measurements \cite{Tran:2008aa}.
+Liao and coworkers developed methods to generate ensembles that all approach the same steady-state, for example one determined by fluxomics measurements \cite{Tran:2008aa}.
 They have used this approach for model reduction \cite{Tan:2011aa}, strain engineering \cite{Contador2009221,Tan:2012aa} and to study the robustness of non-native pathways and network failure \cite{Lee:2014aa}.
@@
 We and others have used ensemble approaches, generated using both sampling and optimization techniques,
-to robustly simulate a wide variety of signal transduction processes \citep{Luan:2007aa,Song:2009aa,Tasseff:2010aa,Tasseff:2011aa,Nayak:2011aa}, neutrophil trafficking in sepsis \cite{Song:2012aa},
+that have robustly simulated a wide variety of signal transduction processes \citep{Luan:2007aa,Song:2009aa,Tasseff:2010aa,Tasseff:2011aa,Nayak:2011aa}, neutrophil trafficking in sepsis \cite{Song:2012aa},
 patient specific coagulation behavior \cite{Luan:2010aa}, uncertainty quantification in metabolic kinetic models \cite{Andreozzi:2016aa} and to capture cell to cell variation \cite{Lequieu:2011aa}.
+Further, ensemble approaches have been used in synthetic biology to sample possible biocircuit configurations and to design folding behaviors of novel mRNAs \cite{Otero-Muras2014,Taneda2015}.
 Thus, ensemble approaches are widely used to robustly simulate a variety of biochemical systems.
@@
 Identification of biochemical models with hundreds or even thousands of states and parameters may not be tractable as a single objective optimization problem.
-Further, large models require significant training data perhaps taken from diverse sources, for example different laboratories or cell-lines.
-These data are often heterogenous, and contain intrinsic conflicts that complicate parameter estimation.
-Parameter ensembles which optimally balance tradeoffs between submodels and conflicts in training data can
-lead to robust model performance. Multiobjective optimization is an ensemble generation technique that naturally balances conflicting training data.
+Further, large models require significant training data perhaps taken from diverse sources.
+These data often contain intrinsic conflicts resulting from, for example, the use of different cell lines or cross laboratory variability.
+Parameter ensembles that optimally balance tradeoffs between submodels and conflicts in training data can
+lead to robust model performance. 
+Multiobjective optimization is an ensemble generation technique that naturally balances conflicting training data and is preferred over single objective optimization in cases where optimizing one objective will hurt the performance of other objectives \cite{Bandyopadhyay2013}.
 Previously, we developed the Pareto Optimal Ensemble Technique (POETs) algorithm to address the challenge of competing or conflicting objectives.
@@
 Further, the architecture of JuPOETs takes advantage of the first-class function type in Julia allowing user definable behavior for all key aspects of the algorithm, including objective functions, custom diversity generation logic, linear/non-linear parameter constraints (and parameter bounds constraints) as well as custom cooling schedules.
-Julia's ability to naturally call other languages such as Python or C also allows JuPOETS to be used with models implemented in a variety of languages across many platforms.
+Julia's ability to naturally call other languages such as Python or C also allows JuPOETs to be used with models implemented in a variety of languages across many platforms.
 Additionally, Julia offers a built-in package manager which is directly integrated with GitHub,
@@
 as well as $\mathcal{C}$ problem specific constraints $g_i\left(t,\mathbf{x}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p}\right),i=1,\hdots,\mathcal{C}$.
-
 JuPOETs integrates simulated annealing with Pareto optimality to estimate parameter sets on or near the optimal tradeoff surface between competing training objectives
@@
 a parameter set as an input and returns a $\mathcal{K}\times{1}$ objective vector. Candidate parameter sets are generated by the user supplied \texttt{neighbor} function.
-The error vector associated with $\mathbf{p}_{i+1}$ is ranked using the builtin Pareto \texttt{rank} function, by comparing the current error at iteration $i+1$ to the error archive $\mathcal{O}_{i}$ (all error vectors up to iteration $i-1$ meeting a ranking criteria).
+The error vector associated with $\mathbf{p}_{i+1}$ is ranked using the builtin Pareto \texttt{rank} function, by comparing the current error at iteration $i+1$ to the error archive $\mathcal{O}_{i}$ (all error vectors up to iteration $i-1$ meeting a ranking criterion).
 Pareto rank is a measure of distance from the trade-off surface;
 parameter sets on or near the optimal trade-off surface between the objectives have a rank equal to $0$ (no other current parameter sets are better). 
+These rank zero parameter sets define the Pareto optimal group for the ensemble, wherein Pareto optimality is defined as a parameter set not being dominated by any other sets within the ensemble.
+The surface within objective space on which these Pareto optimal parameter sets lie is called the tradeoff surface.
 Sets with increasing non-zero rank are progressively further away from the optimal trade-off surface.
@@
 We used the Pareto rank to inform the SA calculation.
-The parameter set $\mathbf{p}_{i+1}$ was accepted or rejected by the SA, by calculating an \texttt{acceptance} probability $\mathcal{P}\left(\mathbf{p}_{i+1}\right)$:
+The parameter set $\mathbf{p}_{i+1}$ was accepted or rejected by the SA at each iteration, by calculating an \texttt{acceptance} probability $\mathcal{P}\left(\mathbf{p}_{i+1}\right)$:
 \begin{equation}\label{eqn_costMOSA}
@@
 \end{equation}
-where $T$ is the computational annealing temperature.
+where $T$ is the computational annealing temperature that provides control over how strictly Pareto rank is enforced. 
 As $\texttt{rank}\left(\mathcal{O}_{i+1}\left(\mathbf{p}_{i+1}\right)\mid \mathcal{O}_{i}\right)\rightarrow{0}$, the acceptance probability moves toward one,
 ensuring that we explore parameter sets along the Pareto surface.
-Occasionally, (depending upon $T$) a parameter set with a high Pareto rank was accepted by the SA allowing a more diverse search of the parameter space.
-However, as $T$ is reduced, the probability of accepting a high-rank set occurring decreases.
-Parameter sets could also be accepted by the SA but $\emph{not}$ permanently archived in $\mathcal{S}_{i}$.
-Only parameter sets with rank less than or equal to threshold (rank $\leq$4 by default) were included in $\mathcal{S}_{i}$, where the archive was re-ranked and filtered after
-every new parameter set was accepted.
+Occasionally, (depending upon $T$) a parameter set with a high Pareto rank is  accepted by the SA allowing a more diverse search of the parameter space.
+However, as $T$ is reduced, the probability of accepting a high-rank set decreases.
+Parameter sets could also be accepted by the SA but $\emph{not}$ permanently archived in $\mathcal{S}_{i}$, where $\mathcal{S}_{i}$ is the solution archive.
+Only parameter sets with rank less than or equal to a threshold (rank $\leq$4 by default) are included in $\mathcal{S}_{i}$, where the archive is re-ranked and filtered after accepting
+every new parameter set.
 Parameter bounds were implemented in the \texttt{neighbor} function as box constraints, while problem specific constraints were implemented in \texttt{objective} using a penalty method:
@@
 The synthetic data was generated by randomly varying these parameters.
-JuPOETs produced an ensemble of approximately $\dim{\mathcal{S}}\simeq$ 13,000 parameters that captured the mean of the measured data sets for extracellular metabolites and cellmass
-(Fig. \ref{fig:fig-biochemical}A and B). JuPOETs minimized the difference between the simulated and measured values for A$_{e}$, B$_{e}$, C$_{e}$ and cellmass,
+JuPOETs produced an ensemble of approximately $\dim{\mathcal{S}}\simeq$ 13,000 parameter sets that captured the mean of the measured data sets for extracellular metabolites and cellmass
+(Fig. \ref{fig:fig-biochemical}A and B). JuPOETs minimized the difference between the simulated and measured values for extracellular metabolites A$_{e}$, B$_{e}$, C$_{e}$ and cellmass,
 where the residual for each data set was treated as a single objective (leading to four objectives).
@@
 Many evolutionary approaches e.g., the non-dominated sorting genetic algorithm (NSGA) family of algorithms, have been adapted to solve multiobjective optimization problems \cite{DEB2002,HUBAND2006}. 
-It is unclear if JuPOETs will perform as well as these other approaches;
-one potential advantage that JuPOETs may have is the local refinement step which temporarily reduces the problem to a single objective formulation.
+Since there is a lack of open source NSGA implementations in Julia, we were not able to compare JuPOETs to similar approaches.
+One potential advantage that JuPOETs may have is the local refinement step which temporarily reduces the problem to a single objective formulation.
 Previously, this hybrid approach led to better convergence on a proof-of-concept signal transduction model \cite{Song:2010aa}.
+Such hybrid methods have also shown to be efficient in NSGA implementations on biochemical systems \cite{Otero-Muras2014,Sendin2006}. 
 For many real world parameter estimation problems, the bulk of the execution time is spent evaluating the objective functions.
@@
   C: Trade-off plots between the four training objectives. The quantity $O_{j}$ denotes the jth training objective.
-  Each point represents a member of the parameter ensemble, where gray denotes rank 0 sets, while black denotes rank 1 sets.
+  Each point represents a member of the parameter ensemble, where gray denotes rank 0 sets, while black denotes rank 1 sets. Ensembles were generated using POETs without employing local refinement. 
    }\label{fig:fig-biochemical}

### 2016-09-21 response_to_reviewers/Paper_v2.tex
More changes

@@
 the problem of estimating model parameters from experimental data remains challenging.
-Ensemble approaches address parameter uncertainty in systems biology and other fields like weather prediction \cite{Battogtokh:2002,Kuepfer:2007, Brown:2003,Palmer:2005}
+Ensemble approaches address parameter uncertainty in systems biology and other fields like weather prediction \cite{Battogtokh:2002,Kuepfer:2007aa, Brown:2003,Palmer:2005}
 by using parameter families instead of single best-fit parameter sets.

### 2016-11-17 response_to_reviewers/Paper_v2.tex
Some of reviewer 1 changes

@@
 
-Identification of biochemical models with hundreds or even thousands of states and parameters may not be tractable as a single objective optimization problem.
-Further, large models require significant training data perhaps taken from diverse sources.
-These data often contain intrinsic conflicts resulting from, for example, the use of different cell lines or cross laboratory variability.
-Parameter ensembles that optimally balance tradeoffs between submodels and conflicts in training data can
-lead to robust model performance.
-Multiobjective optimization is an ensemble generation technique that naturally balances conflicting training data and is preferred over single objective optimization in cases where optimizing one objective will hurt the performance of other objectives \cite{Bandyopadhyay2013}.
+Identification of biochemical models often requires significant training data perhaps taken from diverse sources.
+These data often can contain intrinsic conflicts resulting from, for example, the use of different cell lines or cross laboratory variability.
+Parameter ensembles that optimally balance these conflicts in training data can lead to more robust model performance.
+Multiobjective optimization is an ensemble generation technique that naturally balances conflicting training data \cite{Handl:2007aa}.
+Multiobjective optimization has been used us and others to identify biochemical models [REFHERE].
 Previously, we developed the Pareto Optimal Ensemble Technique (POETs) algorithm to address the challenge of competing or conflicting objectives.
@@
 \section*{Implementation}
-In this study, we present an open-source implementation of the Pareto optimal ensemble technique in the Julia programming language (JuPOETs).
+In this software note, we present an open-source implementation of the Pareto optimal ensemble technique in the Julia programming language (JuPOETs).
 JuPOETs offers many advantages and improvements compared to Octave-POETs.
@@
 \end{equation}
-subject to:
+subject to the constraints:
 \begin{eqnarray}\nonumber
@@
 \end{equation}using a modified simulated annealing approach.
+The quantity $O_{j}$ denotes the $j^{th}$ objective function ($j=1,2,\hdots,\mathcal{K}$), typically the sum of squared errors for the $j^{th}$ data set for biochemical modeling applications.
+The terms $\mathbf{f}(t,\mathbf{x}(t,\mathbf{p}),\dot{\mathbf{x}}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p})$ denote the system of model equations (e.g., differential equations,
+differential algebraic equations or linear/non-linear algebraic equations) where $\mathbf{p}$ denotes the decision variable vector e.g., unknown model parameters ($\mathcal{D}\times~1$).
+In typical biochemical modeling applications, the model equations $\mathbf{f}\left(\cdot\right)$
+are a system of continuous real-valued non-linear differential equations that comprise a kinetic model, but other types of models e.g., stoichiometric models are also common.
 The quantity $t$ denotes time, $\mathbf{x}\left(t,\mathbf{p}\right)$ denotes the model state (with an initial state $\mathbf{x}_{0}$), and $\mathbf{u}(t)$ denotes an input vector.
-The terms $\mathbf{f}(t,\mathbf{x}(t,\mathbf{p}),\dot{\mathbf{x}}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p})$ denote the system of model equations (e.g., differential equations,
-differential algebraic equations or linear/non-linear algebraic equations) where $\mathbf{p}$ denotes the unknown parameter vector ($\mathcal{D}\times~1$).
-The parameter search can be subject to parameter bound constraints, where $\mathcal{L}$ and $\mathcal{U}$ denote the lower and upper parameter bounds, respectively
+The decision variables (e.g., kinetic parameters) can be subject to bounds constraints, where $\mathcal{L}$ and $\mathcal{U}$ denote the lower and upper bounds, respectively
 as well as $\mathcal{C}$ problem specific constraints $g_i\left(t,\mathbf{x}(t,\mathbf{p}),\mathbf{u}(t),\mathbf{p}\right),i=1,\hdots,\mathcal{C}$.
-JuPOETs integrates simulated annealing with Pareto optimality to estimate parameter sets on or near the optimal tradeoff surface between competing training objectives
+The decision variables $\mathbf{p}$ are typically real-valued kinetic constants, or metabolic fluxes in the case of stoichiometric models.
+However, other variables types e.g., binary or categorical decision variables can also be accommodated.
+
+%Pareto rank measures the quality of a parameter solution relative to the optimal tradeoff surface between competing training objectives.
+%The surface within objective space on which these Pareto optimal parameter sets lie is called the tradeoff surface.
+
+JuPOETs integrates simulated annealing (SA) \cite{Kirkpatrick:1983aa} with Pareto optimality to estimate parameter sets on or near the optimal tradeoff surface between competing training objectives
 (Fig. \ref{fig:fig-MOSAalgorithm} and Algorithm \ref{algo:JuPOETs-pcode}).
-The central idea of POETs is a mapping between the value of the objective vector evaluated at $\mathbf{p}_{i+1}$ (parameter guess at iteration $i+1$) and Pareto rank.
+A tradeoff surface defines the best possible error value for every conflicting objective, and Pareto rank is a measure of distance away from the optimal tradeoff surface.
+Thus, the central idea underlying POETs is a mapping between the value of the objective vector evaluated at $\mathbf{p}_{i+1}$ (parameter guess at iteration $i+1$) and Pareto rank (Fig.
+\ref{fig:fig-MOSAalgorithm}).
+A parameter set $\mathbf{p}_{i+1}$ lies along the optimal tradeoff surface if no other parameter guess leads to decreased error for every objective.
 JuPOETs calculates the performance of a candidate parameter set $\mathbf{p}_{i+1}$ by calling the user defined \texttt{objective} function; \texttt{objective} takes
-a parameter set as an input and returns a $\mathcal{K}\times{1}$ objective vector. Candidate parameter sets are generated by the user supplied \texttt{neighbor} function.
-The error vector associated with $\mathbf{p}_{i+1}$ is ranked using the builtin Pareto \texttt{rank} function, by comparing the current error at iteration $i+1$ to the error archive $\mathcal{O}_{i}$ (all error vectors up to iteration $i-1$ meeting a ranking criterion).
+a parameter set as an input, evaluates the model equations, and using this solution, returns the $\mathcal{K}\times{1}$ objective vector.
+Candidate parameter sets are generated by the user supplied \texttt{neighbor} function; the default implementation of \texttt{neighbor} is a random perturbation, however other perturbation logic can be implemented by the user.
+The error vector associated with $\mathbf{p}_{i+1}$ is ranked using the builtin Pareto \texttt{rank} function, by comparing the error at iteration $i+1$ to the error archive $\mathcal{O}_{i}$ (all error vectors up to iteration $i$ meeting a ranking criterion).
 Pareto rank is a measure of distance from the trade-off surface;
@@
 These rank zero parameter sets define the Pareto optimal group for the ensemble, wherein Pareto optimality is defined as a parameter set not being dominated by any other sets within the ensemble.
-The surface within objective space on which these Pareto optimal parameter sets lie is called the tradeoff surface.
 Sets with increasing non-zero rank are progressively further away from the optimal trade-off surface.

### 2016-11-17 response_to_reviewers/Paper_v2.tex
More reviewer 1 changes

@@
 \parttitle{Results}
-In this study, we present a multiobjective based technique to estimate parameter or models ensembles, the Pareto Optimal Ensemble Technique in the Julia programming language (JuPOETs).
+In this software note, we present a multiobjective based technique to estimate parameter or models ensembles, the Pareto Optimal Ensemble Technique in the Julia programming language (JuPOETs).
 JuPOETs integrates simulated annealing with Pareto optimality to estimate ensembles on or near the optimal tradeoff surface between competing training objectives.
@@
 
+%JuPOETs offers many advantages and improvements compared to Octave-POETs.
+
 \section*{Implementation}
 In this software note, we present an open-source implementation of the Pareto optimal ensemble technique in the Julia programming language (JuPOETs).
-JuPOETs offers many advantages and improvements compared to Octave-POETs.
 JuPOETs takes advantage of the unique features of Julia.
@@
 \end{equation}
-subject to the constraints:
+subject to the model equations and constraints:
 \begin{eqnarray}\nonumber
@@
   \mathcal{L} \leq \mathbf{p} \leq \mathcal{U}
-\end{equation}using a modified simulated annealing approach.
+\end{equation}
 The quantity $O_{j}$ denotes the $j^{th}$ objective function ($j=1,2,\hdots,\mathcal{K}$), typically the sum of squared errors for the $j^{th}$ data set for biochemical modeling applications.
@@
 
-JuPOETs integrates simulated annealing (SA) \cite{Kirkpatrick:1983aa} with Pareto optimality to estimate parameter sets on or near the optimal tradeoff surface between competing training objectives
-(Fig. \ref{fig:fig-MOSAalgorithm} and Algorithm \ref{algo:JuPOETs-pcode}).
-A tradeoff surface defines the best possible error value for every conflicting objective, and Pareto rank is a measure of distance away from the optimal tradeoff surface.
+JuPOETs integrates simulated annealing (SA) \cite{Kirkpatrick:1983aa} with Pareto ranking to estimate parameter sets on or near the optimal tradeoff surface between competing objectives (Fig.
+\ref{fig:fig-MOSAalgorithm} and Algorithm \ref{algo:JuPOETs-pcode}).
+A tradeoff surface defines the best possible performance for every conflicting objective, such that an increase in the performance of one objective does not decrease the performance of at least one other objective.
+Pareto rank is a measure of distance away from the optimal tradeoff surface.
 Thus, the central idea underlying POETs is a mapping between the value of the objective vector evaluated at $\mathbf{p}_{i+1}$ (parameter guess at iteration $i+1$) and Pareto rank (Fig.
@@
 However, because both the neighbor and objective functions are user defined, different constraint implementations are easily defined.
+
+To use JuPOETs, the user specifies the \texttt{neighbor}, \texttt{acceptance}, \texttt{cooling} and \texttt{objective} functions along with an initial decision variable guess.
+If the user is operating JuPOETs in hybrid mode, then a \texttt{local\_refinement} function pointer must also be specified; the default \texttt{local\_refinement} implementation is simply a pass through function.
+Default implementations of the \texttt{neighbor}, \texttt{acceptance} and \texttt{cooling} functions can be used directly, or they can be overridden by user defined logic.
+However, the user must provide an implementation of the \texttt{objective} function and provide an initial decision variable guess.
+In addition, there are several user configurable parameters that can be adjusted to control the performance of JuPOETs:
+\texttt{maximum\_number\_of\_iterations} controls the number of iterations per temperature (default 20);
+\texttt{rank\_cutoff} controls the upper rank bound on the solution archive (default 5);
+\texttt{temperature\_min} controls the minimum temperature after which JuPOETs returns the error and solution archives (default 0.001);
+\texttt{show\_trace} controls the level of output shown to the user (default true).
+After the completion of the run, JuPOETs returns the parameter solution archive $\mathcal{S}$ and objective archive $\mathcal{O}$.
+The parameter solution archive $\mathcal{S}$ contains is an $\mathcal{D}\times\mathcal{A}$ array, where $\mathcal{A}$ denotes the number of solutions in the archive when
+JuPOETs terminated.
+On the other hand, the objective archive $\mathcal{O}$ is an $\mathcal{K}\times\mathcal{A}$ array containing the performance values for each objective corresponding the columns of $\mathcal{S}$.
+Lastly, JuPOETs returns the rank archive $\mathcal{R}$ which is an $\mathcal{A}\times{1}$ array of Pareto ranks corresponding to the
+columns of $\mathcal{S}$.
 
@@
 
-JuPOETs identified optimal or nearly optimal solutions significantly faster than Octave-POETs for a suite of multiobjective test problems (Table \ref{table:objective-table}).
+JuPOETs identified optimal or nearly optimal solutions significantly faster than Octave-POETs for a suite of multiobjective algebraic test problems (Table \ref{table:objective-table}).
+The algebraic test problems were constrained non-linear functions with bound constraints and additional non-linear constraints on the decision variables in one case.
+The problems had up to three-dimensional continuous real-valued decision vectors, and each case had two objective functions.
 The wall-clock time for JuPOETs and Octave-POETs was measured for 10 independent trials for each of the test problems.
@@
 On average, JuPOETs identified optimal or near optimal solutions for the suite of test problems six-fold faster (60s versus 400s)
-than Octave-POETs (Fig. \ref{fig:fig-timing}). JuPOETs produced the characteristic tradeoff curves for each test problem, given both parameter bound and
+than Octave-POETs (Fig. \ref{fig:fig-timing}). JuPOETs produced the characteristic tradeoff curves for each test problem, given both decision variable bound and
 problem constraints (Fig. \ref{fig:fig-surfaces}).
-Thus, JuPOETs estimated an ensemble of solutions to constrained multiobjective optimization test problems significantly faster than the current Octave implementation.
+Thus, JuPOETs estimated an ensemble of solutions to constrained multiobjective algebraic test problems significantly faster than the current Octave implementation.
 Next, we tested JuPOETs on a proof-of-concept biochemical model identification problem.
 
-JuPOETs estimated an ensemble of biochemical models that was consistent with the mean of synthetic training data (Fig. \ref{fig:fig-biochemical}).
+JuPOETs estimated an ensemble of biochemical model parameters that were consistent with the mean of synthetic training data (Fig. \ref{fig:fig-biochemical}).
 Four synthetic training data sets were generated from a prototypical biochemical network consisting of 6 metabolites and 7 reactions (Fig. \ref{fig:fig-biochemical}, inset right).
-We considered a common case in which the same measurements were made on four hypothetical cell types, each having the same biological connectivity but different performance.
-Network dynamics were modeled using the hybrid cybernetic model with elementary modes (HCM-EM) approach of Ramkrishna and coworkers \cite{2008_kim_varner_ramkrishna_BiotechProg}.
-In the HCM-EM approach, metabolic networks are first decomposed into a set of elementary modes (EMs) (chemically balanced steady-state pathways, see \cite{Schuster:2000aa}).
+We considered a common case in which the same extracellular measurements of $A_{e},B_{e},C_{e}$ and cellmass were made on four hypothetical cell types, each having the same biological connectivity but different performance.
+Network dynamics were modeled using the hybrid cybernetic model with elementary modes (HCM) approach of Ramkrishna and coworkers \cite{2008_kim_varner_ramkrishna_BiotechProg}.
+In the HCM approach, metabolic networks are first decomposed into a set of elementary modes (EMs) (chemically balanced steady-state pathways, see \cite{Schuster:2000aa}).
 Dynamic combinations of elementary modes are then used to characterize network behavior.
 Each elementary mode is catalyzed by a pseudo enzyme; thus, each mode has both kinetic and enzyme synthesis parameters.
-The proof of concept network generated 6 EMs, resulting in 13 model parameters.
-The synthetic data was generated by randomly varying these parameters.
+The proof of concept network generated 6 EMs, resulting in 13 model parameters (continuos real-valued decision variables).
+The synthetic training data was generated by randomly varying these parameters.
+
+The general form of the biochemical model optimization test problem was given by:
+\begin{equation}
+  \min_{\mathbf{p}}\left(O_{1},\hdots,O_{\mathcal{K}}\right)
+\end{equation}subject to model and bounds constraints. We considered four training data sets, thus $\mathcal{K}=4$.
+Each objective $O_{j},~j=1,\hdots,\mathcal{K}$ quantified the squared difference between the simulated ($x_{i}$) and measured extracellular species abundance ($y_{i}$)
+in the $j^{th}$ data set:
+\begin{equation}
+  O_{j} = \sum_{i}\sum_{\tau}\left(x_{i}(\tau)-y_{i}(\tau)\right)^2\qquad j=1,\hdots,\mathcal{K}
+\end{equation}where the abundance of extracellular species $i$ ($x_{i}$), the pseudo enzyme $e_{l}$ (catalyzes flux through mode $l$), and cellmass were governed by the model equations:
+\begin{eqnarray}\nonumber
+	\frac{dx_{i}}{dt}  & = &  \sum_{j = 1}^{\mathcal{R}}\sum_{l = 1}^{\mathcal{L}}\sigma_{ij}z_{jl}q_{l}\left(\mathbf{e},\mathbf{p},\mathbf{x}\right)c \qquad{i=1,\hdots,\mathcal{M}}\\\nonumber
+  \frac{de_{l}}{dt}  & = & \alpha_{l} + r_{El}\left(\mathbf{p},\mathbf{x}\right)u_{l} - \left(\beta_{l}+r_{G}\right)e_{l} \qquad l=1,\hdots,\mathcal{L} \\\nonumber
+  \frac{dc}{dt} & = & r_{G}c
+\end{eqnarray}
+where $\mathcal{R}$ and $\mathcal{M}$ denote the number of reactions and extracellular species in the model and $\mathcal{L}$ denotes the number of elementary modes.
+The quantity $\sigma_{ij}$ denotes the stoichiometric coefficient for species $i$ in reaction $j$ and $z_{jl}$ denotes the normalized flux for reaction $j$ in mode $l$.
+If $\sigma_{ij}>0$, species $i$ is produced by reaction $j$;
+if $\sigma_{ij}<0$, species $i$ is consumed by reaction $j$; if $\sigma_{ij} = 0$, species $i$ is not connected with reaction $j$.
+Extracellular species, cellmass and pseudo-enzyme were subject to the initial conditions $\mathbf{x}\left(t_{o}\right) = \mathbf{x}_{o}$, $c(t_{o})=c_{o}$ and $e_{l}=0.5$, respectively.
+The term $q_{l}\left(\mathbf{e},\mathbf{p},\mathbf{x}\right)$ denotes the specific uptake/secretion rate for mode $l$
+where $\mathbf{e}$ denotes the pseudo enzyme vector, $\mathbf{p}$ denotes the unknown kinetic parameter vector (decision variables),
+$\mathbf{x}$ denotes the extracellular species vector, and $c$ denotes the cell mass;
+$q_{l}\left(\mathbf{e},\mathbf{p},\mathbf{x}\right)$ is the product of a kinetic term ($\bar{q}_{l}$) and a control variable governing enzyme activity.
+Flux through each mode was catalyzed by a pseudo enzyme $e_{l}$, synthesized at the regulated specific rate $r_{E,l}\left(\mathbf{p},\mathbf{x}\right)$, and constitutively at the rate $\alpha_{l}$.
+The term $u_{l}$ denotes the cybernetic variable controlling the synthesis of enzyme $l$.
+The term $\beta_{l}$ denotes the rate constant governing non-specific enzyme degradation, and $r_{G}$ denotes the specific growth rate through all modes.
+The specific uptake/secretion rates and the specific rate of enzyme synthesis were modeled using saturation kinetics.
+The specific growth rate was given by:
+\begin{equation}\nonumber
+	r_{G}  = \sum_{l = 1}^{\mathcal{L}}z_{\mu l}q_{l}\left(\mathbf{e},\mathbf{p},\mathbf{x}\right)
+\end{equation}
+where $z_{\mu l}$ denotes the growth flux $\mu$ through mode $l$.
+The control variables $u_{l}$ and $v_{l}$, which control the synthesis and activity of each enzyme respectively, were given by:
+\begin{equation}
+	u_{l} = \frac{z_{sl}\bar{q}_{l}}{\sum\limits_{l = 1}^{\mathcal{L}}z_{sl}\bar{q}_{l}}
+\end{equation}
+and
+\begin{equation}
+    v_{l} = \frac{z_{sl}\bar{q}_{l}}{\max\limits_{l=1,\hdots,\mathcal{L}}z_{sl}\bar{q}_{l}}
+\end{equation}
+where $z_{sl}$ denotes the uptake flux of substrate $s$ through mode $l$.
+Each unknown kinetic parameter was continuous and real-valued, and subject to bounds constraints: $\mathcal{L} \leq \mathbf{p} \leq \mathcal{U}$.
+
 JuPOETs produced an ensemble of approximately $\dim{\mathcal{S}}\simeq$ 13,000 parameter sets that captured the mean of the measured data sets for extracellular metabolites and cellmass

### 2016-11-17 response_to_reviewers/Paper_v2.tex
Changes to Alogirithm 1

@@
 To use JuPOETs, the user specifies the \texttt{neighbor}, \texttt{acceptance}, \texttt{cooling} and \texttt{objective} functions along with an initial decision variable guess.
-If the user is operating JuPOETs in hybrid mode, then a \texttt{local\_refinement} function pointer must also be specified; the default \texttt{local\_refinement} implementation is simply a pass through function.
+If the user is operating JuPOETs in hybrid mode, then a \texttt{refinement} function pointer must also be specified; the default \texttt{local\_refinement} implementation is simply a pass through function.
 Default implementations of the \texttt{neighbor}, \texttt{acceptance} and \texttt{cooling} functions can be used directly, or they can be overridden by user defined logic.
@@
   \SetKwFunction{acceptance}{acceptance}
+  \SetKwFunction{refinement}{refinement}
   \SetKwFunction{rank}{rank}
@@
   \BlankLine
-  initialize: $\mathcal{R}$, $\mathcal{S}$ and $\mathcal{O}$ using initial guess\;
+  initialize: $\mathcal{R}$, $\mathcal{S}$ and $\mathcal{O}$ using initial guess $\mathbf{p}_{o}$\;
   initialize: T $\leftarrow$1.0\;
@@
   initialize: Maximum number of steps per temperature $\mathcal{I}$\;
+
+  \BlankLine
+  \BlankLine
+
+  \tcp{Locally refine initial guess (single objective problem)}
+  $\mathbf{p}_{o}$ $\leftarrow$ user-function:\refinement{$\mathbf{p}_{o}$}\;
 

### 2016-11-18 response_to_reviewers/Paper_v2.tex
Almost done with reviewer 1

@@
 To use JuPOETs, the user specifies the \texttt{neighbor}, \texttt{acceptance}, \texttt{cooling} and \texttt{objective} functions along with an initial decision variable guess.
-If the user is operating JuPOETs in hybrid mode, then a \texttt{refinement} function pointer must also be specified; the default \texttt{local\_refinement} implementation is simply a pass through function.
+If the user is operating JuPOETs in hybrid mode, then a \texttt{refinement} function pointer must also be specified; the default \texttt{refinement} implementation is simply a pass through function.
 Default implementations of the \texttt{neighbor}, \texttt{acceptance} and \texttt{cooling} functions can be used directly, or they can be overridden by user defined logic.
@@
 \texttt{show\_trace} controls the level of output shown to the user (default true).
-After the completion of the run, JuPOETs returns the parameter solution archive $\mathcal{S}$ and objective archive $\mathcal{O}$.
+After the completion of the run, JuPOETs returns the parameter solution archive $\mathcal{S}$, objective archive $\mathcal{O}$ and rank archive $\mathcal{R}$.
 The parameter solution archive $\mathcal{S}$ contains is an $\mathcal{D}\times\mathcal{A}$ array, where $\mathcal{A}$ denotes the number of solutions in the archive when
@@
 
-  \Input{User specified neighbor, objective, acceptance and cooling functions. Initial parameter guess ($\mathcal{D}\times~1$)}
-  \Output{Rank archive $\mathcal{R}$, parameter solution archive $\mathcal{S}$ and objective archive $\mathcal{O}$}
+  \Input{User specified objective function, and initial guess ($\mathcal{D}\times~1$).
+  User can also specify custom neighbor, acceptance. cooling and refinement functions or use the default functions provided.}
+  \BlankLine
+  \Output{Rank archive $\mathcal{R}$ ($\mathcal{A}\times{1}$),
+  parameter solution archive $\mathcal{S}$ ($\mathcal{D}\times\mathcal{A}$)
+  and objective archive $\mathcal{O}$ ($\mathcal{K}\times\mathcal{A}$), where $\mathcal{A}$ denotes the number of accepted solutions}
 
@@
 
-  \tcp{Locally refine initial guess (single objective problem)}
+  \tcp{Call to local refinement function (single objective problem)}
   $\mathbf{p}_{o}$ $\leftarrow$ user-function:\refinement{$\mathbf{p}_{o}$}\;
@@
 \vspace{0.05in}
-\caption{Pseudo-code for the main run-loop of JuPOETs.
-The user specifies the \texttt{neighbor}, \texttt{acceptance}, \texttt{cooling} and \texttt{objective} functions along with an initial parameter guess.
+\caption{Pseudo-code for the JuPOETs run-loop.
+The user must specify the \texttt{objective} function and an initial parameter guess.
+The user can optionally specify the \texttt{neighbor}, \texttt{acceptance}, \texttt{cooling} and \texttt{refinement} functions (or use the default implementations).
 The rank archive $\mathcal{R}$, solution archive $\mathcal{S}$ and objective archive $\mathcal{O}$ are initialized from the initial guess.
-The initial guess is perturbed in the \texttt{neighbor} function, which generates a new solution whose performance is evaluated using the user supplied \texttt{objective} function.
+The initial guess (potentially following a single objective local refinement step)
+is perturbed in the \texttt{neighbor} function, which generates a new solution whose performance is evaluated using the user supplied \texttt{objective} function.
 The new solution and objective values are then added to the respective archives and ranked using the builtin \texttt{rank} function.
@@
 However, if the solution is not accepted, it is removed from the archive and discarded.
-The computational temperature is adjusted using the user supplied \texttt{cooling} function after each $\mathcal{I}$ iterations.
-}\label{algo:JuPOETs-pcode}
+The temperature is adjusted using the user supplied \texttt{cooling} function after each $\mathcal{I}$ iterations.
+When JuPOETs terminates, the parameter solution archive $\mathcal{S}$, objective archive $\mathcal{O}$ and rank archive $\mathcal{R}$ are retuned to the caller.}\label{algo:JuPOETs-pcode}
 \end{algorithm}
@@
 
-The general form of the biochemical model optimization test problem was given by:
+The general form of the biochemical test problem was given by:
 \begin{equation}
   \min_{\mathbf{p}}\left(O_{1},\hdots,O_{\mathcal{K}}\right)
-\end{equation}subject to model and bounds constraints. We considered four training data sets, thus $\mathcal{K}=4$.
+\end{equation}subject to model and bounds constraints. We considered four training data sets ($\mathcal{K}=4$),
+each of which contained time-series measurements of $A_{e},B_{e},C_{e}$ and cellmass.
 Each objective $O_{j},~j=1,\hdots,\mathcal{K}$ quantified the squared difference between the simulated ($x_{i}$) and measured extracellular species abundance ($y_{i}$)
@@
   O_{j} = \sum_{i}\sum_{\tau}\left(x_{i}(\tau)-y_{i}(\tau)\right)^2\qquad j=1,\hdots,\mathcal{K}
-\end{equation}where the abundance of extracellular species $i$ ($x_{i}$), the pseudo enzyme $e_{l}$ (catalyzes flux through mode $l$), and cellmass were governed by the model equations:
+\end{equation}where, $i$ denotes the species index and $\tau$ denotes the time index.
+The abundance of extracellular species $i$ ($x_{i}$), the pseudo enzyme $e_{l}$ (catalyzes flux through mode $l$), and cellmass were governed by the model equations:
 \begin{eqnarray}\nonumber
@@
 
+%JuPOETs offers improved performance and is highly adaptable compared with the previous implementation in Octave.
+
 \section*{Conclusions}
-JuPOETs is a promising approach for the estimation of parameter and model ensembles using multiobjective optimization.
-JuPOETs is a significant advance over the previous POETs implementation.
-It offers improved performance and is highly adaptable to different problem types.
-We demonstrated JuPOETs on a suite of test problems, and a proof-of-concept biochemical model.
-However, there are several areas that could be explored further to improve JuPOETs.
+In this software note, we presented JuPOETs, a multiobjective technique to estimate parameter ensembles in the Julia programming language.
+JuPOETs is open source, and available for download under an MIT license from the JuPOETs GitHub repository at https://github.com/varnerlab/POETs.jl.
+We demonstrated JuPOETs on a suite of algebraic test problems, and a proof-of-concept ODE based biochemical model.
+While JuPOETs outperformed (and was significantly more flexible) than the previous Octave implementation, there are several areas that could be explored further.
 First, JuPOETs should be compared with other multiobjective evolutionary algorithms (MOEAs) to determine its relative performance on test and real world problems.
-Many evolutionary approaches e.g., the non-dominated sorting genetic algorithm (NSGA) family of algorithms, have been adapted to solve multiobjective optimization problems \cite{DEB2002,HUBAND2006}.
-Since there is a lack of open source NSGA implementations in Julia, we were not able to compare JuPOETs to similar approaches.
-One potential advantage that JuPOETs may have is the local refinement step which temporarily reduces the problem to a single objective formulation.
-Previously, this hybrid approach led to better convergence on a proof-of-concept signal transduction model \cite{Song:2010aa}.
-Such hybrid methods have also shown to be efficient in NSGA implementations on biochemical systems \cite{Otero-Muras2014,Sendin2006}.
-For many real world parameter estimation problems, the bulk of the execution time is spent evaluating the objective functions.
-One strategy to improve performance could be to optimize surrogates \cite{SURROGATES}, while another would be parallel execution of the objective functions.
+Many evolutionary approaches e.g., the non-dominated sorting genetic algorithm (NSGA) family of algorithms, have been adapted to solve multiobjective problems \cite{DEB2002,HUBAND2006}.
+However, since there is a lack of open source Julia implementations of these alternative approaches, we did not benchmark the relative performance of JuPOETs in this note.
+One advantage that JuPOETs may have when compared to a strictly evolutionary approaches,
+is the inclusion of a local refinement step (hybrid mode), which temporarily reduces the problem to a single objective formulation.
+Previously, POETs run in hybrid mode led to better convergence on a proof-of-concept signal transduction model compared to the same approach without the hybrid refinement step \cite{Song:2010aa}.
+Other hybrid multiobjective methods have also been shown to be more efficient than evolutionary approaches alone, for a variety of biochemical optimization problems \cite{Otero-Muras2014,Sendin2006}.
+Thus, there are several different algorithms that we can use to benchmark, and improve the performance of JuPOETs, after we implement them in Julia.
+Another strategy to improve the performance of JuPOETs is to reduce the number (or cost) of function evaluations that are required to obtain optimal or near optimal solutions.
+For example, in many real world parameter estimation problems, the bulk of the execution time is spent evaluating the objective functions.
+One strategy to improve JuPOETs performance could be to optimize surrogates \cite{SURROGATES}, while another would be parallel execution of the objective functions.
 Currently, JuPOETs serially evaluates the objective function vector.
-However, parallel evaluation of the objective functions e.g., using the \texttt{\@parallel} Julia macro or other techniques,
-could be easily implemented without changing the main run loop of JuPOETs.
-Because of the flexible function pointer architecture of JuPOETs, the only changes required are in the user defined objective function.
-Taken together, JuPOETs has demonstrated improved flexibility, and performance over POETs in parameter identification and ensemble generation for multiple objectives.
-JuPOETs has the potential for widespread use due to the flexibility of the implementation, and the high level syntax and distribution tools native to Julia.
+However, parallel evaluation of the objective functions e.g., using the \texttt{\@parallel} Julia macro or other techniques, could be implemented without significantly changing the JuPOETs run loop.
+Taken together, JuPOETs demonstrated improved flexibility, and performance over POETs in parameter identification and ensemble generation for multiple objectives.
+JuPOETs has the potential for widespread use due to the flexibility of the implementation, and the high level syntax and distribution tools native to the Julia programming language.
 
@@
 %using a variety of techniques without alterating JuPOETs.
+%Because of the flexible function pointer architecture of JuPOETs, the only changes required are in the user defined objective function.
 

### 2016-11-18 response_to_reviewers/Paper_v2.tex
First round of reviews complete

@@
 patient specific coagulation behavior \cite{Luan:2010aa}, uncertainty quantification in metabolic kinetic models \cite{Andreozzi:2016aa} and to capture cell to cell variation \cite{Lequieu:2011aa}.
-Further, ensemble approaches have been used in synthetic biology to sample possible biocircuit configurations and to design folding behaviors of novel mRNAs \cite{Otero-Muras2014,Taneda2015}.
+Further, ensemble approaches have been used in synthetic biology to sample possible biocircuit configurations \cite{Otero-Muras:2014aa}.
 Thus, ensemble approaches are widely used to robustly simulate a variety of biochemical systems.
 
-Identification of biochemical models often requires significant training data perhaps taken from diverse sources.
-These data often can contain intrinsic conflicts resulting from, for example, the use of different cell lines or cross laboratory variability.
-Parameter ensembles that optimally balance these conflicts in training data can lead to more robust model performance.
-Multiobjective optimization is an ensemble generation technique that naturally balances conflicting training data \cite{Handl:2007aa}.
-Multiobjective optimization has been used us and others to identify biochemical models [REFHERE].
-Previously, we developed the Pareto Optimal Ensemble Technique (POETs) algorithm to address the challenge of competing or conflicting objectives.
-POETs, which integrates simulated annealing (SA) and multiobjective optimization through the notion of Pareto rank, estimates parameter ensembles which optimally trade-off between
-competing (and potentially conflicting) experimental objectives \cite{Song:2010aa}.
+Identification of biochemical models requires significant training data perhaps taken from diverse sources.
+These real-world data sets often contain intrinsic conflicts resulting from, for example, the use of different cell lines,
+different measurement technologies, different reagent vendors or lots, uncontrollable experimental artifacts or general cross laboratory variability.
+Parameter ensembles that optimally balance these inherent conflicts lead to more robust model performance.
+Multiobjective optimization is an ensemble generation technique that naturally balances conflicts in noisy training data \cite{Handl:2007aa}.
+Multiobjective optimization has been used to identify signal transduction models \cite{Lequieu:2011aa,Song:2009aa},
+for the design of synthetic circuits \cite{Otero-Muras:2014aa}, to design the folding behaviors of novel RNAs \cite{Taneda2015}, to design bioprocesses \cite{Banga2006},
+and to understand bacterial adaptation \cite{Angione:2015aa}.
+Thus, it is a widely used approach for a variety of biochemical applications.
+Previously, we developed the Pareto Optimal Ensemble Technique (POETs) algorithm to address the challenge of competing or conflicting training objectives.
+POETs, which integrates simulated annealing (SA) and multiobjective optimization through the notion of Pareto rank, estimates parameter ensembles which optimally
+trade-off between competing (and potentially conflicting) experimental objectives \cite{Song:2010aa}.
 However, the previous implementation of POETs, in the Octave programming language \cite{CITE_Octave}, suffered from poor performance and was not configurable.
@@
 In this software note, we present an open-source implementation of the Pareto optimal ensemble technique in the Julia programming language (JuPOETs).
-JuPOETs takes advantage of the unique features of Julia.
+JuPOETs takes advantage of the unique features of Julia to address many of the shortcomings of the previous implementation. 
 Julia is a cross-platform, high-performance programming language for technical computing that has performance comparable to C but with syntax similar to MATLAB/Octave and Python \cite{Julia}.
@@
 Previously, POETs run in hybrid mode led to better convergence on a proof-of-concept signal transduction model compared to the same approach without the hybrid refinement step \cite{Song:2010aa}.
-Other hybrid multiobjective methods have also been shown to be more efficient than evolutionary approaches alone, for a variety of biochemical optimization problems \cite{Otero-Muras2014,Sendin2006}.
+Other hybrid multiobjective methods have also been shown to be more efficient than evolutionary approaches alone, for a variety of biochemical optimization problems \cite{Otero-Muras:2014aa,Sendin2006}.
 Thus, there are several different algorithms that we can use to benchmark, and improve the performance of JuPOETs, after we implement them in Julia.

### 2016-11-21 response_to_reviewers/Paper_v2.tex
Submitted version with changes

@@
 
-JuPOETs integrates simulated annealing (SA) \cite{Kirkpatrick:1983aa} with Pareto ranking to estimate parameter sets on or near the optimal tradeoff surface between competing objectives (Fig.
+JuPOETs integrates simulated annealing (SA) \cite{Kirkpatrick:1983aa} with Pareto ranking to estimate decision variables on or near the optimal tradeoff surface between competing objectives (Fig.
 \ref{fig:fig-MOSAalgorithm} and Algorithm \ref{algo:JuPOETs-pcode}).
 A tradeoff surface defines the best possible performance for every conflicting objective, such that an increase in the performance of one objective does not decrease the performance of at least one other objective.
-Pareto rank is a measure of distance away from the optimal tradeoff surface.
-Thus, the central idea underlying POETs is a mapping between the value of the objective vector evaluated at $\mathbf{p}_{i+1}$ (parameter guess at iteration $i+1$) and Pareto rank (Fig.
+Pareto rank is a scalar measure of distance away from the optimal tradeoff surface (low rank is near the surface, while higher ranks are progressively further away).
+Thus, the central idea underlying POETs is a mapping between the value of the objective vector evaluated at $\mathbf{p}_{i+1}$ (decision variable guess at iteration $i+1$) and the scalar Pareto rank (Fig.
 \ref{fig:fig-MOSAalgorithm}).
+Traditional simulated annealing uses a scalar performance value e.g., simulation error to make a probabilistic decision to keep or reject a set of decision variables; decision variables with better performance are always accepted, while those with worse performance are sometimes accepted depending upon a parameter called the temperature.
+On the other hand, JuPOETs makes this same decision using the Pareto rank instead of a single performance objective.
+The problem of estimating biochemical model parameters from experimental data is typically posed as an error minimization problem over continuous real-valued decision variables (model parameters) subject to the model equations.
 A parameter set $\mathbf{p}_{i+1}$ lies along the optimal tradeoff surface if no other parameter guess leads to decreased error for every objective.
@@
 The error vector associated with $\mathbf{p}_{i+1}$ is ranked using the builtin Pareto \texttt{rank} function, by comparing the error at iteration $i+1$ to the error archive $\mathcal{O}_{i}$ (all error vectors up to iteration $i$ meeting a ranking criterion).
-Pareto rank is a measure of distance from the trade-off surface;
-parameter sets on or near the optimal trade-off surface between the objectives have a rank equal to $0$ (no other current parameter sets are better).
+Parameter sets on or near the optimal trade-off surface between the objectives have a rank equal to $0$ (no other current parameter sets are better).
 These rank zero parameter sets define the Pareto optimal group for the ensemble, wherein Pareto optimality is defined as a parameter set not being dominated by any other sets within the ensemble.
@@
 \end{equation}
-where $T$ is the computational annealing temperature that provides control over how strictly Pareto rank is enforced.
+where $T$ is the simulated annealing temperature; the temperature provides control over how strictly decreasing Pareto rank is enforced.
 As $\texttt{rank}\left(\mathcal{O}_{i+1}\left(\mathbf{p}_{i+1}\right)\mid \mathcal{O}_{i}\right)\rightarrow{0}$, the acceptance probability moves toward one,
@@
 Occasionally, (depending upon $T$) a parameter set with a high Pareto rank is  accepted by the SA allowing a more diverse search of the parameter space.
-However, as $T$ is reduced, the probability of accepting a high-rank set decreases.
+However, as $T$ is reduced as a function of iteration count (using the \texttt{cooling} function), the probability of accepting a high-rank set decreases.
 Parameter sets could also be accepted by the SA but $\emph{not}$ permanently archived in $\mathcal{S}_{i}$, where $\mathcal{S}_{i}$ is the solution archive.
@@
 To use JuPOETs, the user specifies the \texttt{neighbor}, \texttt{acceptance}, \texttt{cooling} and \texttt{objective} functions along with an initial decision variable guess.
-If the user is operating JuPOETs in hybrid mode, then a \texttt{refinement} function pointer must also be specified; the default \texttt{refinement} implementation is simply a pass through function.
 Default implementations of the \texttt{neighbor}, \texttt{acceptance} and \texttt{cooling} functions can be used directly, or they can be overridden by user defined logic.
 However, the user must provide an implementation of the \texttt{objective} function and provide an initial decision variable guess.
+Lastly, if the user is operating JuPOETs in hybrid mode, then a \texttt{refinement} function pointer must also be specified.
+Hybrid mode temporarily switches the search from a multiobjective to a single objective problem, where the sum of the objective functions can be used to update the best (or initial) parameter guess.
+The specific hybrid mode search logic is up to the user; by default hybrid mode is off, and the default \texttt{refinement} implementation is simply a pass through function.
+However, we have shown previously that POETs operated in hybrid mode (where the single objective problem used a pattern search approach)
+had better performance that POETs alone \cite{Song:2010aa}. Thus, hybrid mode is generally recommended for most applications.
 In addition, there are several user configurable parameters that can be adjusted to control the performance of JuPOETs:
@@
 columns of $\mathcal{S}$.
+One technical note, if JuPOETs is run from multiple starting locations, and the archives from each of these runs is combined into a single collective archive, the combined parameter rank archive may become invalid.
+In these cases, it is required to re-rank the parameter sets using the built-in \texttt{rank} function to produce a collective parameter ranking.
 
@@
 while simultaneously estimating parameter sets that performed well on each individual objective function.
+
+Currently, JuPOETs does not consider parameter identifiability when constructing parameter ensembles.
+Although JuPOETs produces parameter estimates that give model performance similar to the training data,
+we do not have strict statistical confidence that the \textit{true} parameter values are contained within the ensemble.
+However, despite this, ensembles produced by POETs can be predictive \cite{Lequieu:2011aa,Song:2009aa}.
+Thus, JuPOETs produces a collection of parameters that are constrained by the performance of the model,
+and not by specific hypotheses regarding the individual values of the raw model parameters.
+Of course, knowledge of specific parameter values, or the relationship between parameter combinations, can be used to inform the search through either bounds or problem specific constraints (for example, as demonstrated in the first example problem.)
 
